### 2012-11-21 paper.tex
deleted blindtext

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Short Sectioned Assignment
% LaTeX Template
% Version 1.0 (5/5/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%   PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[paper=a4, fontsize=11pt]{scrartcl} % A4 paper and 11pt font size

% philipp: added for code listings
\usepackage{listings}

\usepackage{natbib}
\bibliographystyle{plain}

\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\usepackage{fourier} % Use the Adobe Utopia font for the document - comment this line to return to the LaTeX default
\usepackage[english]{babel} % English language/hyphenation
\usepackage{amsmath,amsfonts,amsthm} % Math packages

\usepackage{lipsum} % Used for inserting dummy 'Lorem ipsum' text into the template
\usepackage{todonotes}

\usepackage{sectsty} % Allows customizing section commands
\allsectionsfont{\centering \normalfont\scshape} % Make all sections centered, the default font and small caps

\usepackage{fancyhdr} % Custom headers and footers
\pagestyle{fancyplain} % Makes all pages in the document conform to the custom headers and footers
\fancyhead{} % No page header - if you want one, create it in the same way as the footers below
\fancyfoot[L]{} % Empty left footer
\fancyfoot[C]{} % Empty center footer
\fancyfoot[R]{\thepage} % Page numbering for right footer
% \renewcommand{\thesection}{Data Set \arabic{section}}
\renewcommand{\headrulewidth}{0pt} % Remove header underlines
\renewcommand{\footrulewidth}{0pt} % Remove footer underlines
\setlength{\headheight}{13.6pt} % Customize the height of the header

\numberwithin{equation}{section} % Number equations within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{figure}{section} % Number figures within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{table}{section} % Number tables within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)

\setlength\parindent{0pt} % Removes all indentation from paragraphs - comment this line for an assignment with lots of text

%----------------------------------------------------------------------------------------
%   TITLE SECTION
%----------------------------------------------------------------------------------------

\newcommand{\horrule}[1]{\rule{\linewidth}{#1}} % Create horizontal rule command with 1 argument of height

\title{ 
\normalfont \normalsize 
\textsc{Vienna University of Technology} \\ [25pt] % Your university, school and/or department name(s)
\horrule{0.5pt} \\[0.4cm] % Thin top horizontal rule
\huge Experiments in Machine Learning \\ % The assignment title
\horrule{2pt} \\[0.5cm] % Thick bottom horizontal rule
}

\author{Benjamin Kiesl \and Philipp Steinwender \and Robert Sch\"{a}fer} % Your name

\date{\normalsize\today} % Today's date or a custom date

\begin{document}

\maketitle % Print the title

%----------------------------------------------------------------------------------------
%   PROBLEM 1
%----------------------------------------------------------------------------------------

\todo[inline]{Chose 5 different classifiers, from at least 3 different types
of learning algorithms

Argue \& justify choice (part of grading...)

Experiment with different parameter settings - And report on it !

Compare results among classifiers and datasets

Evaluate effect of pre-processing (e.g. different strategies
for missing values)
}

\section{Echocardiogram data}

The Echocardiogram data set consists of 132 entries with 13 attributes respectively and has 132 missing values in total. Originally, these entries referred to patients that suffered from a heart attack at some point in the past \cite{uci-repo}. The prediction task is to determine if the patient will survive at least one year.

\subsection{Preprocessing}
The name attribute of the patients was replaced by a dummy for all entries. For this reason, this attribute can be entirely removed. Another unnecessary attribute is ``group'' which is just meaningless and can be ignored \cite{uci-repo}. Another derived attribute is ``mult'' and can be ignored as well.

Some of the patients survived less than 12 months even if they were still alive at the end of the survey. In this case these patients joined the survey too recently to be followed at least one year. To make a reasonable prediction if a patient will survive at least one year, those entries have to be left out for the training data set because they would irritate any learning algorithm. \todo[inline]{maybe, give an example to show that those patients confuse a machine learning algorithm}

Yet, there are some entries, which can be wiped out from the whole data set as well:

\missingfigure{Show the crappy entry with the 76 years old patient who has only null values}

So long, these preprocessing tasks are just removing attributes (column-wise) or removing certain entries (row-wise), depending on their values.

Unfortunately, there is another peculiarity with the actual prediction attribute ``alive-at-1'', here is the official description:
\begin{quotation} ``Boolean-valued. Derived from the first two attributes. 0 means patient was either dead after 1 year or had been followed for less than 1 year.  1 means patient was alive at 1 year.''\cite{uci-repo} \end{quotation}
De facto, the values don't fit to this description. The attribute is $1$ if the patient was followed less than one year and $0$ otherwise. Very astonishing: Even if the attribute is supposed to be a derived attribute, there are several \texttt{null} values, making the corresponding values useless for training data.

\missingfigure{Show the crappy null valued entries}
The only way to amend this is to add the prediction value by hand: At first, all patients who joined the survey too recently, have to be left out (i.e. survived less than one year but still alive). Then, the ``alive-at-one'' attribute has to be overwritten with the correct value (i.e. $1$ if the patient survived at least one year).
\todo{Compare results with/without preprocessing}

After preprocessing only 95 of 132 instances remain and it is still a set full of gaps with 45 missing values in total. After preprocessing it is also a very imbalanced data set for machine learning: Only 4 patients died within 12 months, a ratio of 1 to 24. As the documentation of the UCI repository suggests ``The most difficult part of this problem is correctly predicting that the patient will NOT survive. (Part of the difficulty seems to be the size of the data set.)''\cite{uci-repo}. Indeed this is rather difficult to predict with only four valid instances.

\subsection{Decision Tree}
%------------------------------------------------
On the first glimpse, it is an easy task for a decision based algorithm:

\begin{verbatim}
J48 pruned tree
------------------

survival <= 12
|   survival <= 11: 0 (4.0)
|   survival > 11: 1 (4.0)
survival > 12: 1 (87.0)
\end{verbatim}

That's correct, if the patient survived $ <= 12 $ months, he survived at least 12 month. What a statement!

The only remarkable thing is, that the J48 implementation under default settings seems to enforce a certain depth of the tree, even if the splitting on the same attribute don't make any sense ($ 11 $ is yet a smaller number than $ 12 $, isn't it?). E.g. choosing a DecisionStump results in the expected single splitting: 
\begin{verbatim}
Class distributions
survival <= 11.5
0   1   
1.0 0.0 
survival > 11.5
0   1   
0.0 1.0 
survival is missing
0   1   
0.042105263157894736    0.9578947368421052
\end{verbatim}
Since there is no missing value for the survival attribute, this branch can be ignored.

But allowing the machine learning algorithms to see the origin attributes from which we derived the prediction attribut leads to absurdity, since the actual prediction task is to say whether a patient will survive one year after suffering a heart attack or not \emph{depending on the medical characteristics}.

\subsection{re-pre-processed the data set}

As mentioned in the description of the dataset on the UCI repo site we can remove the attributes group, name and mold as they do not have any meaning for the prediction attribute.

We also have to remove the survival and still-alive attributes, as the prediction attribute derives from them. Otherwise, a classification algorithm could predict all instances by taking into account only these two attributes.

A further refinement of the data set has been done by deleting all instances, where the outcome was unknown, as this cannot be used for training.

\section{Adult data}

The Adult or Census Income data sets contains 48842 instances of information about persons and if they earn more than 50.000 a year. The data set has 14 attributes and missing values.

The instances origin from the 1994 Census Database and were extracted by Barry Becker. For having a clean data set, only records with age above 16 with minimium one working hour per week have been selected from the database.

\subsection{Preprocessing}

We can remove the final weight -- but we have to find out what it means !!.

We can also remove the $education-num$ attribute as it is only a numeric representation of the education attribute.

At first we try to discretize all numeric attributes with 10 bins of equal ranges.

\subsection{Decision Tree}

\subsection{Random Forrest}

\subsection{Support Vector Machine}

\subsection{Naive Bayes}

We use 66\% of the data set for training and the rest for evaluation.

For the first run, we use the default setting and get the following result:

\begin{verbatim}
Correctly Classified Instances        9104               82.2329 %
Incorrectly Classified Instances      1967               17.7671 %
Total Number of Instances            11071

Precision   Recall  F-Measure  Class
 0.592       0.786   0.675      >50K
 0.927       0.834   0.878      <=50K
\end{verbatim}

Second run: we use the discretizer with 20 bin.

\begin{verbatim}
Correctly Classified Instances        9287               83.8858 %
Incorrectly Classified Instances      1784               16.1142 %
Total Number of Instances            11071

=== Detailed Accuracy By Class ===

Precision   Recall  F-Measure  Class
 0.621       0.805   0.701      >50K
 0.934       0.849   0.89       <=50K
\end{verbatim}

As expected, the result is slightly better.

A better result should be possible if we let the dicretizer split the values in part of nearly equal size. The result:

\begin{verbatim}
Correctly Classified Instances        9179               82.9103 %
Incorrectly Classified Instances      1892               17.0897 %
Total Number of Instances            11071

Precision   Recall  F-Measure  Class
0.602        0.805   0.689      >50K
0.933        0.836   0.882      <=50K
\end{verbatim}

It got worse. \todo{why does discretizer with equal bin size not give a better result?}

The Naive Bayes does not have a lot of settings. It use a built in supervised discretizer or a kernel estimator. But the result is always the same with around 80\% correctly classified instances in the test run.

\subsection{k-Nearest}

\section{Contraceptive Method Choice Data}

The Contraceptive Method Choice Data Set consists of information about the demographic and socio-economic characteristics of married women. The task is to predict which kind of contraceptive methods a woman uses. These are categorized as no use, short term or long term.

\subsection{Preprocessing}

The attributes are already in a numerical and categorical form, whereby the categories are encoded in an ordinal form. These go from 1 to 4 where 1 is low/bad and 4 is high/good.

\todo{check if ordinal is the correct term when i mean, that the attributes are encoded in numbers, where 1<2, 1<4, 4>2, ...}

\subsection{Desicion Tree}
\subsection{Random Forrest}
\subsection{Naive Bayes}
\subsection{Support VectorMachine}
\subsection{k-Nearest}

\bibliography{references}

\end{document}
%------------------------------------------------

### 2012-11-21 paper.tex
worked on decision trees for echocardiogramm

@@
 \usepackage{natbib}
+\usepackage{listings}
 \bibliographystyle{plain}
@@
 
-\begin{verbatim}
+\begin{lstlisting}
 J48 pruned tree
@@
 survival > 12: 1 (87.0)
-\end{verbatim}
+\end{lstlisting}
 
@@
 choosing a DecisionStump results in the expected single splitting:
-\begin{verbatim}
+\begin{lstlisting}
 Class distributions
@@
 0.042105263157894736    0.9578947368421052
-\end{verbatim}
+\end{lstlisting}
 Since there is no missing value for the survival attribute, this branch can be ignored.
@@
 
-\subsection{re-pre-processed the data set}
+\subsection{re-preprocessed the data set}
 
-As mentioned in the description of the dataset on the UCI repo site we can remove the attributes group, name and mold as they do not have any meaning for the prediction attribute.
+As mentioned in the description of the dataset on the UCI repository site we can remove the attributes ``group'', ``name'' and ``mult'' as they do not have any meaning for the prediction attribute.
 
-We also have to remove the survival and still-alive attributes, as the prediction attribute derives from them.
+We also have to remove the ``survival'' and ``still-alive'' attributes, as the prediction attribute derives from them.
 Otherwise, a classification algorithm could predict all instances by taking into account only these two attributes.
@@
 
+\subsection{Decision Tree on re-preprocessed data}
+
+The J48 Decision Tree shows a more useful tree on the re-preprocessed data:
+\begin{lstlisting}
+J48 pruned tree
+------------------
+
+epss <= 13100: 1 (90.63/1.95)
+epss > 13100
+|   fractional-shortening <= 190: 1 (2.25/0.05)
+|   fractional-shortening > 190: 0 (2.11/0.11)
+\end{lstlisting}
+
+The documentation of the data set explains the two attributes as follows:
+\begin{quote}
+``epss -- E-point septal separation, another measure of contractility. Larger numbers are increasingly abnormal.''
+\end{quote}
+\begin{quote}
+`` fractional-shortening -- a measure of contracility around the heart lower numbers are increasingly abnormal''
+\end{quote}
+
+Only the rule for the fractional shortening should be the other way round, because a low fractional shortening is a rather pathological sign.
+Although the decision tree has a high precision of almost 95\% it performs very imprecise, because none of the patients who died after one year are correctly classified:
+
+\begin{lstlisting}
+=== Confusion Matrix ===
+
+  a  b   <-- classified as
+  0  4 |  a = 0
+  1 90 |  b = 1
+
+\end{lstlisting}
+
+Using an unpruned tree doesn't improve the results:
+\begin{lstlisting}
+J48 unpruned tree
+------------------
+
+pericardial-effusion = 0
+|   epss <= 13100: 1 (76.62/1.95)
+|   epss > 13100
+|   |   epss <= 15600: 0 (2.19/0.16)
+|   |   epss > 15600: 1 (2.19/0.03)
+pericardial-effusion = 1: 1 (14.0)
+\end{lstlisting}
+Here, the existence of a pericardial effusion (fluid around the heart) is assumed to be a healthy characteristic, but it is not. Accordingly, the precision of the J48 decreases from 95\% to 88\%.
 \section{Adult data}

### 2012-11-21 paper.tex

@@
 
-\missingfigure{Show the crappy null valued entries}
-The only way to amend this is to add the prediction value by hand: At first, all patients who joined the survey too recently, have to be left out (i.e.
-survived less than one year but still alive).
-Then, the ``alive-at-one'' attribute has to be overwritten with the correct value (i.e.
-$1$ if the patient survived at least one year).
+We thus inferred the prediction attribute ``alive-at-1'' as explained in the description of the whole dataset: For all patients where the months of survival are greater than or equal to 12 we set the ``alive-at-1''-attribute to 1, otherwise we set it to 0.
+
+The only thing which remained then was to a
+
 \todo{Compare results with/without preprocessing}
-
-After preprocessing only 95 of 132 instances remain and it is still a set full of gaps with 45 missing values in total.
-After preprocessing it is also a very imbalanced data set for machine learning: Only 4 patients died within 12 months, a ratio of 1 to 24.
-As the documentation of the UCI repository suggests ``The most difficult part of this problem is correctly predicting that the patient will NOT survive.
-(Part of the difficulty seems to be the size of the data set.)''\cite{uci-repo}.
-Indeed this is rather difficult to predict with only four valid instances.
 

### 2012-11-22 paper.tex
asdf
finished decision based algorithms on echocardiogram data so far -- simple decision tree deserves a little more detailed evaluation

@@
 %----------------------------------------------------------------------------------------
+
+\todo[inline]{defined:
+
+validation: cross-validation with 10 for big data sets (9 training, 1 test), 3 for the small echo cardio set.
+}
 
@@
 J48 pruned tree
-------------------
 
@@
 === Confusion Matrix ===
-
   a  b   <-- classified as
@@
 J48 unpruned tree
-------------------
 
@@
 Here, the existence of a pericardial effusion (fluid around the heart) is assumed to be a healthy characteristic, but it is not. Accordingly, the precision of the J48 decreases from 95\% to 88\%.
+
+\subsubsection{Random Forest}
+Let's see if a random forest in place of a simple decision tree can make better result out of the echocardiogram data set.
+With default settings, i.e.
+default number of features ($ log_2(7) + 1 = 2 $) and 10 random trees with unlimited tree depth, the random forest shows us a rather disappointing result:
+
+\begin{lstlisting}
+=== Confusion Matrix ===
+  a  b   <-- classified as
+  0  4 |  a = 0
+  2 91 |  b = 1
+\end{lstlisting}
+And the statistical measures:
+\begin{lstlisting}
+=== Detailed Accuracy By Class ===
+                 TP Rate  FP Rate  Precision  Recall  F-Measure    Class
+                 0        0.022    0          0       0            0
+                 0.978    1        0.958      0.978   0.968        1
+Weighted Avg.    0.938    0.96     0.918      0.938   0.928     
+\end{lstlisting}
+\paragraph{Parameter: Number of trees}
+\todo[inline]{proper line break here}
+Reducing the number of the trees continuously decreases the quality of the model, as shown in the accuracy details for 5 random trees:
+\begin{lstlisting}
+=== Detailed Accuracy By Class ===
+                 TP Rate  FP Rate  Precision  Recall  F-Measure    Class
+                 0        0.032    0          0       0            0  
+                 0.968    1        0.957      0.968   0.963        1   
+Weighted Avg.    0.928    0.96     0.918      0.928   0.923    
+\end{lstlisting}
+.. and for a single random tree:
+\begin{lstlisting}
+=== Detailed Accuracy By Class ===
+                 TP Rate  FP Rate  Precision  Recall  F-Measure    Class
+                 0        0.054    0          0       0            0  
+                 0.946    1        0.957      0.946   0.951        1   
+Weighted Avg.    0.907    0.961    0.917      0.907   0.912    
+\end{lstlisting}
+Reducing the number of the trees increases the number of false positives but doesn't catch any true positives.
+\paragraph{Limiting the tree-depth}
+Limitations in the depths of the random trees has also almost no effect on the accuracy of random forests on this data set.
+The only change to the outcome can be observed as soon the depth of the trees is limited to 6:
+\begin{lstlisting}
+    === Confusion Matrix ===
+    a  b   <-- classified as
+    0  4 |  a = 0
+    1 92 |  b = 1
+\end{lstlisting}
+Only one false positive has gone. Not really a remarkable change in the quality. This does not change if the tree-depth is further reduced.
+\paragraph{Increasing the number of features}
+Increasing the number of features chosen by each random tree seems to has also a low effect on the quality, because until a certain border of 5 chosen features the outcome is exactly the same as it was with default settings.
+But if we configure the random forests with 6 features we can spot our first true positive:
+\begin{lstlisting}
+=== Confusion Matrix ===
+  a  b   <-- classified as
+  1  3 |  a = 0
+  2 91 |  b = 1
+\end{lstlisting}
+Which gives us the best achievable accuracy that we can reach with decision based algorithms:
+\begin{lstlisting}
+=== Detailed Accuracy By Class ===
+                TP Rate  FP Rate  Precision  Recall  F-Measure   Class
+                0.25     0.022    0.333      0.25    0.286       0
+                0.978    0.75     0.968      0.978   0.973       1
+Weighted Avg.   0.948    0.72     0.942      0.948   0.945      
+\end{lstlisting}
 \section{Adult data}

### 2012-11-22 paper.tex
added chapter decision tree for the adult data set

@@
 \subsection{Decision Tree}
+Using a decision tree yields a precision about 86\%. Varying the parameters for this algorithm influences the quality of the results only slightly, as the following examples illustrate.
+
+First of all, with the default settings of the J48 implementation of weka, i.e.
+using pruning with a confidence factor of 0.25 together with MDLCorrection and Subtree raising and at least 2 instances per leaf of the tree, we get the following summary of the quality:
+
+\begin{lstlisting}
+=== Detailed Accuracy By Class ===
+                TP Rate  FP Rate  Precision  Recall  F-Measure  Class
+                0.615    0.062    0.759      0.615   0.68       >50K
+                0.938    0.385    0.885      0.938   0.911      <=50K
+Weighted Avg.   0.86     0.307    0.854      0.86    0.855      
+\end{lstlisting}
+
+The decision tree that was used for this run has a size of 413 and has 311 leaves.
+
+\paragraph{Raising the confidence factor}
+Subsequently increasing the confidence factor, which result in less pruning, leads to a continuous worse quality. If we set the confidence factor to 0.5 we get the following outcome:
+\begin{lstlisting}
+=== Detailed Accuracy By Class ===
+                TP Rate  FP Rate  Precision  Recall  F-Measure  Class
+                0.632    0.071    0.738      0.632   0.681      >50K
+                0.929    0.368    0.888      0.929   0.908      <=50K
+Weighted Avg.   0.857    0.296    0.852      0.857   0.854      
+\end{lstlisting}
+.. for a confidence factor of 0.75 we get the following:
+\begin{lstlisting}
+=== Detailed Accuracy By Class ===
+                TP Rate  FP Rate  Precision  Recall  F-Measure  Class
+                0.655    0.094    0.689      0.655   0.672      >50K
+                0.906    0.345    0.892      0.906   0.899      <=50K
+Weighted Avg.   0.846    0.285    0.843      0.846   0.844     
+\end{lstlisting}
+
+\paragraph{Other parameters}
+Varying other parameters has a similar effect like the one mentioned above.
+Either the quality of the J48 algorithm decreases a little or it stays the same, thus never exceeding an overall quality of 86\%.
+To sum it up: Using an unpruned tree yields the same quality as a tree with a confidence factor of 0.75.
+Smoothing the counts at the leaves with laplace has no effect.
+Either without MDLcorrection or without Subtree raising during pruning, in both cases the precision falls from 86\% to 85.5\%.
+If we rise the minimum of instances per leaves, e.g.
+to 20 instances and 50 instances, the precision falls from 86\% to 85\% and 84\$ respectively.
 

### 2012-11-23 paper.tex
finished chapter random forests on census income  -- last commit in the train to cologne, going to sleep now

@@
 With default settings, i.e.
-default number of features ($ log_2(7) + 1 = 2 $) and 10 random trees with unlimited tree depth, the random forest shows us a rather disappointing result:
+default number of features \todo[inline]{calculate!} and 10 random trees with unlimited tree depth, the random forest shows us a rather disappointing result:
 
@@
 
+With default settings, a random forest achieves an accuracy as follows:
+\begin{lstlisting}
+=== Detailed Accuracy By Class ===
+                TP Rate  FP Rate  Precision  Recall  F-Measure  Class
+                0.615    0.108    0.644      0.615   0.629      >50K
+                0.892    0.385    0.88       0.892   0.886      <=50K
+Weighted Avg.   0.825    0.318    0.823      0.825   0.824     
+\end{lstlisting}
+which is surprisingly a worse result than from the J48 algorithm.
+
+\paragraph{Number of Random Trees}
+As expected, reducing the number of random trees reduces the quality, with 5 random trees the precision lays at 82.6\% with only 3 random trees it lays at 82.3\%.
+Interestingly, the quality gets higher if we take even more than 10 random trees into account: With 15 trees, the precision is at 83.2\% and with 25 random trees it is at 83.3\%.
+
+\paragraph{Limitations of the maximum depth of the trees}
+We have a sign for overfitting under default settings: Reducing the maximum depth has a positive effect on the quality.
+From size 20 over 10 until the size of 7 the precision of the random forest rises until 85\%.
+Reducing the depth even further lets the precision fall again.
+Here is the detailed accuracy for a random forest which can grow only to a height of 6:
+
+\begin{lstlisting}
+=== Detailed Accuracy By Class ===
+                TP Rate  FP Rate  Precision  Recall  F-Measure  Class
+                0.606    0.064    0.749      0.606   0.67       >50K
+                0.936    0.394    0.882      0.936   0.908      <=50K
+Weighted Avg.   0.856    0.314    0.85       0.856   0.851     
+\end{lstlisting}
+We can push the accuracy even further, if we take more random trees, e.g. twenty:
+\begin{lstlisting}
+=== Detailed Accuracy By Class ===
+                TP Rate  FP Rate  Precision  Recall  F-Measure  Class
+                0.593    0.058    0.763      0.593   0.667      >50K
+                0.942    0.407    0.879      0.942   0.909      <=50K
+Weighted Avg.   0.858    0.323    0.851      0.858   0.851      
+\end{lstlisting}
+
+\paragraph{Number of features}
+Increasing the number of features doesn't support the quality at all.
+The default number of features for this data set with 13 attributes is \todo[inline]{calculate!}.
+With a number of 5 randomly chosen features, the precision is yet at 85\% (with maximum depth of tree = 6) but with 10 randomly chosen features the precision falls back to 84.5\%.
 \subsection{Support Vector Machine}

### 2012-11-23 paper.tex
Signed-off-by: schubiduu <b.kiesl@atn.nu>

@@
 
-\todo[inline]{Chose 5 different classifiers, from at least 3 different types
+\todo[inline]{Choose 5 different classifiers, from at least 3 different types
 of learning algorithms
@@
 
-The Echocardiogram data set consists of 132 entries with 13 attributes respectively and has 132 missing values in total.
-Originally, these entries referred to patients that suffered from a heart attack at some point in the past \cite{uci-repo}.
-The prediction task is to determine if the patient will survive at least one year.
+The Echocardiogram data set originally consists of 132 entries with 13 attributes respectively.
+The entries of the dataset refer to patients that suffered from a heart attack at some point in the past \cite{uci-repo}.
+The prediction task is to determine from several different attributes if the patient will survive at least one year, beginning from the day when the heart attack occured.
+The dataset contains missing values and there's a lot of other preprocessing steps (described later) which have to be done.
+The dataset is very small, especially after preprocessing and a further problem is that the number of patients which died within one year after their heart attack is very small (4) which makes the classification very hard.
+It is nevertheless an interesting question wether some classification algorithm can perform well on that data.
 
 \subsection{Preprocessing}
-The name attribute of the patients was replaced by a dummy for all entries.
-For this reason, this attribute can be entirely removed.
-Another unnecessary attribute is ``group'' which is just meaningless and can be ignored \cite{uci-repo}.
-Another derived attribute is ``mult'' and can be ignored as well.
+\paragraph{}The patient's attribute ``name'' was replaced by a dummy for all entries.
+For this reason, this attribute could be entirely removed.
+Other unnecessary attributes were ``group'' and ``mult'' which were also (as explained in the description) just meaningless and could thus be ignored \cite{uci-repo}.
+There were also two attributes ``wall-motion-score'' and ``wall-motion-index'' where the description of the dataset said that one should only use the ``wall-motion-index'' for classification issues and so ``wall-motion-score'' was also removed.
 
-Some of the patients survived less than 12 months even if they were still alive at the end of the survey.
-In this case these patients joined the survey too recently to be followed at least one year.
-To make a reasonable prediction if a patient will survive at least one year, those entries have to be left out for the training data set because they would irritate any learning algorithm.
-\todo[inline]{maybe, give an example to show that those patients confuse a machine learning algorithm}
+\paragraph{}The actual prediction attribute ``alive-at-1'' (denoting that a patient lived at least one year after having a heart attack) had to be inferred from two other attributes, namely ``survival'' (number of months that the test-person survived after the heart attack) and ``still-alive'' (a binary variable denoting if a patient is still alive at the time when the data was recorded).
+It is possible that the ``survival'' variable has a value lower than 12 but the ``still-alive'' variable is set to true, meaning that the test-person still lived.
+This is the case if the data for a test-person was recorded in the last 12 months and so this data cannot be used for the classification and was thus removed.
 
-Yet, there are some entries, which can be wiped out from the whole data set as well:
+\paragraph{}It should be clear that the variables ``survival'' and ``still-alive'' should not be used for classification since they constitute the final prediction attribute.
 
-\missingfigure{Show the crappy entry with the 76 years old patient who has only null values}
+\paragraph{}The last preprocessing step was the replacement of missing values (data imputation).
+Most of the data was complete but there were still some data records having missing values which we replaced by the mean values of the specific variables.
+One data record did contain completeley useless data and was thus removed (see at the picture below).
 
-So long, these preprocessing tasks are just removing attributes (column-wise) or removing certain entries (row-wise), depending on their values.
+\paragraph{}
 
-Unfortunately, there is another peculiarity with the actual prediction attribute ``alive-at-1'', here is the official description:
-\begin{quotation} ``Boolean-valued.
+%So long, these preprocessing tasks are just removing attributes (column-wise) or removing certain entries (row-wise), depending on their values.
+%
+%Unfortunately, there is another peculiarity with the actual prediction attribute ``alive-at-1'', here is the official description:
+%\begin{quotation} ``Boolean-valued.
 Derived from the first two attributes.
@@
 1 means patient was alive at 1 year.''\cite{uci-repo} \end{quotation}
-De facto, the values don't fit to this description.
+%De facto, the values don't fit to this description.
 The attribute is $1$ if the patient was followed less than one year and $0$ otherwise.
 Very astonishing: Even if the attribute is supposed to be a derived attribute, there are several \texttt{null} values, making the corresponding values useless for training data.
+%
+%We thus inferred the prediction attribute ``alive-at-1'' as explained in the description of the whole dataset: For all patients where the months of survival are greater than or equal to 12 we set the ``alive-at-1''-attribute to 1, otherwise we set it to 0.
 
-We thus inferred the prediction attribute ``alive-at-1'' as explained in the description of the whole dataset: For all patients where the months of survival are greater than or equal to 12 we set the ``alive-at-1''-attribute to 1, otherwise we set it to 0.
-
-The only thing which remained then was to a
+%The only thing which remained then was to a
 
@@
 But allowing the machine learning algorithms to see the origin attributes from which we derived the prediction attribut leads to absurdity, since the actual prediction task is to say whether a patient will survive one year after suffering a heart attack or not \emph{depending on the medical characteristics}.
-
-\subsection{re-preprocessed the data set}
-
-As mentioned in the description of the dataset on the UCI repository site we can remove the attributes ``group'', ``name'' and ``mult'' as they do not have any meaning for the prediction attribute.
-
-We also have to remove the ``survival'' and ``still-alive'' attributes, as the prediction attribute derives from them.
-Otherwise, a classification algorithm could predict all instances by taking into account only these two attributes.
-
-A further refinement of the data set has been done by deleting all instances, where the outcome was unknown, as this cannot be used for training.
 
@@
 \end{lstlisting}
+
+\subsection{Support Vector Machine}
+
+\paragraph{}We used the support vector machine implementation called `SMO' in WEKA.
+Since a support vector machine uses the numeric coordinates of the data records to do the classification we tried to normalize the data in advance to obtain better results.
+Since this normalization did not have any impact on the results we followed that the support vector machine algorithm of the WEKA-toolkit automatically normalizes all the data records.
+
+\paragraph{}The build of the model took no time since the dataset is that small.
+The same holds for the evaluation.
+We tried all the different kernel-functions for the classification but the result didn't change.
+The following table shows the results of the evaluation and if you take a close look on the results, you can see that all data records were classified to the same class which means that the SVM algorithm wasn't able to cope with the small dataset:
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|lllllll|}
+	\hline
+	TP Rate   			&FP Rate   &Precision   &Recall  &F-Measure   &ROC Area  &Class\\
+	\hline
+	0         			&0         &0         	&0       &  0         & 0.5      &0\\
+  1         			&1         &0.959     	&1       &0.979       &	0.5       &1\\
+  \hline
+	Weighted Avg.		&0.959     &0.959      	&0.919   &0.959     	&0.939     &0.5\\
+	\hline
+\end{tabular}
+\caption{Results of the SVM-classification for the echocardiogram dataset}
+\end{table}
+
+\vspace{6pt}
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|lll|}
+	\hline
+	Classified as	& &\\
+	a		&	b & \\
+	\hline
+	0   & 4  	&  a = 0\\
+  0   & 93 	&  b = 1\\
+  \hline
+\end{tabular}
+\caption{Confusion matrix}
+\end{table}
+
+\begin{lstlisting}
+
+\end{lstlisting}
+
 \section{Adult data}
@@
 
-We can remove the final weight -- but we have to find out what it means !!.
-
-We can also remove the $education-num$ attribute as it is only a numeric representation of the education attribute.
-
-At first we try to discretize all numeric attributes with 10 bins of equal ranges.
+The ``Adult'' dataset has missing values which we all replaced by the mean value of the respective attribute.
+We did no further preprocessing.
+As for the Echocardiogram dataset there was no normalization neccessary for the SVM or the kNN algorithms.
 

### 2012-11-24 paper.tex
note on validation

--- 2012-11-22 paper.tex
@@
 
-We thus inferred the prediction attribute ``alive-at-1'' as explained in the description of the whole dataset: For all patients where the months of survival are greater than or equal to 12 we set the ``alive-at-1''-attribute to 1, otherwise we set it to 0.
+\missingfigure{Show the crappy null valued entries}
+The only way to amend this is to add the prediction value by hand: At first, all patients who joined the survey too recently, have to be left out (i.e.
+survived less than one year but still alive).
+Then, the ``alive-at-one'' attribute has to be overwritten with the correct value (i.e.
+$1$ if the patient survived at least one year).
+\todo{Compare results with/without preprocessing}
 
-The only thing which remained then was to a
-
-\todo{Compare results with/without preprocessing}
+After preprocessing only 95 of 132 instances remain and it is still a set full of gaps with 45 missing values in total.
+After preprocessing it is also a very imbalanced data set for machine learning: Only 4 patients died within 12 months, a ratio of 1 to 24.
+As the documentation of the UCI repository suggests ``The most difficult part of this problem is correctly predicting that the patient will NOT survive.
+(Part of the difficulty seems to be the size of the data set.)''\cite{uci-repo}.
+Indeed this is rather difficult to predict with only four valid instances.
 
@@
 === Confusion Matrix ===
+
   a  b   <-- classified as
@@
 Here, the existence of a pericardial effusion (fluid around the heart) is assumed to be a healthy characteristic, but it is not. Accordingly, the precision of the J48 decreases from 95\% to 88\%.
-
-\subsubsection{Random Forest}
-Let's see if a random forest in place of a simple decision tree can make better result out of the echocardiogram data set.
-With default settings, i.e.
-default number of features ($ log_2(7) + 1 = 2 $) and 10 random trees with unlimited tree depth, the random forest shows us a rather disappointing result:
-
-\begin{lstlisting}
-=== Confusion Matrix ===
-  a  b   <-- classified as
-  0  4 |  a = 0
-  2 91 |  b = 1
-\end{lstlisting}
-And the statistical measures:
-\begin{lstlisting}
-=== Detailed Accuracy By Class ===
-                 TP Rate  FP Rate  Precision  Recall  F-Measure    Class
-                 0        0.022    0          0       0            0
-                 0.978    1        0.958      0.978   0.968        1
-Weighted Avg.    0.938    0.96     0.918      0.938   0.928     
-\end{lstlisting}
-\paragraph{Parameter: Number of trees}
-\todo[inline]{proper line break here}
-Reducing the number of the trees continuously decreases the quality of the model, as shown in the accuracy details for 5 random trees:
-\begin{lstlisting}
-=== Detailed Accuracy By Class ===
-                 TP Rate  FP Rate  Precision  Recall  F-Measure    Class
-                 0        0.032    0          0       0            0  
-                 0.968    1        0.957      0.968   0.963        1   
-Weighted Avg.    0.928    0.96     0.918      0.928   0.923    
-\end{lstlisting}
-.. and for a single random tree:
-\begin{lstlisting}
-=== Detailed Accuracy By Class ===
-                 TP Rate  FP Rate  Precision  Recall  F-Measure    Class
-                 0        0.054    0          0       0            0  
-                 0.946    1        0.957      0.946   0.951        1   
-Weighted Avg.    0.907    0.961    0.917      0.907   0.912    
-\end{lstlisting}
-Reducing the number of the trees increases the number of false positives but doesn't catch any true positives.
-\paragraph{Limiting the tree-depth}
-Limitations in the depths of the random trees has also almost no effect on the accuracy of random forests on this data set.
-The only change to the outcome can be observed as soon the depth of the trees is limited to 6:
-\begin{lstlisting}
-    === Confusion Matrix ===
-    a  b   <-- classified as
-    0  4 |  a = 0
-    1 92 |  b = 1
-\end{lstlisting}
-Only one false positive has gone. Not really a remarkable change in the quality. This does not change if the tree-depth is further reduced.
-\paragraph{Increasing the number of features}
-Increasing the number of features chosen by each random tree seems to has also a low effect on the quality, because until a certain border of 5 chosen features the outcome is exactly the same as it was with default settings.
-But if we configure the random forests with 6 features we can spot our first true positive:
-\begin{lstlisting}
-=== Confusion Matrix ===
-  a  b   <-- classified as
-  1  3 |  a = 0
-  2 91 |  b = 1
-\end{lstlisting}
-Which gives us the best achievable accuracy that we can reach with decision based algorithms:
-\begin{lstlisting}
-=== Detailed Accuracy By Class ===
-                TP Rate  FP Rate  Precision  Recall  F-Measure   Class
-                0.25     0.022    0.333      0.25    0.286       0
-                0.978    0.75     0.968      0.978   0.973       1
-Weighted Avg.   0.948    0.72     0.942      0.948   0.945      
-\end{lstlisting}
 \section{Adult data}
@@
 \subsection{Decision Tree}
-Using a decision tree yields a precision about 86\%. Varying the parameters for this algorithm influences the quality of the results only slightly, as the following examples illustrate.
-
-First of all, with the default settings of the J48 implementation of weka, i.e.
-using pruning with a confidence factor of 0.25 together with MDLCorrection and Subtree raising and at least 2 instances per leaf of the tree, we get the following summary of the quality:
-
-\begin{lstlisting}
-=== Detailed Accuracy By Class ===
-                TP Rate  FP Rate  Precision  Recall  F-Measure  Class
-                0.615    0.062    0.759      0.615   0.68       >50K
-                0.938    0.385    0.885      0.938   0.911      <=50K
-Weighted Avg.   0.86     0.307    0.854      0.86    0.855      
-\end{lstlisting}
-
-The decision tree that was used for this run has a size of 413 and has 311 leaves.
-
-\paragraph{Raising the confidence factor}
-Subsequently increasing the confidence factor, which result in less pruning, leads to a continuous worse quality. If we set the confidence factor to 0.5 we get the following outcome:
-\begin{lstlisting}
-=== Detailed Accuracy By Class ===
-                TP Rate  FP Rate  Precision  Recall  F-Measure  Class
-                0.632    0.071    0.738      0.632   0.681      >50K
-                0.929    0.368    0.888      0.929   0.908      <=50K
-Weighted Avg.   0.857    0.296    0.852      0.857   0.854      
-\end{lstlisting}
-.. for a confidence factor of 0.75 we get the following:
-\begin{lstlisting}
-=== Detailed Accuracy By Class ===
-                TP Rate  FP Rate  Precision  Recall  F-Measure  Class
-                0.655    0.094    0.689      0.655   0.672      >50K
-                0.906    0.345    0.892      0.906   0.899      <=50K
-Weighted Avg.   0.846    0.285    0.843      0.846   0.844     
-\end{lstlisting}
-
-\paragraph{Other parameters}
-Varying other parameters has a similar effect like the one mentioned above.
-Either the quality of the J48 algorithm decreases a little or it stays the same, thus never exceeding an overall quality of 86\%.
-To sum it up: Using an unpruned tree yields the same quality as a tree with a confidence factor of 0.75.
-Smoothing the counts at the leaves with laplace has no effect.
-Either without MDLcorrection or without Subtree raising during pruning, in both cases the precision falls from 86\% to 85.5\%.
-If we rise the minimum of instances per leaves, e.g.
-to 20 instances and 50 instances, the precision falls from 86\% to 85\% and 84\$ respectively.
 

### 2012-11-24 paper.tex
Added a LOT of things

--- 2012-11-23 paper.tex
@@
 It is nevertheless an interesting question wether some classification algorithm can perform well on that data.
+Since it is very small we performed a 3-fold-cross-validation on this dataset.
 
@@
 Since a support vector machine uses the numeric coordinates of the data records to do the classification we tried to normalize the data in advance to obtain better results.
-Since this normalization did not have any impact on the results we followed that the support vector machine algorithm of the WEKA-toolkit automatically normalizes all the data records.
+Since this normalization did not have any impact on the results (we tried this not only on this dataset) we followed that the support vector machine algorithm of the WEKA-toolkit automatically normalizes all the numeric values.
 
-\paragraph{}The build of the model took no time since the dataset is that small.
+\paragraph{}The build of the model took not even a second since the dataset is that small.
 The same holds for the evaluation.
 We tried all the different kernel-functions for the classification but the result didn't change.
-The following table shows the results of the evaluation and if you take a close look on the results, you can see that all data records were classified to the same class which means that the SVM algorithm wasn't able to cope with the small dataset:
+The following table shows the results of the evaluation and if you take a close look on the results, you can see that all data records were classified to the same class which means that the SVM algorithm wasn't able to cope with the small number of data records which should be classified with value 0:
 
@@
 \centering
-\begin{tabular}{|lllllll|}
+\begin{tabular}{|llllllll|}
 	\hline
-	TP Rate   			&FP Rate   &Precision   &Recall  &F-Measure   &ROC Area  &Class\\
+									&TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure   &ROC Area  &Class\\
 	\hline
-	0         			&0         &0         	&0       &  0         & 0.5      &0\\
-  1         			&1         &0.959     	&1       &0.979       &	0.5       &1\\
+									&0         	&0         &0         	&0       &  0         & 0.5      &0\\
+  								&1         	&1         &0.959     	&1       &0.979       &	0.5      &1\\
   \hline
-	Weighted Avg.		&0.959     &0.959      	&0.919   &0.959     	&0.939     &0.5\\
+	Weighted Avg.		&0.959     	&0.959     &0.919   		&0.959   &0.939     	&0.5	&\\
 	\hline
@@
 \end{tabular}
-\caption{Confusion matrix}
+\caption{Confusion matrix for SVM}
 \end{table}
 
-\begin{lstlisting}
+\subsection{k-Nearest Neighbour}
 
-\end{lstlisting}
+As with the SVM-algorithm there was no normalization necessary (we tested this on different datasets).
+In WEKA there are different implementations of the kNN-algorithm available and we chose (for all our datasets) the brute force kNN-algorithm since this is the classic implementation of the kNN-algorithm where we were interested in the running time.
+We tried different parameters concerning the number of neighbours which are considered for classification.
+Changes on that parameter didn't make any differences.
+As with the SVM, the kNN-algorithm was not able to cope with the small number of data records which should be classified with value 0.
+The creation of the model doesn't take any time with a kNN-algorithm since there is no model which would be built: kNN is a lazy classification algorithm.
+The evaluation of the model took not even a second.
+In the following are the results of the classification.
+The results are exactly the same as for the SVM.
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|llllllll|}
+	\hline
+									&TP Rate   	&FP Rate   &Precision &Recall  &F-Measure   &ROC Area  &Class\\
+	\hline
+									&0         	&0         &0         &0       &  0         & 0.5      &0\\
+  								&1         	&1         &0.959     &1       &0.979       &	0.5      &1\\
+  \hline
+	Weighted Avg.		&0.959     	&0.959     &0.919   	&0.959   &0.939     	&0.5				&\\
+	\hline
+\end{tabular}
+\caption{Results of the kNN-classification for the echocardiogram dataset}
+\end{table}
+
+\vspace{6pt}
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|lll|}
+	\hline
+	Classified as	& &\\
+	a		&	b & \\
+	\hline
+	0   & 4  	&  a = 0\\
+  0   & 93 	&  b = 1\\
+  \hline
+\end{tabular}
+\caption{Confusion matrix for kNN}
+\end{table}
+
+\subsection{Comparison}
 
@@
 
-The Adult or Census Income data sets contains 48842 instances of information about persons and if they earn more than 50.000 a year. The data set has 14 attributes and missing values.
+\paragraph{}The Adult or Census Income data sets contains 48842 instances of information about persons and the task is to predict if a person earns more than 50.000 dollar a year.
+The data set has 14 attributes and missing values.
 
-The instances origin from the 1994 Census Database and were extracted by Barry Becker.
+\paragraph{}The instances origin from the 1994 Census Database and were extracted by Barry Becker.
 For having a clean data set, only records with age above 16 with minimium one working hour per week have been selected from the database.
+We performed a 10-fold-cross-validation on this dataset.
 
@@
 
-\subsection{Random Forrest}
+\subsection{Random Forest}
 
@@
 With a number of 5 randomly chosen features, the precision is yet at 85\% (with maximum depth of tree = 6) but with 10 randomly chosen features the precision falls back to 84.5\%.
+
 \subsection{Support Vector Machine}
+
+From the different kernel functions for the SVM-algorithm we used the PolyKernel on this dataset.
+As with the dataset before the normalization of the numeric data had no inpact on the validation results.
+One interesting aspect is that the creation of the model and the 10-fold-cross-validation took 135 minutes since the dataset is big.
+The SVM-algorithm was able to correctly classify 84,9\% of the instances.
+The following two tables show further validation results including a confusion matrix:
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|llllllll|}
+	\hline
+									&TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure &ROC Area  	&Class\\
+	\hline
+									&0.568     	&0.062     &0.744     	&0.568   &0.644    	&0.753    	&	>50K\\
+  								&0.938     	&0.432     &0.873     	&0.938   &0.904     &0.753    	&$\leq$50K\\
+  \hline
+	Weighted Avg.   & 0.849    & 0.343     &0.842   		&0.849   &0.842     & 0.753			&\\
+	\hline
+\end{tabular}
+\caption{Results of the SVM-classification for the adult dataset}
+\end{table}
+
+\vspace{6pt}
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|lll|}
+	\hline
+	Classified as	& &\\
+	a		&	b & \\
+	\hline
+	4454   	& 3387  	&  a = >50K\\
+  1531   	& 23189 	&  b = $\leq$50K\\
+  \hline
+\end{tabular}
+\caption{Confusion matrix for SVM}
+\end{table}
 
@@
 
-\subsection{k-Nearest}
+\subsection{k-Nearest Neighbour}
+
+We tested the k-nearest neighbour algorithm with different values for $k$.
+For $k$=1 (nearest neighbour) the algorithm was able to correctly classify 79,4\% of the instances.
+We tested the algorithm also with values 10,40 and 100 for $k$ and obtained the best results with $k=40$, where 83,48\% of the instances were correctly classified.
+Again the normalization of the input had no influence on the results.
+The evaluation of this algorithm took 120 seconds.
+In the following tables we present further results of the evaluation with $k=40$ including a confusion matrix.
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|llllllll|}
+	\hline
+									&TP Rate   &FP Rate   &Precision   &Recall  &F-Measure &ROC Area  	&Class\\
+	\hline
+									&0.593     &0.088      &0.68      &0.593     &0.633    &0.884    		&>50K\\
+									&0.912     &0.407      &0.876     &0.912     &0.893    &0.884    		&$\leq$50K\\
+  \hline
+	Weighted Avg.   &0.835     &0.331      &0.829     &0.835     &0.831    &0.884 \\
+	\hline
+\end{tabular}
+\caption{Results of the kNN-classification for the adult dataset}
+\end{table}
+
+\vspace{6pt}
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|lll|}
+	\hline
+	Classified as	& &\\
+	a		&	b & \\
+	\hline
+	4647  	&3194  		&  a = >50K\\
+  2186 		&22534 		&  b = $\leq$50K\\
+  \hline
+\end{tabular}
+\caption{Confusion matrix for kNN}
+\end{table}
+
+\subsection{Comparison}
 
@@
 The Contraceptive Method Choice Data Set consists of information about the demographic and socio-economic characteristics of married women.
-The task is to predict which kind of contraceptive methods a woman uses.
+There are 1473 instances with 9 different attributes.
+Examples of these attributes are the wife's age, husbands education, wife's religion, etc.
+The task is to predict which kind of contraceptive method a woman uses.
 These are categorized as no use, short term or long term.
+We performed a 10-fold-cross-validation on this dataset.
 
@@
 \subsection{Desicion Tree}
-\subsection{Random Forrest}
+\subsection{Random Forest}
 \subsection{Naive Bayes}
 \subsection{Support VectorMachine}
-\subsection{k-Nearest}
+
+We tested different kernel functions on this dataset and obtained the best results with the NormalizedPolyKernel.
+The whole process of the evaluation, including the building of the models, took 28 seconds.
+Normalization of the data had again no influence on the results.
+The SVM algorithm was able to correctly classify 49,2\% of the instances.
+The following tables show further results of the evaluation with the NormalizedPolyKernel including a confusion matrix.
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|llllllll|}
+	\hline
+	&								TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure   &ROC Area  	&Class\\
+	\hline
+									&0.583     	&0.355     &0.55      	&0.583   &0.566   		&0.628    	&1\\
+                 	&0.393     	&0.181     &0.389     	&0.393   &0.391      	&0.648    	&2\\
+                 	&0.444     	&0.252     &0.484     	&0.444   &0.463      	&0.614    	&3\\
+  \hline
+	Weighted Avg.   & 0.849    & 0.343      &0.842   &  0.849     &0.842     & 0.753			&\\
+	\hline
+\end{tabular}
+\caption{Results of the SVM-classification for the contraception dataset}
+\end{table}
+
+\vspace{6pt}
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|llll|}
+	\hline
+	Classified as	& & &\\
+	a		&	b & c& \\
+	\hline
+	367 &105 &157 &   a = 1\\
+ 	117 &131 &85 	&   b = 2\\
+ 	183 &101 &227 &   c = 3\\
+  \hline
+\end{tabular}
+\caption{Confusion matrix for SVM}
+\end{table}
+
+\subsection{k-Nearest Neighbour}
+
+We tested the kNN-algorithm with a lot of different values for $k$ and obtained the best results for $k=7$ where 48,88\% of the instances were correctly classified.
+The evaluation took not even a second.
+As with the other datasets the normalization of the input had no influence on the results.
+In the following tables we present further results of the evaluation with $k=7$ including a confusion matrix.
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|llllllll|}
+	\hline
+	&								TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure   &ROC Area  	&Class\\
+	\hline
+									&0.569     	&0.351     &0.547     	&0.569   &0.558      	&0.65     	&1\\
+                 	&0.414     	&0.166     &0.422     	&0.414   &0.418      	&0.677    	&2\\
+                 	&0.438     	&0.279     &0.455     	&0.438   &0.447      	&0.603    	&3\\
+  \hline
+	Weighted Avg.   &0.489     	&0.284     &0.487     	&0.489   &0.488      	&0.64			&\\
+	\hline
+\end{tabular}
+\caption{Results of the SVM-classification for the contraception dataset}
+\end{table}
+
+\vspace{6pt}
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|llll|}
+	\hline
+	Classified as	& & &\\
+	a		&	b & c& \\
+	\hline
+	358	& 99 	&172 &   a = 1\\
+  99 	&	138 &96  &   b = 2\\
+ 	197 & 90 	&224 &   c = 3\\
+  \hline
+\end{tabular}
+\caption{Confusion matrix for SVM}
+\end{table}
+
+\subsection{Comparison}
 

### 2012-11-24 paper.tex
Signed-off-by: schubiduu <b.kiesl@atn.nu>
asdf

@@
 
-\section{Echocardiogram data}
+\section{Preface}
 
-The Echocardiogram data set originally consists of 132 entries with 13 attributes respectively.
+\paragraph{}For the first exercise 3 data sets were chosen on which five classification algorithms were tested to practice basic machine learning tasks and see how these algorithms behave with their various settings.
+For the elaboration of this paper the Machine Learning Software Weka was used.
+It offered all necessary algorithms.
+
+\paragraph{}As data sets we chose as follows:
+\begin{itemize}
+\item Echo Cardiogram, a data set about heart attacks and if a patient survived one year after the attack.
+\item Adult / Census Income, a data set about people and their basic information and if they earn more than 50k a year.
+\item Contraceptive Method Choice, a data set about women and their environment and basic information and what contraceptive method they use.
+\end{itemize}
+
+\paragraph{}Five classification algorithms were chosen that we used on all data sets. Each chapter will also dicuss how the classifiers perform an the data and why.
+
+\section{Echo Cardiogram data}
+
+\paragraph{}The Echocardiogram data set originally consists of 132 entries with 13 attributes respectively.
 The entries of the dataset refer to patients that suffered from a heart attack at some point in the past \cite{uci-repo}.
@@
 \subsection{Decision Tree}
+\todo{robert: do you want to keep this section?
+on the one hand it shows what happens if we run a desicion tree with attributes on which the the prediction attribute depends, but on the other hand we know we should not do that.
+please decide.
+i'm ok with both (philipp)}
 %------------------------------------------------
@@
 
-\subsection{Decision Tree on re-preprocessed data}
+\subsection{Decision Tree on preprocessed data}
 
-The J48 Decision Tree shows a more useful tree on the re-preprocessed data:
+The J48 Decision Tree shows a more useful tree on the preprocessed data:
+
 \begin{lstlisting}
@@
 \end{tabular}
-\caption{Results of the SVM-classification for the contraception dataset}
+\caption{Results of the kNN-classification for the contraception dataset}
 \end{table}
@@
 \end{tabular}
-\caption{Confusion matrix for SVM}
+\caption{Confusion matrix for kNN}
 \end{table}

### 2012-11-24 paper.tex
calculated default features -- introduced booktabs package

--- 2012-11-24 paper.tex
@@
 \usepackage{listings}
+\usepackage{booktabs}
 \bibliographystyle{plain}
@@
 ------------------
-
 survival <= 12
@@
 J48 pruned tree
-
 epss <= 13100: 1 (90.63/1.95)
@@
 J48 unpruned tree
-
 pericardial-effusion = 0
@@
 With default settings, i.e.
-default number of features \todo[inline]{calculate!} and 10 random trees with unlimited tree depth, the random forest shows us a rather disappointing result:
+default number of features $ log_2(7) + 1 \approx 4 $, and 10 random trees with unlimited tree depth, the random forest shows us a rather disappointing result:
 
@@
 And the statistical measures:
-\begin{lstlisting}
-=== Detailed Accuracy By Class ===
-                 TP Rate  FP Rate  Precision  Recall  F-Measure    Class
-                 0        0.022    0          0       0            0
-                 0.978    1        0.958      0.978   0.968        1
-Weighted Avg.    0.938    0.96     0.918      0.938   0.928     
-\end{lstlisting}
-\paragraph{Parameter: Number of trees}
+\begin{table}
+    \begin{tabular}{lllllll}
+        \toprule 
+        \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+              &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class \\ \midrule
+              &  0       & 0.022   & 0         & 0      & 0         & 0     \\
+              &  0.978   & 1       & 0.958     & 0.978  & 0.968     & 1     \\
+Weighted Avg. &  0.938   & 0.96    & 0.918     & 0.938  & 0.928     &       \\ \bottomrule
+    \end{tabular}
+\end{table}
+
+\paragraph{Number of Random Trees}
 \todo[inline]{proper line break here}
@@
 Reducing the number of the trees increases the number of false positives but doesn't catch any true positives.
-\paragraph{Limiting the tree-depth}
+\paragraph{Limiting the maximum depth of the trees}
 Limitations in the depths of the random trees has also almost no effect on the accuracy of random forests on this data set.
@@
 \end{lstlisting}
-which is surprisingly a worse result than from the J48 algorithm.
+which is surprisingly worse than the J48 algorithm.
 
@@
 
-\paragraph{Limitations of the maximum depth of the trees}
+\paragraph{Limiting the maximum depth of the trees}
 We have a sign for overfitting under default settings: Reducing the maximum depth has a positive effect on the quality.
@@
 Increasing the number of features doesn't support the quality at all.
-The default number of features for this data set with 13 attributes is \todo[inline]{calculate!}.
-With a number of 5 randomly chosen features, the precision is yet at 85\% (with maximum depth of tree = 6) but with 10 randomly chosen features the precision falls back to 84.5\%.
-
+The default number of features for this data set with 13 attributes is $ log_2(13) + 1 \approx 5 $.
+Choosing 6 attributes brinchoose gs us a precision of 84,9\% and with 10 randomly chosen features the precision falls back to 84.5\%.
+We can tease the most out of the random forest if we decrease the number of features a little.
+To be concise, we get the best achievable result, if we take only 3 randomly chosen
+features.
+\begin{lstlisting}
+=== Detailed Accuracy By Class ===
+                 TP Rate  FP Rate  Precision  Recall  F-Measure Class
+                 0.577    0.052    0.779      0.577   0.663     >50K
+                 0.948    0.423    0.876      0.948   0.911     <=50K
+Weighted Avg.    0.859    0.333    0.853      0.859   0.851    
+\end{lstlisting}
+Which is a quite similar accuracy like the best of the J48 algorithm. 
 \subsection{Support Vector Machine}

### 2012-11-24 paper.tex
fine tuned booktabs table example
restyled tables

--- 2012-11-24 paper.tex
@@
 
+% ------------------------------------
+% E C H O   C A R D I O    S E T
+% ------------------------------------
+
 \section{Echo Cardiogram data}
@@
 J48 pruned tree
-------------------
-
 survival <= 12
@@
 
-That's correct, if the patient survived $ <= 12 $ months, he survived at least 12 month. What a statement!
+\paragraph{}That's correct, if the patient survived $ <= 12 $ months, he survived at least 12 month. What a statement!
 
-The only remarkable thing is, that the J48 implementation under default settings seems to enforce a certain depth of the tree, even if the splitting on the same attribute don't make any sense ($ 11 $ is yet a smaller number than $ 12 $, isn't it?).
+\paragraph{}The only remarkable thing is, that the J48 implementation under default settings seems to enforce a certain depth of the tree, even if the splitting on the same attribute don't make any sense ($ 11 $ is yet a smaller number than $ 12 $, isn't it?).
 E.g.
@@
 \end{lstlisting}
-Since there is no missing value for the survival attribute, this branch can be ignored.
+\paragraph{}Since there is no missing value for the survival attribute, this branch can be ignored.
 
-But allowing the machine learning algorithms to see the origin attributes from which we derived the prediction attribut leads to absurdity, since the actual prediction task is to say whether a patient will survive one year after suffering a heart attack or not \emph{depending on the medical characteristics}.
+\paragraph{}But allowing the machine learning algorithms to see the origin attributes from which we derived the prediction attribut leads to absurdity, since the actual prediction task is to say whether a patient will survive one year after suffering a heart attack or not \emph{depending on the medical characteristics}.
 
@@
 
-The J48 Decision Tree shows a more useful tree on the preprocessed data:
+\paragraph{}The J48 Decision Tree shows a more useful tree on the preprocessed data:
 
@@
 
-The documentation of the data set explains the two attributes as follows:
+\paragraph{}The documentation of the data set explains the two attributes as follows:
 \begin{quote}
@@
 
-Only the rule for the fractional shortening should be the other way round, because a low fractional shortening is a rather pathological sign.
+\paragraph{}Only the rule for the fractional shortening should be the other way round, because a low fractional shortening is a rather pathological sign.
 Although the decision tree has a high precision of almost 95\% it performs very imprecise, because none of the patients who died after one year are correctly classified:
 
-\begin{lstlisting}
-=== Confusion Matrix ===
-  a  b   <-- classified as
-  0  4 |  a = 0
-  1 90 |  b = 1
-
-\end{lstlisting}
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|ll|}
+\hline
+ & a &	b \\
+\hline
+a = 0  & 0 & 4 	\\
+b = 1 & 1 & 90 	\\
+\hline
+\end{tabular}
+\caption{Confusion matrix for Decision tree}
+\end{table}
 
@@
 
-\begin{lstlisting}
-=== Confusion Matrix ===
-  a  b   <-- classified as
-  0  4 |  a = 0
-  2 91 |  b = 1
-\end{lstlisting}
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|ll|}
+\hline
+ & a &	b \\
+\hline
+a=0 & 0 & 4 \\
+b=1 & 2 & 91\\
+\hline
+\end{tabular}
+\caption{Confusion matrix for Random Forest}
+\end{table}
+
 And the statistical measures:
@@
 The only change to the outcome can be observed as soon the depth of the trees is limited to 6:
-\begin{lstlisting}
-    === Confusion Matrix ===
-    a  b   <-- classified as
-    0  4 |  a = 0
-    1 92 |  b = 1
-\end{lstlisting}
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|ll|}
+\hline
+ & a &	b \\
+\hline
+a=0 & 0 & 4 \\
+b=1 & 1 & 92\\
+\hline
+\end{tabular}
+\caption{Confusion matrix for Random Forest, limited depth}
+\end{table}
+
 Only one false positive has gone. Not really a remarkable change in the quality. This does not change if the tree-depth is further reduced.
@@
 But if we configure the random forests with 6 features we can spot our first true positive:
-\begin{lstlisting}
-=== Confusion Matrix ===
-  a  b   <-- classified as
-  1  3 |  a = 0
-  2 91 |  b = 1
-\end{lstlisting}
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|ll|}
+\hline
+ & a &	b \\
+\hline
+a=0 & 1 & 3 \\
+b=1 & 2 & 91\\
+\hline
+\end{tabular}
+\caption{Confusion matrix for Random Forest, increased number of features}
+\end{table}
+
 Which gives us the best achievable accuracy that we can reach with decision based algorithms:
@@
 \centering
-\begin{tabular}{|lll|}
+\begin{tabular}{|l|ll|}
 \hline
-	Classified as	& &\\
-	a		&	b & \\
+ & a &	b \\
 \hline
-	0   & 4  	&  a = 0\\
-  0   & 93 	&  b = 1\\
+a=0 & 0 & 4 \\
+b=1 & 0 & 93\\
 \hline
@@
 \subsection{Comparison}
+
+% ------------------------------------
+% A D U L T   D A T A    S E T
+% ------------------------------------
 
@@
 \centering
-\begin{tabular}{|lll|}
+\begin{tabular}{|l|ll|}
 \hline
-	Classified as	& &\\
-	a		&	b & \\
+              &    a &     b \\
 \hline
-	4454   	& 3387  	&  a = >50K\\
-  1531   	& 23189 	&  b = $\leq$50K\\
+ a = >50k      & 4454 &  3387 \\
+ b = $\leq$50k & 1531 & 23189\\
 \hline
@@
 \centering
-\begin{tabular}{|lll|}
+\begin{tabular}{|l|ll|}
 	\hline
-	Classified as	& &\\
-	a		&	b & \\
+	 &              a    & b     \\
 	\hline
-	4647  	&3194  		&  a = >50K\\
-  2186 		&22534 		&  b = $\leq$50K\\
+    a = >50K      &	4647 & 3194  \\
+    b = $\leq$50K & 2186 & 22534 \\
     \hline
@@
 \subsection{Comparison}
+
+% ------------------------------------
+% C O N T R A S E P T I V E    S E T
+% ------------------------------------
 
@@
 \centering
-\begin{tabular}{|llll|}
+\begin{tabular}{|l|lll|}
 	\hline
-	Classified as	& & &\\
-	a		&	b & c& \\
+	      &a   &b   &c \\
 	\hline
-	367 &105 &157 &   a = 1\\
- 	117 &131 &85 	&   b = 2\\
- 	183 &101 &227 &   c = 3\\
+	a = 1 &367 &105 &157 \\
+ 	b = 2 &117 &131 &85 \\
+ 	c = 3 &183 &101 &227 \\
   \hline
@@
 \centering
-\begin{tabular}{|llll|}
+\begin{tabular}{|l|lll|}
 	\hline
-	Classified as	& & &\\
-	a		&	b & c& \\
+	      &a   &b   &c \\
 	\hline
-	358	& 99 	&172 &   a = 1\\
-  99 	&	138 &96  &   b = 2\\
- 	197 & 90 	&224 &   c = 3\\
+    a = 1 &358 &99  &172\\
+    b = 2 &99  &138 &96 \\
+ 	c = 3 &197 &90  &224 \\
   \hline

### 2012-11-24 paper.tex

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%x%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Short Sectioned Assignment
@@
 J48 pruned tree
+------------------
 survival <= 12
@@
 J48 pruned tree
+------------------
 epss <= 13100: 1 (90.63/1.95)
@@
 And the statistical measures:
-\begin{lstlisting}
-=== Detailed Accuracy By Class ===
-                 TP Rate  FP Rate  Precision  Recall  F-Measure    Class
-                 0        0.022    0          0       0            0
-                 0.978    1        0.958      0.978   0.968        1
-Weighted Avg.    0.938    0.96     0.918      0.938   0.928     
-\end{lstlisting}
-\paragraph{Parameter: Number of trees}
+\begin{table*}[h]\centering
+    \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+        \toprule 
+        \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+              &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class \\ \midrule
+              &  0       & 0.022   & 0         & 0      & 0         & 0     \\
+              &  0.978   & 1       & 0.958     & 0.978  & 0.968     & 1     \\
+Weighted Avg. &  0.938   & 0.96    & 0.918     & 0.938  & 0.928     &       \\ \bottomrule
+    \end{tabular*}
+\caption{Random Forest -- Default Settings} 
+\end{table*}
+
+\paragraph{Number of Random Trees}
 \todo[inline]{proper line break here}

### 2012-11-24 paper.tex
inermediate step, building proper tables

@@
 
+\usepackage{natbib}
 % philipp: added for code listings
-\usepackage{listings}
-
-\usepackage{natbib}
 \usepackage{listings}
@@
 \paragraph{Number of Random Trees}
-\todo[inline]{proper line break here}
-Reducing the number of the trees continuously decreases the quality of the model, as shown in the accuracy details for 5 random trees:
-\begin{lstlisting}
-=== Detailed Accuracy By Class ===
-                 TP Rate  FP Rate  Precision  Recall  F-Measure    Class
-                 0        0.032    0          0       0            0  
-                 0.968    1        0.957      0.968   0.963        1   
-Weighted Avg.    0.928    0.96     0.918      0.928   0.923    
-\end{lstlisting}
-.. and for a single random tree:
-\begin{lstlisting}
-=== Detailed Accuracy By Class ===
-                 TP Rate  FP Rate  Precision  Recall  F-Measure    Class
-                 0        0.054    0          0       0            0  
-                 0.946    1        0.957      0.946   0.951        1   
-Weighted Avg.    0.907    0.961    0.917      0.907   0.912    
-\end{lstlisting}
-Reducing the number of the trees increases the number of false positives but doesn't catch any true positives.
+Reducing the number of the trees continuously decreases the quality of the model, tables \ref{tab:echo:rand:5t} and \ref{tab:echo:rand:1t} show the regression of quality.
+This method increases the number of false positives but doesn't catch any true positives.
+\begin{table*}[H]\centering
+    \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+        \toprule 
+        \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+               &  TP Rate & FP Rate & Precision & Recall & F-Measure &   Class  \\ \midrule
+               &  0       & 0.032   & 0         & 0      & 0         &   0      \\
+               &  0.968   & 1       & 0.957     & 0.968  & 0.963     &   1      \\     
+Weighted Avg.  &  0.928   & 0.96    & 0.918     & 0.928  & 0.923     &          \\ \bottomrule
+    \end{tabular*}
+\caption{Random Forest -- 5 Random Trees} 
+\label{tab:echo:rand:5t}
+\end{table*}
+
+\begin{table*}[H]\centering
+    \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+        \toprule 
+        \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+               &  TP Rate & FP Rate & Precision & Recall & F-Measure &   Class \\ \midrule
+               &  0       & 0.054   & 0         & 0      & 0         &   0     \\ 
+               &  0.946   & 1       & 0.957     & 0.946  & 0.951     &   1     \\ 
+Weighted Avg.  &  0.907   & 0.961   & 0.917     & 0.907  & 0.912     &         \\ \bottomrule
+    \end{tabular*}
+\caption{Random Forest -- 1 Random Tree} 
+\label{tab:echo:rand:1t}
+\end{table*}
 \paragraph{Limiting the maximum depth of the trees}
@@
 
-Which gives us the best achievable accuracy that we can reach with decision based algorithms:
-\begin{lstlisting}
-=== Detailed Accuracy By Class ===
-                TP Rate  FP Rate  Precision  Recall  F-Measure   Class
-                0.25     0.022    0.333      0.25    0.286       0
-                0.978    0.75     0.968      0.978   0.973       1
-Weighted Avg.   0.948    0.72     0.942      0.948   0.945      
-\end{lstlisting}
+Which gives us the best achievable accuracy that we can reach with decision based algorithms, shown in table \ref{tab:echo:rand:6f}.
+\begin{table*}[H]\centering
+    \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+        \toprule 
+        \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+               &  TP Rate &  FP Rate &  Precision &  Recall &  F-Measure &  Class \\ \midrule
+               &  0.25    &  0.022   &  0.333     &  0.25   &  0.286     &  0     \\
+               &  0.978   &  0.75    &  0.968     &  0.978  &  0.973     &  1     \\
+Weighted Avg.  &  0.948   &  0.72    &  0.942     &  0.948  &  0.945     &        \\ \bottomrule
+    \end{tabular*}
+\caption{Random Forest -- 6 randomly chosen features} 
+\label{tab:echo:rand:6f}
+\end{table*}
 
@@
 First of all, with the default settings of the J48 implementation of weka, i.e.
-using pruning with a confidence factor of 0.25 together with MDLCorrection and Subtree raising and at least 2 instances per leaf of the tree, we get the following summary of the quality:
+using pruning with a confidence factor of 0.25 together with MDLCorrection and Subtree raising and at least 2 instances per leaf of the tree, we get a summary of the quality as shown in table \ref{tab:adult:dec:default}:
 
-\begin{lstlisting}
-=== Detailed Accuracy By Class ===
-                TP Rate  FP Rate  Precision  Recall  F-Measure  Class
-                0.615    0.062    0.759      0.615   0.68       >50K
-                0.938    0.385    0.885      0.938   0.911      <=50K
-Weighted Avg.   0.86     0.307    0.854      0.86    0.855      
-\end{lstlisting}
+\begin{table*}[H]\centering
+    \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+        \toprule 
+        \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+               & TP Rate & FP Rate & Precision & Recall &  F-Measure &  Class \\ \midrule  
+               & 0.615   & 0.062   & 0.759     & 0.615  &  0.68      &  >50K  \\  
+               & 0.938   & 0.385   & 0.885     & 0.938  &  0.911     &  <=50K \\  
+Weighted Avg.  & 0.86    & 0.307   & 0.854     & 0.86   &  0.855     &        \\  \bottomrule 
+    \end{tabular*}
+\caption{Decision Tree -- Default Settings} 
+\label{tab:adult:dec:default}
+\end{table*}
 

### 2012-11-25 paper.tex
converted all accuracy details to tables, still misplaced at the end of the paper

@@
 \usepackage{booktabs}
+\usepackage{morefloats}
 \bibliographystyle{plain}
@@
 \paragraph{Raising the confidence factor}
-Subsequently increasing the confidence factor, which result in less pruning, leads to a continuous worse quality. If we set the confidence factor to 0.5 we get the following outcome:
-\begin{lstlisting}
-=== Detailed Accuracy By Class ===
-                TP Rate  FP Rate  Precision  Recall  F-Measure  Class
-                0.632    0.071    0.738      0.632   0.681      >50K
-                0.929    0.368    0.888      0.929   0.908      <=50K
-Weighted Avg.   0.857    0.296    0.852      0.857   0.854      
-\end{lstlisting}
-.. for a confidence factor of 0.75 we get the following:
-\begin{lstlisting}
-=== Detailed Accuracy By Class ===
-                TP Rate  FP Rate  Precision  Recall  F-Measure  Class
-                0.655    0.094    0.689      0.655   0.672      >50K
-                0.906    0.345    0.892      0.906   0.899      <=50K
-Weighted Avg.   0.846    0.285    0.843      0.846   0.844     
-\end{lstlisting}
+Subsequently increasing the confidence factor, which result in less pruning, leads to a continuous worse quality.
+If we set the confidence factor to 0.5 we get an outcome as shown in table \ref{tab:adult:dec:cf5} and if we set it to 0.75 the quality gets even worse as seen in \ref{tab:adult:dec:cf75}.
+\begin{table*}[b]\centering
+    \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+        \toprule 
+        \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+              &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class    \\  \midrule
+              &  0.632   & 0.071   & 0.738     & 0.632  & 0.681     & >50K     \\ 
+              &  0.929   & 0.368   & 0.888     & 0.929  & 0.908     & <=50K    \\ 
+Weighted Avg. &  0.857   & 0.296   & 0.852     & 0.857  & 0.854     &          \\  \bottomrule
+    \end{tabular*}
+\caption{Decision Tree -- Confidence Factor of 0.5} 
+\label{tab:adult:dec:cf5}
+\end{table*}
+
+\begin{table*}[b]\centering
+    \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+        \toprule 
+        \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+              &  TP Rate & FP Rate & Precision & Recall &  F-Measure &  Class  \\ \midrule        
+              &  0.655   & 0.094   & 0.689     & 0.655  &  0.672     &  >50K   \\          
+              &  0.906   & 0.345   & 0.892     & 0.906  &  0.899     &  <=50K  \\          
+Weighted Avg. &  0.846   & 0.285   & 0.843     & 0.846  &  0.844     &         \\ \bottomrule          
+    \end{tabular*}
+\caption{Decision Tree -- Confidence Factor of 0.75} 
+\label{tab:adult:dec:cf75}
+\end{table*}
 
@@
 
-With default settings, a random forest achieves an accuracy as follows:
-\begin{lstlisting}
-=== Detailed Accuracy By Class ===
+With default settings, a random forest gains an accuracy on the census income data set that can be seen in table \ref{tab:adult:rand:default}:
+\begin{table*}[b]\centering
+  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+      \toprule 
+      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
                 TP Rate  FP Rate  Precision  Recall  F-Measure  Class
@@
 Weighted Avg.   0.825    0.318    0.823      0.825   0.824     
-\end{lstlisting}
+    \end{tabular*}
+\caption{Random Forest -- Default Settings} 
+\label{tab:adult:rand:default}
+\end{table*}
 which is surprisingly worse than the J48 algorithm.
@@
 
-\begin{lstlisting}
-=== Detailed Accuracy By Class ===
+\begin{table*}[b]\centering
+  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+      \toprule 
+      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
                 TP Rate  FP Rate  Precision  Recall  F-Measure  Class
@@
 Weighted Avg.   0.856    0.314    0.85       0.856   0.851     
-\end{lstlisting}
+    \end{tabular*}
+\caption{Random Forest -- 7 Random Trees} 
+\label{tab:adult:rand:7t}
+\end{table*}
 We can push the accuracy even further, if we take more random trees, e.g. twenty:
-\begin{lstlisting}
-=== Detailed Accuracy By Class ===
-                TP Rate  FP Rate  Precision  Recall  F-Measure  Class
-                0.593    0.058    0.763      0.593   0.667      >50K
-                0.942    0.407    0.879      0.942   0.909      <=50K
-Weighted Avg.   0.858    0.323    0.851      0.858   0.851      
-\end{lstlisting}
+\begin{table*}[b]\centering
+  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+      \toprule 
+      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+                & TP Rate  & FP Rate  & Precision &  Recall &  F-Measure  & Class  \\  \midrule          
+                & 0.593    & 0.058    & 0.763     &  0.593  &  0.667      & >50K   \\     
+                & 0.942    & 0.407    & 0.879     &  0.942  &  0.909      & <=50K  \\     
+Weighted Avg.   & 0.858    & 0.323    & 0.851     &  0.858  &  0.851      &        \\  \bottomrule
+    \end{tabular*}
+\caption{Random Forest -- 20 Random Trees} 
+\label{tab:adult:rand:20t}
+\end{table*}
 
@@
 features.
-\begin{lstlisting}
-=== Detailed Accuracy By Class ===
-                 TP Rate  FP Rate  Precision  Recall  F-Measure Class
-                 0.577    0.052    0.779      0.577   0.663     >50K
-                 0.948    0.423    0.876      0.948   0.911     <=50K
-Weighted Avg.    0.859    0.333    0.853      0.859   0.851    
-\end{lstlisting}
+\begin{table*}[b]\centering
+  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+      \toprule 
+      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+               &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class \\   \midrule
+               &  0.577   & 0.052   & 0.779     & 0.577  & 0.663     & >50K  \\   
+               &  0.948   & 0.423   & 0.876     & 0.948  & 0.911     & <=50K \\   
+Weighted Avg.  &  0.859   & 0.333   & 0.853     & 0.859  & 0.851     &       \\   \bottomrule
+    \end{tabular*}
+\caption{Random Forest -- 3 Random Features} 
+\label{tab:adult:rand:3f}
+\end{table*}
 Which is a quite similar accuracy like the best of the J48 algorithm. 

### 2012-11-25 paper.tex
forgot to delimit columns for two tables

@@
       \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
-                TP Rate  FP Rate  Precision  Recall  F-Measure  Class
-                0.615    0.108    0.644      0.615   0.629      >50K
-                0.892    0.385    0.88       0.892   0.886      <=50K
-Weighted Avg.   0.825    0.318    0.823      0.825   0.824     
+              &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class    \\  \midrule
+              &  0.615   & 0.108   & 0.644     & 0.615  & 0.629     & >50K     \\     
+              &  0.892   & 0.385   & 0.88      & 0.892  & 0.886     & <=50K    \\ 
+Weighted Avg. &  0.825   & 0.318   & 0.823     & 0.825  & 0.824     &          \\ \bottomrule
     \end{tabular*}
@@
       \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
-                TP Rate  FP Rate  Precision  Recall  F-Measure  Class
-                0.606    0.064    0.749      0.606   0.67       >50K
-                0.936    0.394    0.882      0.936   0.908      <=50K
-Weighted Avg.   0.856    0.314    0.85       0.856   0.851     
+              &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class    \\  \midrule
+              &  0.606   & 0.064   & 0.749     & 0.606  & 0.67      & >50K     \\ 
+              &  0.936   & 0.394   & 0.882     & 0.936  & 0.908     & <=50K    \\ 
+Weighted Avg. &  0.856   & 0.314   & 0.85      & 0.856  & 0.851     &          \\  \bottomrule
     \end{tabular*}

### 2012-11-25 paper.tex
forced tables to be placed at current position with \Floatbarrier     maybe a poor typographic decision ? improvements welcome

@@
 
+\usepackage{float}
+\usepackage{placeins}
 \usepackage{natbib}
@@
 And the statistical measures:
-\begin{table*}[h]\centering
+\begin{table*}[htb]\centering
     \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
@@
 \end{table*}
+\FloatBarrier
 
@@
 This method increases the number of false positives but doesn't catch any true positives.
-\begin{table*}[H]\centering
+\begin{table*}[htb]\centering
     \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
@@
 \end{table*}
+\FloatBarrier
 
-\begin{table*}[H]\centering
+\begin{table*}[htb]\centering
     \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
@@
 \end{table*}
+\FloatBarrier
 \paragraph{Limiting the maximum depth of the trees}
@@
 Which gives us the best achievable accuracy that we can reach with decision based algorithms, shown in table \ref{tab:echo:rand:6f}.
-\begin{table*}[H]\centering
+\begin{table*}[htb]\centering
     \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
@@
 
-\begin{table*}[H]\centering
+\begin{table*}[htb]\centering
     \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
@@
 If we set the confidence factor to 0.5 we get an outcome as shown in table \ref{tab:adult:dec:cf5} and if we set it to 0.75 the quality gets even worse as seen in \ref{tab:adult:dec:cf75}.
-\begin{table*}[b]\centering
+\begin{table*}[htb]\centering
     \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
@@
 \end{table*}
+\FloatBarrier
 
-\begin{table*}[b]\centering
+\begin{table*}[htb]\centering
     \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
@@
 With default settings, a random forest gains an accuracy on the census income data set that can be seen in table \ref{tab:adult:rand:default}:
-\begin{table*}[b]\centering
+\begin{table*}[htb]\centering
   \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
@@
 \end{table*}
+\FloatBarrier
 which is surprisingly worse than the J48 algorithm.
@@
 
-\begin{table*}[b]\centering
+\begin{table*}[htb]\centering
   \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
@@
 \end{table*}
+\FloatBarrier
 We can push the accuracy even further, if we take more random trees, e.g. twenty:
-\begin{table*}[b]\centering
+\begin{table*}[htb]\centering
   \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
@@
 features.
-\begin{table*}[b]\centering
+\begin{table*}[htb]\centering
   \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
@@
 \end{table*}
+\FloatBarrier
 Which is a quite similar accuracy like the best of the J48 algorithm. 

### 2012-11-25 paper.tex
inserted todo markers

--- 2012-11-25 paper.tex
@@
 
+\subsection{Naive Bayes}
+
+\todo{todo marker}
+
 \subsection{Comparison}
+
+\todo{todo marker}
 
@@
 
+\todo{rewrite this section}
+
 We use 66\% of the data set for training and the rest for evaluation.
@@
 
-Second run: we use the discretizer with 20 bin.
+Second run: we use the discretizer with 20 bins.
 
@@
 
+\todo{todo marker}
+
 % ------------------------------------
@@
 
-The attributes are already in a numerical and categorical form, whereby the categories are encoded in an ordinal form. These go from 1 to 4 where 1 is low/bad and 4 is high/good.
-
-\todo{check if ordinal is the correct term when i mean, that the attributes are encoded in numbers, where 1<2, 1<4, 4>2, ...}
+The values of the data sets's attributes are already in a numerical and categorical form, whereby the categories are encoded in an ordinal form.
+These go from 1 to 4 where 1 means low/bad and 4 means high/good.
+This set has no missing values and the attributes do not depend on each other.
+So we can use the data as is.
 
 \subsection{Desicion Tree}
+
+\todo{todo marker}
+
 \subsection{Random Forest}
+
+\todo{todo marker}
+
 \subsection{Naive Bayes}
+
+\todo{todo marker}
+
 \subsection{Support VectorMachine}
@@
 
+\todo{todo marker}
+
 \bibliography{references}

### 2012-11-25 paper.tex
2x bayes

@@
 
-\todo{todo marker}
+For the fist try, the unsupervised discretizer was used. But the result was bad. The classifier could predict one single instance for for patients that did not survive one year.
+
+For the second try we prepare the data set with the PKIDiscretizer.
+The description says it "Discretizes numeric attributes using equal frequency binning, where the number of bins is equal to the square root of the number of non-missing values." But as we can see in \ref{tab:echo:bayes:1c} this discretizer does not perform better.
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|ll|}
+\hline
+ & a &	b \\
+\hline
+a=0 & 0 & 4 \\
+b=1 & 1 & 92\\
+\hline
+\end{tabular}
+\caption{Confusion matrix for kNN}
+\label{tab:echo:bayes:1c}
+\end{table}
+
+The log output of Naive Bayes shows that each bin has 1-2 negative (alive-at-1 = 0) instances. This means that each bin has the same low possibility for TP-prediction.
+
+So we have to boost the instances we want to predict.
+For that the SMOTE filter is used.
+It produces more instances of a given class index by looking at near neighbours with the same class value.
+The filter was applied several times, until the number of instances that have alive-at-1 = 0 1.2 times more than the alive-at-1 = 1 instances (128:93).
+The SMOTE filter used 3 nearest neighbours and produced 100\% new instances in each run.
+In this way we raised the numbers of instances with an outcome we want to predict, as for this training set it is better to predict if a patient is at risk to not survive one year.
+
+We have to supply a separate set for validation as we kind of destroy the data set when generating new instances.
+For that the original training set was taken.
+Although it is not good to validate a classifier with the same set as the training set, we can use this approach because our training set does not have a lot in common with the original set as we added a lot of generated instances.
+
+The confusion matrix \ref{tab:echo:bayes:2c} shows, that we predicted three out of four correct instances and also had more errors when predicting alive-at-1 = 0 than  alive-at-1 = 1.
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|ll|}
+\hline
+ & a &	b \\
+\hline
+a=0 & 3 & 1 \\
+b=1 & 9 & 84\\
+\hline
+\end{tabular}
+\caption{Confusion matrix for kNN}
+\label{tab:echo:bayes:2c}
+\end{table}
+
+Also the detailed accuracy table \ref{tab:echo:bayes:2d} shows that this approach gained a high recall with a low precision for alive-at-1 = 0 and both high precision and recall for alive-at-1 = 1.
+
+\begin{table*}[htb]\centering
+  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+      \toprule 
+      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+              &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class    \\  \midrule
+              &   0.75   & 0.097   & 0.25      & 0.75   & 0.375     & 0 \\
+              &   0.903  & 0.25    & 0.988     & 0.903  & 0.944     & 1 \\
+Weighted Avg. &   0.897  & 0.244   & 0.958     & 0.897  & 0.92      &   \\ \bottomrule
+  \end{tabular*}
+\caption{Naive Bayes with SMOTE filter} 
+\label{tab:echo:bayes:2d}
+\end{table*}
 
@@
 
-\todo{rewrite this section}
+As further preprocessing for Naive Bayes it is important to dicretize the numerical attributes.
+The PKIDescretizer was used again as it produces better bin sizes than the standard discretizer without any configuration.
 
-We use 66\% of the data set for training and the rest for evaluation.
+After the first run with the Simple Naive Bayes algorithm we gain the result shown in table \label{tab:adult:bayes:1d} and a confusion matrix \ref{tab:adult:bayes:1c}.
 
-For the first run, we use the default setting and get the following result:
+\begin{table*}[htb]\centering
+  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+      \toprule 
+      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+               &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class \\   \midrule
+               &  0.806   & 0.163   & 0.611     & 0.806  & 0.695     & >50K  \\
+               &  0.837   & 0.194   & 0.932     & 0.837  & 0.882     & <=50K \\
+Weighted Avg.  &  0.83    & 0.186   & 0.854     & 0.83   & 0.837     &       \\   \bottomrule
+    \end{tabular*}
+\caption{Naive Bayes -- preprocessed with PKIDiscretizer} 
+\label{tab:adult:bayes:1d}
+\end{table*}
 
-\begin{verbatim}
-Correctly Classified Instances        9104               82.2329 %
-Incorrectly Classified Instances      1967               17.7671 %
-Total Number of Instances            11071
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|ll|}
+\hline
+              &    a &     b \\
+\hline
+ a = >50k      & 6321 &1520   \\
+ b = $\leq$50k & 4025 &20695 \\
+\hline
+\end{tabular}
+\caption{Confusion matrix for Naive Bayes}
+\label{tab:adult:bayes:1c}
+\end{table}
 
-Precision   Recall  F-Measure  Class
- 0.592       0.786   0.675      >50K
- 0.927       0.834   0.878      <=50K
-\end{verbatim}
-
-Second run: we use the discretizer with 20 bins.
-
-\begin{verbatim}
-Correctly Classified Instances        9287               83.8858 %
-Incorrectly Classified Instances      1784               16.1142 %
-Total Number of Instances            11071
-
-=== Detailed Accuracy By Class ===
-
-Precision   Recall  F-Measure  Class
- 0.621       0.805   0.701      >50K
- 0.934       0.849   0.89       <=50K
-\end{verbatim}
-
-As expected, the result is slightly better.
-
-A better result should be possible if we let the dicretizer split the values in part of nearly equal size. The result:
-
-\begin{verbatim}
-Correctly Classified Instances        9179               82.9103 %
-Incorrectly Classified Instances      1892               17.0897 %
-Total Number of Instances            11071
-
-Precision   Recall  F-Measure  Class
-0.602        0.805   0.689      >50K
-0.933        0.836   0.882      <=50K
-\end{verbatim}
-
-It got worse. \todo{why does discretizer with equal bin size not give a better result?}
-
-The Naive Bayes does not have a lot of settings.
-It use a built in supervised discretizer or a kernel estimator.
-But the result is always the same with around 80\% correctly classified instances in the test run.
+By using different discretizers and also do not seperately discretize in the preprocessing section of weka but instead let the NaiveBayers classifier use it's own discretizers namely KernelEstimator or SupervisedDiscretizer the outcome of precision and recall differ a bit by 0.05 to 0.1 higher and lower.
+It does obviously do not really make a difference.
 

### 2012-11-25 paper.tex
added chapter about decision trees on contraceptive data set

--- 2012-11-25 paper.tex
@@
 If we rise the minimum of instances per leaves, e.g.
-to 20 instances and 50 instances, the precision falls from 86\% to 85\% and 84\$ respectively.
+to 20 instances and 50 instances, the precision falls from 86\% to 85\% and 84\% respectively.
 
@@
 
-\todo{todo marker}
+With default settings, the J48 algorithm comes off very badly, in table \ref{tab:contra:dec:default} one can see a that it classifies about the half of all instances correctly.
+It must be apologetically supplemented that the task includes more distinct values, three instead of two nominals, to be predicted.
+
+\begin{table*}[htb]\centering
+  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+      \toprule 
+      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+               &  TP Rate & FP Rate & Precision & Recall & F-Measure &  Class \\ \midrule 
+               &  0.609   & 0.287   & 0.613     & 0.609  & 0.611     &  1     \\  
+               &  0.378   & 0.138   & 0.445     & 0.378  & 0.409     &  2     \\      
+               &  0.507   & 0.318   & 0.458     & 0.507  & 0.481     &  3     \\      
+Weighted Avg.  &  0.521   & 0.264   & 0.521     & 0.521  & 0.52      &        \\ \bottomrule     
+    \end{tabular*}
+\caption{Decision Tree on Contraceptive Data Set -- Default Settings} 
+\label{tab:contra:dec:default}
+\end{table*}
+\FloatBarrier
+
+\paragraph{Raising the confidence factor}
+As expected, raising the confidence factor worsens the result.
+For a series of a factor 0.25, 0.5, 0.25 and a totally unpruned tree, the quality decreases from originally 0.52\% to 50.4\%, 48.6\% and finally 48.7\%.
+The used decision tree grows from a size of 157 leaves and a maximum depth of 263 to a huge tree with 417 leaves and 655 branches.
+As said, this is a definitely wrong direction of optimizing.
+
+\paragraph{Raising the minimum of instances per leaf}
+We can notice a certain improvement to the outcome if we ensure that every leaf of the decision tree has to contain more instances, thereby preventing the tree from deep nesting.
+We can make the most out of it if we set the minimum to 23 leaves, the yield can be seen in table \ref{tab:contra:dec:23l}.
+This tree has only 31 leaves and 49 branches.
+Further enhancing of this parameter cuts the amount of correctly classified instances again.
+
+\begin{table*}[htb]\centering
+  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+      \toprule 
+      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+              &   TP Rate & FP Rate & Precision & Recall & F-Measure &  Class \\ \midrule      
+              &   0.595   & 0.231   & 0.657     & 0.595  & 0.624     &  1     \\      
+              &   0.405   & 0.135   & 0.467     & 0.405  & 0.434     &  2     \\      
+              &   0.575   & 0.334   & 0.478     & 0.575  & 0.522     &  3     \\      
+Weighted Avg. &   0.545   & 0.245   & 0.552     & 0.545  & 0.546     &        \\ \bottomrule     
+    \end{tabular*}
+\caption{Decision Tree on Contraceptive Data Set -- at least 23 instances per leaf} 
+\label{tab:contra:dec:23l}
+\end{table*}
+\FloatBarrier
+\paragraph{Other parameters}
+Neither laplace correction nor mdl correction has any positive effect on the output of the algorithm on this data set.
+The absence of MDL correction make the outcome a bit worse, it brings the precision down to 0.551\%.
+But there is one small tweak from which the accuracy of the algorithm can benefit: If binary splits are used on nominal attributes, we get the best result for the J48 on this data set as seen in table \ref{tab:contra:dec:bin}.
+
+\begin{table*}[htb]\centering
+  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+      \toprule 
+      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+               &  TP Rate & FP Rate & Precision & Recall & F-Measure  Class \\ \midrule   
+               &  0.649   & 0.263   & 0.648     & 0.649  & 0.648      1     \\    
+               &  0.402   & 0.13    & 0.475     & 0.402  & 0.436      2     \\        
+               &  0.548   & 0.292   & 0.499     & 0.548  & 0.522      3     \\    
+Weighted Avg.  &  0.558   & 0.243   & 0.557     & 0.558  & 0.557            \\ \bottomrule 
+    \end{tabular*}
+\caption{Decision Tree on Contraceptive Data Set -- using binary splits} 
+\label{tab:contra:dec:bin}
+\end{table*}
+\FloatBarrier
 

### 2012-11-25 paper.tex
bugfix: label to ref

@@
 
-After the first run with the Simple Naive Bayes algorithm we gain the result shown in table \label{tab:adult:bayes:1d} and a confusion matrix \ref{tab:adult:bayes:1c}.
+After the first run with the Simple Naive Bayes algorithm we gain the result shown in table \ref{tab:adult:bayes:1d} and a confusion matrix \ref{tab:adult:bayes:1c}.
 

### 2012-11-25 paper.tex
nb for contraseptive data

--- 2012-11-25 paper.tex
@@
 If we rise the minimum of instances per leaves, e.g.
-to 20 instances and 50 instances, the precision falls from 86\% to 85\% and 84\% respectively.
+to 20 instances and 50 instances, the precision falls from 86\% to 85\% and 84\$ respectively.
 
@@
 These go from 1 to 4 where 1 means low/bad and 4 means high/good.
+The number-of-children attribute is numeric but the values range from 0 to 16 so that a preprocessing is not necessary as one can think of 15 classes (15 distinct values).
 This set has no missing values and the attributes do not depend on each other.
@@
 
-With default settings, the J48 algorithm comes off very badly, in table \ref{tab:contra:dec:default} one can see a that it classifies about the half of all instances correctly.
-It must be apologetically supplemented that the task includes more distinct values, three instead of two nominals, to be predicted.
-
-\begin{table*}[htb]\centering
-  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-      \toprule 
-      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
-               &  TP Rate & FP Rate & Precision & Recall & F-Measure &  Class \\ \midrule 
-               &  0.609   & 0.287   & 0.613     & 0.609  & 0.611     &  1     \\  
-               &  0.378   & 0.138   & 0.445     & 0.378  & 0.409     &  2     \\      
-               &  0.507   & 0.318   & 0.458     & 0.507  & 0.481     &  3     \\      
-Weighted Avg.  &  0.521   & 0.264   & 0.521     & 0.521  & 0.52      &        \\ \bottomrule     
-    \end{tabular*}
-\caption{Decision Tree on Contraceptive Data Set -- Default Settings} 
-\label{tab:contra:dec:default}
-\end{table*}
-\FloatBarrier
-
-\paragraph{Raising the confidence factor}
-As expected, raising the confidence factor worsens the result.
-For a series of a factor 0.25, 0.5, 0.25 and a totally unpruned tree, the quality decreases from originally 0.52\% to 50.4\%, 48.6\% and finally 48.7\%.
-The used decision tree grows from a size of 157 leaves and a maximum depth of 263 to a huge tree with 417 leaves and 655 branches.
-As said, this is a definitely wrong direction of optimizing.
-
-\paragraph{Raising the minimum of instances per leaf}
-We can notice a certain improvement to the outcome if we ensure that every leaf of the decision tree has to contain more instances, thereby preventing the tree from deep nesting.
-We can make the most out of it if we set the minimum to 23 leaves, the yield can be seen in table \ref{tab:contra:dec:23l}.
-This tree has only 31 leaves and 49 branches.
-Further enhancing of this parameter cuts the amount of correctly classified instances again.
-
-\begin{table*}[htb]\centering
-  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-      \toprule 
-      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
-              &   TP Rate & FP Rate & Precision & Recall & F-Measure &  Class \\ \midrule      
-              &   0.595   & 0.231   & 0.657     & 0.595  & 0.624     &  1     \\      
-              &   0.405   & 0.135   & 0.467     & 0.405  & 0.434     &  2     \\      
-              &   0.575   & 0.334   & 0.478     & 0.575  & 0.522     &  3     \\      
-Weighted Avg. &   0.545   & 0.245   & 0.552     & 0.545  & 0.546     &        \\ \bottomrule     
-    \end{tabular*}
-\caption{Decision Tree on Contraceptive Data Set -- at least 23 instances per leaf} 
-\label{tab:contra:dec:23l}
-\end{table*}
-\FloatBarrier
-\paragraph{Other parameters}
-Neither laplace correction nor mdl correction has any positive effect on the output of the algorithm on this data set.
-The absence of MDL correction make the outcome a bit worse, it brings the precision down to 0.551\%.
-But there is one small tweak from which the accuracy of the algorithm can benefit: If binary splits are used on nominal attributes, we get the best result for the J48 on this data set as seen in table \ref{tab:contra:dec:bin}.
-
-\begin{table*}[htb]\centering
-  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-      \toprule 
-      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
-               &  TP Rate & FP Rate & Precision & Recall & F-Measure  Class \\ \midrule   
-               &  0.649   & 0.263   & 0.648     & 0.649  & 0.648      1     \\    
-               &  0.402   & 0.13    & 0.475     & 0.402  & 0.436      2     \\        
-               &  0.548   & 0.292   & 0.499     & 0.548  & 0.522      3     \\    
-Weighted Avg.  &  0.558   & 0.243   & 0.557     & 0.558  & 0.557            \\ \bottomrule 
-    \end{tabular*}
-\caption{Decision Tree on Contraceptive Data Set -- using binary splits} 
-\label{tab:contra:dec:bin}
-\end{table*}
-\FloatBarrier
+\todo{todo marker}
 
@@
 
-\todo{todo marker}
+The input data set does perfectly fit for Naive Bayes.
+But the classifier is not able to correctly predict the instances.
+As shown in tables \ref{tab:contra:bayes:1d} and \ref{tab:contra:bayes:1c} it gains a 50:50 chance of a TP prediction.
+
+\begin{table*}[htb]\centering
+  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+      \toprule 
+      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+               &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class \\   \midrule
+               &  0.52    & 0.223   & 0.635     & 0.52   & 0.572     & 1  \\   \midrule
+               &  0.526   & 0.245   & 0.385     & 0.526  & 0.445     & 2  \\
+               &  0.481   & 0.268   & 0.488     & 0.481  & 0.485     & 3  \\
+Weighted Avg.  &  0.508   & 0.243   & 0.528     & 0.508  & 0.513     &    \\   \bottomrule
+    \end{tabular*}
+\caption{Naive Bayes} 
+\label{tab:contra:bayes:1d}
+\end{table*}
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|lll|}
+\hline
+              &   a &   b &   c  \\
+\hline
+ a = 1        & 327 & 141 & 161  \\
+ b = 2        &  61 & 175 &  97  \\
+ c = 3        & 127 & 138 & 246  \\
+\hline
+\end{tabular}
+\caption{Confusion matrix for Naive Bayes}
+\label{tab:contra:bayes:1c}
+\end{table}
 

### 2012-11-25 paper.tex
nb for contraseptive data

@@
 
-\todo[inline]{defined:
-
-validation: cross-validation with 10 for big data sets (9 training, 1 test), 3 for the small echo cardio set.
-}
-
 \todo[inline]{Choose 5 different classifiers, from at least 3 different types
@@
 \paragraph{}As data sets we chose as follows:
+
 \begin{itemize}
@@
 \subsection{Preprocessing}
+
 \paragraph{}The patient's attribute ``name'' was replaced by a dummy for all entries.
@@
 \paragraph{}The documentation of the data set explains the two attributes as follows:
+
 \begin{quote}
@@
 \end{quote}
+
 \begin{quote}
@@
 Only one false positive has gone. Not really a remarkable change in the quality. This does not change if the tree-depth is further reduced.
+
 \paragraph{Increasing the number of features}
@@
 
-As with the SVM-algorithm there was no normalization necessary (we tested this on different datasets).
+\paragraph{}As with the SVM-algorithm there was no normalization necessary (we tested this on different datasets).
 In WEKA there are different implementations of the kNN-algorithm available and we chose (for all our datasets) the brute force kNN-algorithm since this is the classic implementation of the kNN-algorithm where we were interested in the running time.
@@
 
-For the fist try, the unsupervised discretizer was used. But the result was bad. The classifier could predict one single instance for for patients that did not survive one year.
+\paragraph{}For the fist try, the unsupervised discretizer was used. But the result was bad. The classifier could predict one single instance for for patients that did not survive one year.
 
-For the second try we prepare the data set with the PKIDiscretizer.
+\paragraph{}For the second try we prepare the data set with the PKIDiscretizer.
 The description says it "Discretizes numeric attributes using equal frequency binning, where the number of bins is equal to the square root of the number of non-missing values." But as we can see in \ref{tab:echo:bayes:1c} this discretizer does not perform better.
@@
 
-The log output of Naive Bayes shows that each bin has 1-2 negative (alive-at-1 = 0) instances. This means that each bin has the same low possibility for TP-prediction.
+\paragraph{}The log output of Naive Bayes shows that each bin has 1-2 negative (alive-at-1 = 0) instances. This means that each bin has the same low possibility for TP-prediction.
 
-So we have to boost the instances we want to predict.
+\paragraph{}So we have to boost the instances we want to predict.
 For that the SMOTE filter is used.
@@
 
-We have to supply a separate set for validation as we kind of destroy the data set when generating new instances.
+\paragraph{}We have to supply a separate set for validation as we kind of destroy the data set when generating new instances.
 For that the original training set was taken.
@@
 
-The confusion matrix \ref{tab:echo:bayes:2c} shows, that we predicted three out of four correct instances and also had more errors when predicting alive-at-1 = 0 than  alive-at-1 = 1.
+\paragraph{}The confusion matrix \ref{tab:echo:bayes:2c} shows, that we predicted three out of four correct instances and also had more errors when predicting alive-at-1 = 0 than  alive-at-1 = 1.
 
@@
 
-Also the detailed accuracy table \ref{tab:echo:bayes:2d} shows that this approach gained a high recall with a low precision for alive-at-1 = 0 and both high precision and recall for alive-at-1 = 1.
+\paragraph{}Also the detailed accuracy table \ref{tab:echo:bayes:2d} shows that this approach gained a high recall with a low precision for alive-at-1 = 0 and both high precision and recall for alive-at-1 = 1.
 
@@
 
-The ``Adult'' dataset has missing values which we all replaced by the mean value of the respective attribute.
+\paragraph{}The ``Adult'' dataset has missing values which we all replaced by the mean value of the respective attribute.
 We did no further preprocessing.
@@
 \subsection{Decision Tree}
-Using a decision tree yields a precision about 86\%. Varying the parameters for this algorithm influences the quality of the results only slightly, as the following examples illustrate.
+\paragraph{}Using a decision tree yields a precision about 86\%. Varying the parameters for this algorithm influences the quality of the results only slightly, as the following examples illustrate.
 
-First of all, with the default settings of the J48 implementation of weka, i.e.
+\paragraph{}First of all, with the default settings of the J48 implementation of weka, i.e.
 using pruning with a confidence factor of 0.25 together with MDLCorrection and Subtree raising and at least 2 instances per leaf of the tree, we get a summary of the quality as shown in table \ref{tab:adult:dec:default}:
@@
 
-The decision tree that was used for this run has a size of 413 and has 311 leaves.
+\paragraph{}The decision tree that was used for this run has a size of 413 and has 311 leaves.
 
@@
 Which is a quite similar accuracy like the best of the J48 algorithm.
+
 \subsection{Support Vector Machine}
 
-From the different kernel functions for the SVM-algorithm we used the PolyKernel on this dataset.
+\paragraph{}From the different kernel functions for the SVM-algorithm we used the PolyKernel on this dataset.
 As with the dataset before the normalization of the numeric data had no inpact on the validation results.
@@
 
-As further preprocessing for Naive Bayes it is important to dicretize the numerical attributes.
+\paragraph{}As further preprocessing for Naive Bayes it is important to dicretize the numerical attributes.
 The PKIDescretizer was used again as it produces better bin sizes than the standard discretizer without any configuration.
 
-After the first run with the Simple Naive Bayes algorithm we gain the result shown in table \label{tab:adult:bayes:1d} and a confusion matrix \ref{tab:adult:bayes:1c}.
+\paragraph{}After the first run with the Simple Naive Bayes algorithm we gain the result shown in table \label{tab:adult:bayes:1d} and a confusion matrix \ref{tab:adult:bayes:1c}.
 
@@
 
-By using different discretizers and also do not seperately discretize in the preprocessing section of weka but instead let the NaiveBayers classifier use it's own discretizers namely KernelEstimator or SupervisedDiscretizer the outcome of precision and recall differ a bit by 0.05 to 0.1 higher and lower.
+\paragraph{}By using different discretizers and also do not seperately discretize in the preprocessing section of weka but instead let the NaiveBayers classifier use it's own discretizers namely KernelEstimator or SupervisedDiscretizer the outcome of precision and recall differ a bit by 0.05 to 0.1 higher and lower.
 It does obviously do not really make a difference.
@@
 
-We tested the k-nearest neighbour algorithm with different values for $k$.
+\paragraph{}We tested the k-nearest neighbour algorithm with different values for $k$.
 For $k$=1 (nearest neighbour) the algorithm was able to correctly classify 79,4\% of the instances.
@@
 
-The Contraceptive Method Choice Data Set consists of information about the demographic and socio-economic characteristics of married women.
+\paragraph{}The Contraceptive Method Choice Data Set consists of information about the demographic and socio-economic characteristics of married women.
 There are 1473 instances with 9 different attributes.
@@
 
-The values of the data sets's attributes are already in a numerical and categorical form, whereby the categories are encoded in an ordinal form.
+\paragraph{}The values of the data sets's attributes are already in a numerical and categorical form, whereby the categories are encoded in an ordinal form.
 These go from 1 to 4 where 1 means low/bad and 4 means high/good.
@@
 
-The input data set does perfectly fit for Naive Bayes.
+\paragraph{}The input data set does perfectly fit for Naive Bayes.
 But the classifier is not able to correctly predict the instances.
@@
 
-We tested different kernel functions on this dataset and obtained the best results with the NormalizedPolyKernel.
+\paragraph{}We tested different kernel functions on this dataset and obtained the best results with the NormalizedPolyKernel.
 The whole process of the evaluation, including the building of the models, took 28 seconds.
@@
 
-We tested the kNN-algorithm with a lot of different values for $k$ and obtained the best results for $k=7$ where 48,88\% of the instances were correctly classified.
+\paragraph{}We tested the kNN-algorithm with a lot of different values for $k$ and obtained the best results for $k=7$ where 48,88\% of the instances were correctly classified.
 The evaluation took not even a second.

### 2012-11-25 paper.tex

@@
 The PKIDescretizer was used again as it produces better bin sizes than the standard discretizer without any configuration.
-
-\paragraph{}After the first run with the Simple Naive Bayes algorithm we gain the result shown in table \label{tab:adult:bayes:1d} and a confusion matrix \ref{tab:adult:bayes:1c}.
+After the first run with the Simple Naive Bayes algorithm we gain the result shown in table \ref{tab:adult:bayes:1d} and a confusion matrix \ref{tab:adult:bayes:1c}.
 

### 2012-11-25 paper.tex
added chapter about random forest for contraceptive

--- 2012-11-25 paper.tex
@@
 These go from 1 to 4 where 1 means low/bad and 4 means high/good.
-The number-of-children attribute is numeric but the values range from 0 to 16 so that a preprocessing is not necessary as one can think of 15 classes (15 distinct values).
 This set has no missing values and the attributes do not depend on each other.
@@
 
-\todo{todo marker}
-
-\subsection{Random Forest}
-
-\todo{todo marker}
-
-\subsection{Naive Bayes}
-
-The input data set does perfectly fit for Naive Bayes.
-But the classifier is not able to correctly predict the instances.
-As shown in tables \ref{tab:contra:bayes:1d} and \ref{tab:contra:bayes:1c} it gains a 50:50 chance of a TP prediction.
+With default settings, the J48 algorithm comes off very badly, in table \ref{tab:contra:dec:default} one can see a that it classifies about the half of all instances correctly.
+It must be apologetically supplemented that the task includes more distinct values, three instead of two nominals, to be predicted.
 
@@
                &  TP Rate & FP Rate & Precision & Recall & F-Measure &  Class \\ \midrule 
-               &  0.52    & 0.223   & 0.635     & 0.52   & 0.572     & 1  \\   \midrule
-               &  0.526   & 0.245   & 0.385     & 0.526  & 0.445     & 2  \\
-               &  0.481   & 0.268   & 0.488     & 0.481  & 0.485     & 3  \\
-Weighted Avg.  &  0.508   & 0.243   & 0.528     & 0.508  & 0.513     &    \\   \bottomrule
+               &  0.609   & 0.287   & 0.613     & 0.609  & 0.611     &  1     \\  
+               &  0.378   & 0.138   & 0.445     & 0.378  & 0.409     &  2     \\      
+               &  0.507   & 0.318   & 0.458     & 0.507  & 0.481     &  3     \\      
+Weighted Avg.  &  0.521   & 0.264   & 0.521     & 0.521  & 0.52      &        \\ \bottomrule     
     \end{tabular*}
-\caption{Naive Bayes} 
-\label{tab:contra:bayes:1d}
+\caption{Decision Tree on Contraceptive Data Set -- Default Settings} 
+\label{tab:contra:dec:default}
 \end{table*}
+\FloatBarrier
 
-\begin{table}[h]
-\centering
-\begin{tabular}{|l|lll|}
-\hline
-              &   a &   b &   c  \\
-\hline
- a = 1        & 327 & 141 & 161  \\
- b = 2        &  61 & 175 &  97  \\
- c = 3        & 127 & 138 & 246  \\
-\hline
-\end{tabular}
-\caption{Confusion matrix for Naive Bayes}
-\label{tab:contra:bayes:1c}
-\end{table}
+\paragraph{Raising the confidence factor}
+As expected, raising the confidence factor worsens the result.
+For a series of a factor 0.25, 0.5, 0.25 and a totally unpruned tree, the quality decreases from originally 0.52\% to 50.4\%, 48.6\% and finally 48.7\%.
+The used decision tree grows from a size of 157 leaves and a maximum depth of 263 to a huge tree with 417 leaves and 655 branches.
+As said, this is a definitely wrong direction of optimizing.
+
+\paragraph{Raising the minimum of instances per leaf}
+We can notice a certain improvement to the outcome if we ensure that every leaf of the decision tree has to contain more instances, thereby preventing the tree from deep nesting.
+We can make the most out of it if we set the minimum to 23 leaves, the yield can be seen in table \ref{tab:contra:dec:23l}.
+This tree has only 31 leaves and 49 branches.
+Further enhancing of this parameter cuts the amount of correctly classified instances again.
+
+\begin{table*}[htb]\centering
+  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+      \toprule 
+      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+              &   TP Rate & FP Rate & Precision & Recall & F-Measure &  Class \\ \midrule      
+              &   0.595   & 0.231   & 0.657     & 0.595  & 0.624     &  1     \\      
+              &   0.405   & 0.135   & 0.467     & 0.405  & 0.434     &  2     \\      
+              &   0.575   & 0.334   & 0.478     & 0.575  & 0.522     &  3     \\      
+Weighted Avg. &   0.545   & 0.245   & 0.552     & 0.545  & 0.546     &        \\ \bottomrule     
+    \end{tabular*}
+\caption{Decision Tree on Contraceptive Data Set -- at least 23 instances per leaf} 
+\label{tab:contra:dec:23l}
+\end{table*}
+\FloatBarrier
+\paragraph{Other parameters}
+Neither laplace correction nor mdl correction has any positive effect on the output of the algorithm on this data set.
+The absence of MDL correction make the outcome a bit worse, it brings the precision down to 0.551\%.
+But there is one small tweak from which the accuracy of the algorithm can benefit: If binary splits are used on nominal attributes, we get the best result for the J48 on this data set as seen in table \ref{tab:contra:dec:bin}.
+
+\begin{table*}[htb]\centering
+  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+      \toprule 
+      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+               &  TP Rate & FP Rate & Precision & Recall & F-Measure &  Class \\ \midrule   
+               &  0.649   & 0.263   & 0.648     & 0.649  & 0.648     &  1     \\    
+               &  0.402   & 0.13    & 0.475     & 0.402  & 0.436     &  2     \\        
+               &  0.548   & 0.292   & 0.499     & 0.548  & 0.522     &  3     \\    
+Weighted Avg.  &  0.558   & 0.243   & 0.557     & 0.558  & 0.557     &        \\ \bottomrule 
+    \end{tabular*}
+\caption{Decision Tree on Contraceptive Data Set -- using binary splits} 
+\label{tab:contra:dec:bin}
+\end{table*}
+\FloatBarrier
+
+\subsection{Random Forest}
+Random Forests perform even worse than the J48 with default settings. We can see in table \ref{tab:contra:rand:default} that only the half of all classified instances are classified correctly.
+
+\begin{table*}[htb]\centering
+  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+      \toprule 
+      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+               &  TP Rate & FP Rate & Precision & Recall & F-Measure  & Class  \\  \midrule      
+               &  0.631   & 0.328   & 0.589     & 0.631  & 0.609      & 1      \\       
+               &  0.36    & 0.16    & 0.397     & 0.36   & 0.378      & 2      \\       
+               &  0.456   & 0.274   & 0.469     & 0.456  & 0.462      & 3      \\       
+Weighted Avg.  &  0.509   & 0.271   & 0.504     & 0.509  & 0.506      &        \\  \bottomrule     
+    \end{tabular*}
+\caption{Random Forests on Contraceptive Data Set -- Default Settings} 
+\label{tab:contra:rand:default}
+\end{table*}
+\FloatBarrier
+
+\paragraph{More but shorter trees}
+Random forests perform best on the data set with a maximum depth of 5 branches.
+After this improvement we can lift the precision up to 54.4\%.
+Furthermore, our random forest benefits from adding five more random trees so we can enhance the precision to  54.7\%.
+\paragraph{More Features}
+The most significant improvement can then observed after increasing the number of randomly chosen features.
+By default, the number of chosen features for this data set lays about 4 ( $ log_2(10) + 1 $ ).
+Choosing two more of them results in the best accuracy for random forests on this data set, as shown in table \ref{tab:contra:rand:6f}.
+
+\begin{table*}[htb]\centering
+  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+      \toprule 
+      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+               &  TP Rate & FP Rate & Precision & Recall & F-Measure &  Class \\  \midrule     
+               &  0.604   & 0.231   & 0.661     & 0.604  & 0.631     &  1     \\  
+               &  0.408   & 0.127   & 0.484     & 0.408  & 0.443     &  2     \\  
+               &  0.593   & 0.326   & 0.491     & 0.593  & 0.537     &  3     \\  
+Weighted Avg.  &  0.556   & 0.241   & 0.562     & 0.556  & 0.556     &        \\  \bottomrule 
+    \end{tabular*}
+\caption{Random Forests on Contraceptive Data Set -- with at least 6 chosen features} 
+\label{tab:contra:rand:6f}
+\end{table*}
+\FloatBarrier
+
+\subsection{Naive Bayes}
+
+\todo{todo marker}
 

### 2012-11-25 paper.tex

--- 2012-11-25 paper.tex
@@
       \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
-               &  TP Rate & FP Rate & Precision & Recall & F-Measure  Class \\ \midrule   
-               &  0.649   & 0.263   & 0.648     & 0.649  & 0.648      1     \\    
-               &  0.402   & 0.13    & 0.475     & 0.402  & 0.436      2     \\        
-               &  0.548   & 0.292   & 0.499     & 0.548  & 0.522      3     \\    
-Weighted Avg.  &  0.558   & 0.243   & 0.557     & 0.558  & 0.557            \\ \bottomrule 
+               &  TP Rate & FP Rate & Precision & Recall & F-Measure &  Class \\ \midrule   
+               &  0.649   & 0.263   & 0.648     & 0.649  & 0.648     &  1     \\    
+               &  0.402   & 0.13    & 0.475     & 0.402  & 0.436     &  2     \\        
+               &  0.548   & 0.292   & 0.499     & 0.548  & 0.522     &  3     \\    
+Weighted Avg.  &  0.558   & 0.243   & 0.557     & 0.558  & 0.557     &        \\ \bottomrule 
     \end{tabular*}
@@
 \subsection{Random Forest}
-
-\todo{todo marker}
+Random Forests perform even worse than the J48 with default settings. We can see in table \ref{tab:contra:rand:default} that only the half of all classified instances are classified correctly.
+
+\begin{table*}[htb]\centering
+  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+      \toprule 
+      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+               &  TP Rate & FP Rate & Precision & Recall & F-Measure  & Class  \\  \midrule      
+               &  0.631   & 0.328   & 0.589     & 0.631  & 0.609      & 1      \\       
+               &  0.36    & 0.16    & 0.397     & 0.36   & 0.378      & 2      \\       
+               &  0.456   & 0.274   & 0.469     & 0.456  & 0.462      & 3      \\       
+Weighted Avg.  &  0.509   & 0.271   & 0.504     & 0.509  & 0.506      &        \\  \bottomrule     
+    \end{tabular*}
+\caption{Random Forests on Contraceptive Data Set -- Default Settings} 
+\label{tab:contra:rand:default}
+\end{table*}
+\FloatBarrier
+
+\paragraph{More but shorter trees}
+Random forests perform best on the data set with a maximum depth of 5 branches.
+After this improvement we can lift the precision up to 54.4\%.
+Furthermore, our random forest benefits from adding five more random trees so we can enhance the precision to  54.7\%.
+\paragraph{More Features}
+The most significant improvement can then observed after increasing the number of randomly chosen features.
+By default, the number of chosen features for this data set lays about 4 ( $ log_2(10) + 1 $ ).
+Choosing two more of them results in the best accuracy for random forests on this data set, as shown in table \ref{tab:contra:rand:6f}.
+
+\begin{table*}[htb]\centering
+  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+      \toprule 
+      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+               &  TP Rate & FP Rate & Precision & Recall & F-Measure &  Class \\  \midrule     
+               &  0.604   & 0.231   & 0.661     & 0.604  & 0.631     &  1     \\  
+               &  0.408   & 0.127   & 0.484     & 0.408  & 0.443     &  2     \\  
+               &  0.593   & 0.326   & 0.491     & 0.593  & 0.537     &  3     \\  
+Weighted Avg.  &  0.556   & 0.241   & 0.562     & 0.556  & 0.556     &        \\  \bottomrule 
+    \end{tabular*}
+\caption{Random Forests on Contraceptive Data Set -- with at least 6 chosen features} 
+\label{tab:contra:rand:6f}
+\end{table*}
+\FloatBarrier
 

### 2012-11-25 paper.tex
minor change: changed todos to inline todos to avoid latex warnings
added picture with crappy entry

@@
 
+\usepackage{color}
 \usepackage{float}
@@
 \usepackage{listings}
+\lstset{
+    captionpos=b
+}
+\newcommand{\Hilight}{\makebox[0pt][l]{\color{red}\rule[-4pt]{0.65\linewidth}{14pt}}}
 \usepackage{booktabs}
@@
 Most of the data was complete but there were still some data records having missing values which we replaced by the mean values of the specific variables.
-One data record did contain completeley useless data and was thus removed (see at the picture below).
+One data record did contain completeley useless data and was thus removed (see at in listing \ref{lst:crappy-entry}).
+
+\begin{lstlisting}[escapechar=\%, float=htb, caption={76 years old patient without useful information}, label=lst:crappy-entry]
+20,1,59,0,0.030,21.300,6.290,17,1.310,0.928,name,2,0
+0.25,1,63,1,?,?,?,23,2.300,0.714,name,2,1
+%\Hilight% ?,?,?,77,?,?,?,?,?,2,?,name,2,?
+2,1,56,1,0.040,14,5,?,?,?,name,2,1
+7,1,61,1,0.270,?,?,9,1.500,0.428,name,2,1
+\end{lstlisting}
 
@@
 
-\todo{Compare results with/without preprocessing}
+\todo[inline]{Compare results with/without preprocessing}
 
 \subsection{Decision Tree}
-\todo{robert: do you want to keep this section?
+\todo[inline]{robert: do you want to keep this section?
 on the one hand it shows what happens if we run a desicion tree with attributes on which the the prediction attribute depends, but on the other hand we know we should not do that.

### 2012-11-25 paper.tex
replaced the todo addressed to my by merging preprocessed subsection and first decision tree section

@@
 \paragraph{}It should be clear that the variables ``survival'' and ``still-alive'' should not be used for classification since they constitute the final prediction attribute.
+Listing \ref{lst:silly-tree} shows what happens if a decision tree can take these attributes into account.
+It states that if the patient survived $ <= 12 $ months, he survived at least 12 month.
+Pretty accurate, right?
+The only remarkable thing is, that the J48 implementation under default settings seems to enforce a certain depth of the tree, even if the splitting on the same attribute is redundant since $ x > 11 \rightarrow x > 12 $.
+
+\begin{lstlisting}[caption={Decision tree on 100\% dependant attributes}, label=lst:silly-tree, float=htb]
+J48 pruned tree
+------------------
+    survival <= 12
+    |   survival <= 11: 0 (4.0)
+    |   survival > 11: 1 (4.0)
+    survival > 12: 1 (87.0)
+\end{lstlisting}
 
@@
 Most of the data was complete but there were still some data records having missing values which we replaced by the mean values of the specific variables.
-One data record did contain completeley useless data and was thus removed (see at in listing \ref{lst:crappy-entry}).
+One data record did contain completeley useless data and was thus removed (see in listing \ref{lst:crappy-entry}).
 
@@
 \subsection{Decision Tree}
-\todo[inline]{robert: do you want to keep this section?
-on the one hand it shows what happens if we run a desicion tree with attributes on which the the prediction attribute depends, but on the other hand we know we should not do that.
-please decide.
-i'm ok with both (philipp)}
-%------------------------------------------------
-On the first glimpse, it is an easy task for a decision based algorithm:
-
-\begin{lstlisting}
-J48 pruned tree
-------------------
-survival <= 12
-|   survival <= 11: 0 (4.0)
-|   survival > 11: 1 (4.0)
-survival > 12: 1 (87.0)
-\end{lstlisting}
-
-\paragraph{}That's correct, if the patient survived $ <= 12 $ months, he survived at least 12 month. What a statement!
-
-\paragraph{}The only remarkable thing is, that the J48 implementation under default settings seems to enforce a certain depth of the tree, even if the splitting on the same attribute don't make any sense ($ 11 $ is yet a smaller number than $ 12 $, isn't it?).
-E.g.
-choosing a DecisionStump results in the expected single splitting:
-\begin{lstlisting}
-Class distributions
-survival <= 11.5
-0   1   
-1.0 0.0 
-survival > 11.5
-0   1   
-0.0 1.0 
-survival is missing
-0   1   
-0.042105263157894736    0.9578947368421052
-\end{lstlisting}
-\paragraph{}Since there is no missing value for the survival attribute, this branch can be ignored.
-
-\paragraph{}But allowing the machine learning algorithms to see the origin attributes from which we derived the prediction attribut leads to absurdity, since the actual prediction task is to say whether a patient will survive one year after suffering a heart attack or not \emph{depending on the medical characteristics}.
-
-\subsection{Decision Tree on preprocessed data}
-
-\paragraph{}The J48 Decision Tree shows a more useful tree on the preprocessed data:
+
+The J48 Decision Tree shows a more useful tree on the preprocessed data:
 

### 2012-11-25 paper.tex
looked through the text and made it fit to the shown tables and figures

@@
 \usepackage{booktabs}
+\usepackage{multirow}
 \usepackage{morefloats}
@@
 
-The J48 Decision Tree shows a more useful tree on the preprocessed data:
-
-\begin{lstlisting}
+With default settings J48 Decision Tree seems to show a reasonable tree to some extent as seen in listing \ref{lst:echo:dec:default}.
+
+\begin{lstlisting}[label=lst:echo:dec:default, caption={Vizualization of Decision Tree on echocardiogram data -- Default Settings}]
 J48 pruned tree
@@
 
-\begin{table}[h]
-\centering
-\begin{tabular}{|l|ll|}
-\hline
- & a &	b \\
-\hline
-a = 0  & 0 & 4 	\\
-b = 1 & 1 & 90 	\\
-\hline
+\begin{table*}[htb]\centering
+    \begin{tabular}{@{}ll|ll@{}} 
+\multicolumn{2}{c}{\phantom{bla}}  & \multicolumn{2}{c}{Classified as} \\  
+                        & \phantom{aa}  &  0  &  1  \\ \cmidrule{2-4}
+\multirow{2}{*}{Instances}     &  0  &  0  &  4  \\
+                                   &  1  &  1  &  90 \\ 
 \end{tabular}
 \caption{Confusion matrix for Decision tree}
-\end{table}
+\end{table*}
 
@@
 With default settings, i.e.
-default number of features $ log_2(7) + 1 \approx 4 $, and 10 random trees with unlimited tree depth, the random forest shows us a rather disappointing result:
-
-\begin{table}[h]
-\centering
-\begin{tabular}{|l|ll|}
-\hline
- & a &	b \\
-\hline
-a=0 & 0 & 4 \\
-b=1 & 2 & 91\\
-\hline
+default number of features $ log_2(7) + 1 \approx 4 $, and 10 random trees with unlimited tree depth, the random forest shows us a rather disappointing result as shown in the confusion matrix in table \ref{tab:echo:rand:matrix1} and the accuracy details in table \ref{tab:echo:rand:default}:
+
+\begin{table*}[htb]\centering
+    \begin{tabular}{@{}ll|ll@{}} 
+\multicolumn{2}{c}{\phantom{bla}}  & \multicolumn{2}{c}{Classified as} \\  
+                   & \phantom{aa}        &  0  &  1  \\ \cmidrule{2-4}
+\multirow{2}{*}{Instances}         &  0  &  0  &  4  \\
+                                   &  1  &  2  &  91 \\ 
 \end{tabular}
 \caption{Confusion matrix for Random Forest}
-\end{table}
-
-And the statistical measures:
+\label{tab:echo:rand:matrix1}
+\end{table*}
+
 \begin{table*}[htb]\centering
@@
 \caption{Random Forest -- Default Settings} 
+\label{tab:echo:rand:default}
 \end{table*}
@@
 Limitations in the depths of the random trees has also almost no effect on the accuracy of random forests on this data set.
-The only change to the outcome can be observed as soon the depth of the trees is limited to 6:
-
-\begin{table}[h]
-\centering
-\begin{tabular}{|l|ll|}
-\hline
- & a &	b \\
-\hline
-a=0 & 0 & 4 \\
-b=1 & 1 & 92\\
-\hline
+The only change to the outcome can be observed as soon the depth of the trees is limited to 6 (see table \ref{tab:echo:rand:matrix2}).
+Only one false positive has gone. Not really a remarkable change in the quality. This does not change if the tree-depth is further reduced.
+\begin{table*}[htb]\centering
+    \begin{tabular}{@{}ll|ll@{}} 
+\multicolumn{2}{c}{\phantom{bla}}  & \multicolumn{2}{c}{Classified as} \\  
+                   & \phantom{aa}        &  0  &  1  \\ \cmidrule{2-4}
+\multirow{2}{*}{Instances}         &  0  & 0 & 4 \\
+                                   &  1  & 1 & 92\\ 
 \end{tabular}
 \caption{Confusion matrix for Random Forest, limited depth}
-\end{table}
-
-Only one false positive has gone. Not really a remarkable change in the quality. This does not change if the tree-depth is further reduced.
+\label{tab:echo:rand:matrix2}
+\end{table*}
 
@@
 Increasing the number of features chosen by each random tree seems to has also a low effect on the quality, because until a certain border of 5 chosen features the outcome is exactly the same as it was with default settings.
-But if we configure the random forests with 6 features we can spot our first true positive:
-
-\begin{table}[h]
-\centering
-\begin{tabular}{|l|ll|}
-\hline
- & a &	b \\
-\hline
-a=0 & 1 & 3 \\
-b=1 & 2 & 91\\
-\hline
+But if we configure the random forests with 6 features we can spot our first true positive in the confusion matrix (table \ref{tab:echo:rand:matrix3}).
+This gives us the best achievable accuracy that we can reach with decision based algorithms, shown in table \ref{tab:echo:rand:6f}.
+
+\begin{table*}[htb]\centering
+    \begin{tabular}{@{}ll|ll@{}} 
+\multicolumn{2}{c}{\phantom{bla}}  & \multicolumn{2}{c}{Classified as} \\  
+                   & \phantom{aa}        &  0  &  1  \\ \cmidrule{2-4}
+\multirow{2}{*}{Instances}         &  0  & 1 & 3 \\
+                                   &  1  & 2 & 91\\ 
 \end{tabular}
 \caption{Confusion matrix for Random Forest, increased number of features}
-\end{table}
-
-Which gives us the best achievable accuracy that we can reach with decision based algorithms, shown in table \ref{tab:echo:rand:6f}.
+\label{tab:echo:rand:matrix3}
+\end{table*}
+
 \begin{table*}[htb]\centering
@@
 \paragraph{}First of all, with the default settings of the J48 implementation of weka, i.e.
-using pruning with a confidence factor of 0.25 together with MDLCorrection and Subtree raising and at least 2 instances per leaf of the tree, we get a summary of the quality as shown in table \ref{tab:adult:dec:default}:
+using pruning with a confidence factor of 0.25 together with MDL correction and subtree raising and at least 2 instances per leaf of the tree, we get a summary of the quality as shown in table \ref{tab:adult:dec:default}:
+The decision tree that was used for this run has a size of 413 and has 311 leaves.
 
@@
 \FloatBarrier
-
-\paragraph{}The decision tree that was used for this run has a size of 413 and has 311 leaves.
 
@@
 
-With default settings, a random forest gains an accuracy on the census income data set that can be seen in table \ref{tab:adult:rand:default}:
+With default settings, a random forest gains an accuracy on the census income data set that can be seen in table \ref{tab:adult:rand:default}, which is surprisingly worse than the J48 algorithm.
 \begin{table*}[htb]\centering
@@
 \FloatBarrier
-which is surprisingly worse than the J48 algorithm.
 
@@
 We have a sign for overfitting under default settings: Reducing the maximum depth has a positive effect on the quality.
-From size 20 over 10 until the size of 7 the precision of the random forest rises until 85\%.
+From size 20 over 10 until the size of 7 the precision of the randomeforest rises until 85\%.
 Reducing the depth even further lets the precision fall again.
-Here is the detailed accuracy for a random forest which can grow only to a height of 6:
-
-\begin{table*}[htb]\centering
-  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-      \toprule 
-      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
-              &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class    \\  \midrule
-              &  0.606   & 0.064   & 0.749     & 0.606  & 0.67      & >50K     \\ 
-              &  0.936   & 0.394   & 0.882     & 0.936  & 0.908     & <=50K    \\ 
-Weighted Avg. &  0.856   & 0.314   & 0.85      & 0.856  & 0.851     &          \\  \bottomrule
-    \end{tabular*}
-\caption{Random Forest -- 7 Random Trees} 
-\label{tab:adult:rand:7t}
-\end{table*}
-\FloatBarrier
-We can push the accuracy even further, if we take more random trees, e.g. twenty:
-\begin{table*}[htb]\centering
-  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-      \toprule 
-      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
-                & TP Rate  & FP Rate  & Precision &  Recall &  F-Measure  & Class  \\  \midrule          
-                & 0.593    & 0.058    & 0.763     &  0.593  &  0.667      & >50K   \\     
-                & 0.942    & 0.407    & 0.879     &  0.942  &  0.909      & <=50K  \\     
-Weighted Avg.   & 0.858    & 0.323    & 0.851     &  0.858  &  0.851      &        \\  \bottomrule
-    \end{tabular*}
-\caption{Random Forest -- 20 Random Trees} 
-\label{tab:adult:rand:20t}
+Table \ref{tab:adult:rand:6d} shows the detailed accuracy for a random forest of seven trees which can grow only to a height of 6:
+
+\begin{table*}[htb]\centering
+  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+      \toprule 
+      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+               &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class \\ \midrule
+               &  0.591   & 0.057   & 0.766     & 0.591  & 0.667     & >50K  \\
+               &  0.943   & 0.409   & 0.879     & 0.943  & 0.91      & <=50K \\
+Weighted Avg.  &  0.858   & 0.324   & 0.852     & 0.858  & 0.851     &       \\ \bottomrule
+    \end{tabular*}
+\caption{Random Forest -- 7 Random Trees each not more than 6 branches high} 
+\label{tab:adult:rand:6d}
 \end{table*}
@@
 To be concise, we get the best achievable result, if we take only 3 randomly chosen
-features.
+features. Table \ref{tab:adult:rand:3f} shows these results which are of similar accuracy like the best of the J48 algorithm.
+
 \begin{table*}[htb]\centering
@@
 \FloatBarrier
-Which is a quite similar accuracy like the best of the J48 algorithm.
 

### 2012-11-25 paper.tex
Changed some tables

@@
 \centering
-\begin{tabular}{|llllllll|}
-	\hline
+\begin{tabular}{llllllll}
+	\toprule
 									&TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure   &ROC Area  &Class\\
-	\hline
+	\midrule
 									&0         	&0         &0         	&0       &  0         & 0.5      &0\\
   								&1         	&1         &0.959     	&1       &0.979       &	0.5      &1\\
-  \hline
+
 	Weighted Avg.		&0.959     	&0.959     &0.919   		&0.959   &0.939     	&0.5	&\\
-	\hline
+	\bottomrule
 \end{tabular}
@@
 \centering
-\begin{tabular}{|llllllll|}
-	\hline
+\begin{tabular}{llllllll}
+	\toprule
 									&TP Rate   	&FP Rate   &Precision &Recall  &F-Measure   &ROC Area  &Class\\
-	\hline
+	\midrule
 									&0         	&0         &0         &0       &  0         & 0.5      &0\\
   								&1         	&1         &0.959     &1       &0.979       &	0.5      &1\\
-  \hline
+
 	Weighted Avg.		&0.959     	&0.959     &0.919   	&0.959   &0.939     	&0.5				&\\
-	\hline
+	\bottomrule
 \end{tabular}
@@
 \centering
-\begin{tabular}{|llllllll|}
-	\hline
+\begin{tabular}{llllllll}
+	\toprule
 									&TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure &ROC Area  	&Class\\
-	\hline
+	\midrule
 									&0.568     	&0.062     &0.744     	&0.568   &0.644    	&0.753    	&	>50K\\
   								&0.938     	&0.432     &0.873     	&0.938   &0.904     &0.753    	&$\leq$50K\\
-  \hline
+
 	Weighted Avg.   & 0.849    & 0.343     &0.842   		&0.849   &0.842     & 0.753			&\\
-	\hline
+	\bottomrule
 \end{tabular}
@@
 \centering
-\begin{tabular}{|llllllll|}
-	\hline
+\begin{tabular}{llllllll}
+	\toprule
 									&TP Rate   &FP Rate   &Precision   &Recall  &F-Measure &ROC Area  	&Class\\
-	\hline
+	\midrule
 									&0.593     &0.088      &0.68      &0.593     &0.633    &0.884    		&>50K\\
 									&0.912     &0.407      &0.876     &0.912     &0.893    &0.884    		&$\leq$50K\\
-  \hline
+
 	Weighted Avg.   &0.835     &0.331      &0.829     &0.835     &0.831    &0.884 \\
-	\hline
+	\bottomrule
 \end{tabular}
@@
 
-\subsection{Desicion Tree}
+\subsection{Decision Tree}
 
@@
                &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class \\   \midrule
-               &  0.52    & 0.223   & 0.635     & 0.52   & 0.572     & 1  \\   \midrule
+               &  0.52    & 0.223   & 0.635     & 0.52   & 0.572     & 1  \\  
                &  0.526   & 0.245   & 0.385     & 0.526  & 0.445     & 2  \\
@@
 
-\subsection{Support VectorMachine}
+\subsection{Support Vector Machine}
 
@@
 
-\begin{table}[h]
-\centering
-\begin{tabular}{|llllllll|}
-	\hline
+\begin{table}[htb]
+\centering
+\begin{tabular}{llllllll}
+	\toprule
 	&								TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure   &ROC Area  	&Class\\ 
-	\hline
+	\midrule
 									&0.583     	&0.355     &0.55      	&0.583   &0.566   		&0.628    	&1\\
@@
                  	&0.444     	&0.252     &0.484     	&0.444   &0.463      	&0.614    	&3\\
-  \hline
 	Weighted Avg.   & 0.849    & 0.343      &0.842   &  0.849     &0.842     & 0.753			&\\
-	\hline
+	\bottomrule
 \end{tabular}
@@
 \centering
-\begin{tabular}{|llllllll|}
-	\hline
+\begin{tabular}{llllllll}
+	\toprule
 	&								TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure   &ROC Area  	&Class\\
-	\hline
+	\midrule
 									&0.569     	&0.351     &0.547     	&0.569   &0.558      	&0.65     	&1\\
@@
                  	&0.438     	&0.279     &0.455     	&0.438   &0.447      	&0.603    	&3\\
-  \hline
+
 	Weighted Avg.   &0.489     	&0.284     &0.487     	&0.489   &0.488      	&0.64			&\\
-	\hline
+	\bottomrule
 \end{tabular}

### 2012-11-25 paper.tex
bump
added table for the poor results of decision tree on echocardiogram

@@
 \end{lstlisting}
-Here, the existence of a pericardial effusion (fluid around the heart) is assumed to be a healthy characteristic, but it is not. Accordingly, the precision of the J48 decreases from 95\% to 88\%.
+Here, the decision tree assumes a pericardial effusion (fluid around the heart) to be a healthy characteristic, but it is not.
+Accordingly, the precision of the J48 decreases from 95\% to 88\%.
+Using a higher confidence factor instead of a totally unpruned tree has no effect on the quality or the classification distribution.
+
+\paragraph{Other parameter}
+Neither the laplace correctin to smooth the count of the leaves nor reduced error pruning has any effect on the quality of result.
+Without mdl correction the precision decreases a little to 91.9\%.
+Raising the minimum of instances per leaf has no effect either, which is no suprise since the data set is rather small.
+We couldn't get a better result with the J48 algorithm than to classify all patients to survive at least one year.
+The detailed accuracy for this simple, single-sided classification can be seen in table
+\ref{tab:echo:dec:best}.
+
+\begin{table*}[htb]\centering
+    \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+        \toprule 
+        \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
+               &  TP Rate & FP Rate & Precision & Recall & F-Measure &  Class  \\ \midrule
+               &  0       & 0       & 0         & 0      & 0         &  0      \\     
+               &  1       & 1       & 0.959     & 1      & 0.979     &  1      \\     
+Weighted Avg.  &  0.959   & 0.959   & 0.919     & 0.959  & 0.939     &         \\ \bottomrule
+    \end{tabular*}
+\caption{Decision Tree -- best achievable results} 
+\label{tab:echo:dec:best}
+\end{table*}
+\FloatBarrier
 

### 2012-11-25 paper.tex
deleted all "Detailed Accuracy blabla.. " captions

@@
 
-\todo[inline]{Choose 5 different classifiers, from at least 3 different types
-of learning algorithms
-
-Argue \& justify choice (part of grading...)
-
-Experiment with different parameter settings - And report on it !
-
-Compare results among classifiers and datasets
-
-Evaluate effect of pre-processing (e.g. different strategies
-for missing values)
-}
-
 \section{Preface}
@@
         \toprule 
-        \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
                &  TP Rate & FP Rate & Precision & Recall & F-Measure &  Class  \\ \midrule
@@
         \toprule 
-        \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
               &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class \\ \midrule
@@
         \toprule 
-        \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
                &  TP Rate & FP Rate & Precision & Recall & F-Measure &   Class  \\ \midrule
@@
         \toprule 
-        \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
                &  TP Rate & FP Rate & Precision & Recall & F-Measure &   Class \\ \midrule
@@
         \toprule 
-        \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
                &  TP Rate &  FP Rate &  Precision &  Recall &  F-Measure &  Class \\ \midrule
@@
       \toprule 
-      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
               &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class    \\  \midrule
@@
         \toprule 
-        \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
                & TP Rate & FP Rate & Precision & Recall &  F-Measure &  Class \\ \midrule  
@@
         \toprule 
-        \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
               &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class    \\  \midrule
@@
         \toprule 
-        \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
               &  TP Rate & FP Rate & Precision & Recall &  F-Measure &  Class  \\ \midrule        
@@
       \toprule 
-      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
                &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class \\ \midrule
@@
       \toprule 
-      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
                &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class \\   \midrule
@@
       \toprule 
-      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
                &  TP Rate & FP Rate & Precision & Recall & F-Measure &  Class \\ \midrule 
@@
       \toprule 
-      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
               &   TP Rate & FP Rate & Precision & Recall & F-Measure &  Class \\ \midrule      
@@
       \toprule 
-      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
                &  TP Rate & FP Rate & Precision & Recall & F-Measure &  Class \\ \midrule   
@@
       \toprule 
-      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
                &  TP Rate & FP Rate & Precision & Recall & F-Measure  & Class  \\  \midrule      
@@
       \toprule 
-      \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
                &  TP Rate & FP Rate & Precision & Recall & F-Measure &  Class \\  \midrule     

### 2012-11-25 paper.tex
minor changes -- deleted todo "compare with original data" and some typo

@@
 %The only thing which remained then was to a
-
-\todo[inline]{Compare results with/without preprocessing}
 
@@
 \paragraph{}First of all, with the default settings of the J48 implementation of weka, i.e.
-using pruning with a confidence factor of 0.25 together with MDL correction and subtree raising and at least 2 instances per leaf of the tree, we get a summary of the quality as shown in table \ref{tab:adult:dec:default}:
+using pruning with a confidence factor of 0.25 together with mdl correction and subtree raising and at least 2 instances per leaf of the tree, we get a summary of the quality as shown in table \ref{tab:adult:dec:default}:
 The decision tree that was used for this run has a size of 413 and has 311 leaves.
@@
 Smoothing the counts at the leaves with laplace has no effect.
-Either without MDLcorrection or without Subtree raising during pruning, in both cases the precision falls from 86\% to 85.5\%.
+Either without mdl correction or without Subtree raising during pruning, in both cases the precision falls from 86\% to 85.5\%.
 If we rise the minimum of instances per leaves, e.g.
@@
 Neither laplace correction nor mdl correction has any positive effect on the output of the algorithm on this data set.
-The absence of MDL correction make the outcome a bit worse, it brings the precision down to 0.551\%.
+The absence of mdl correction make the outcome a bit worse, it brings the precision down to 0.551\%.
 But there is one small tweak from which the accuracy of the algorithm can benefit: If binary splits are used on nominal attributes, we get the best result for the J48 on this data set as seen in table \ref{tab:contra:dec:bin}.

### 2012-11-25 paper.tex
added comparison tables

--- 2012-11-25 paper.tex
@@
 \end{lstlisting}
-Here, the decision tree assumes a pericardial effusion (fluid around the heart) to be a healthy characteristic, but it is not.
-Accordingly, the precision of the J48 decreases from 95\% to 88\%.
-Using a higher confidence factor instead of a totally unpruned tree has no effect on the quality or the classification distribution.
-
-\paragraph{Other parameter}
-Neither the laplace correctin to smooth the count of the leaves nor reduced error pruning has any effect on the quality of result.
-Without mdl correction the precision decreases a little to 91.9\%.
-Raising the minimum of instances per leaf has no effect either, which is no suprise since the data set is rather small.
-We couldn't get a better result with the J48 algorithm than to classify all patients to survive at least one year.
-The detailed accuracy for this simple, single-sided classification can be seen in table
-\ref{tab:echo:dec:best}.
-
-\begin{table*}[htb]\centering
-    \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-        \toprule 
-        \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
-               &  TP Rate & FP Rate & Precision & Recall & F-Measure &  Class  \\ \midrule
-               &  0       & 0       & 0         & 0      & 0         &  0      \\     
-               &  1       & 1       & 0.959     & 1      & 0.979     &  1      \\     
-Weighted Avg.  &  0.959   & 0.959   & 0.919     & 0.959  & 0.939     &         \\ \bottomrule
-    \end{tabular*}
-\caption{Decision Tree -- best achievable results} 
-\label{tab:echo:dec:best}
-\end{table*}
-\FloatBarrier
+Here, the existence of a pericardial effusion (fluid around the heart) is assumed to be a healthy characteristic, but it is not. Accordingly, the precision of the J48 decreases from 95\% to 88\%.
 
@@
 \centering
-\begin{tabular}{llllllll}
-	\toprule
-									&TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure   &ROC Area  &Class\\
+\begin{tabular}{lllllll}
+	\toprule
+									&TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure   &Class\\
 	\midrule
-									&0         	&0         &0         	&0       &  0         & 0.5      &0\\
-  								&1         	&1         &0.959     	&1       &0.979       &	0.5      &1\\
-
-	Weighted Avg.		&0.959     	&0.959     &0.919   		&0.959   &0.939     	&0.5	&\\
+									&0         	&0         &0         	&0       &  0              &0\\
+  								&1         	&1         &0.959     	&1       &0.979             &1\\
+
+	Weighted Avg.		&0.959     	&0.959     &0.919   		&0.959   &0.939     		&\\
 	\bottomrule
@@
 \centering
-\begin{tabular}{llllllll}
-	\toprule
-									&TP Rate   	&FP Rate   &Precision &Recall  &F-Measure   &ROC Area  &Class\\
+\begin{tabular}{lllllll}
+	\toprule
+									&TP Rate   	&FP Rate   &Precision &Recall  &F-Measure   &Class\\
 	\midrule
-									&0         	&0         &0         &0       &  0         & 0.5      &0\\
-  								&1         	&1         &0.959     &1       &0.979       &	0.5      &1\\
-
-	Weighted Avg.		&0.959     	&0.959     &0.919   	&0.959   &0.939     	&0.5				&\\
+									&0         	&0         &0         &0       &  0         &0\\
+  								&1         	&1         &0.959     &1       &0.979       &1\\
+
+	Weighted Avg.		&0.959     	&0.959     &0.919   	&0.959   &0.939     					&\\
 	\bottomrule
@@
 \subsection{Comparison}
+
+\begin{table}[h]
+\centering
+\begin{tabular}{llllllll}
+	\toprule
+									&TP Rate   	&FP Rate   &Precision &Recall  &F-Measure   &  Correctly classified\\
+	\midrule
+	Decision Tree		\\
+  Random Forest		\\
+  Naive Bayes			&0.897  		& 0.244   	& 0.958   & 0.897  &0.92      	& \\
+  SVM							&0.959     	&0.959     	&0.919   	&0.959   &0.939     	&	0.959\\
+  kNN							&0.959     	&0.959     	&0.919   	&0.959   &0.939     	&	0.959\\
+	\bottomrule
+\end{tabular}
+\caption{Comparison of the different algorithms for the echocardiogram dataset}
+\end{table}
 
@@
                &  0.577   & 0.052   & 0.779     & 0.577  & 0.663     & >50K  \\   
-               &  0.948   & 0.423   & 0.876     & 0.948  & 0.911     & <=50K \\   
+               &  0.948   & 0.423   & 0.876     & 0.948  & 0.911     & $\leq$50K \\   
 Weighted Avg.  &  0.859   & 0.333   & 0.853     & 0.859  & 0.851     &       \\   \bottomrule
@@
 	\toprule
-									&TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure &ROC Area  	&Class\\
+									&TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure  	&Class\\
 	\midrule
-									&0.568     	&0.062     &0.744     	&0.568   &0.644    	&0.753    	&	>50K\\
-  								&0.938     	&0.432     &0.873     	&0.938   &0.904     &0.753    	&$\leq$50K\\
-
-	Weighted Avg.   & 0.849    & 0.343     &0.842   		&0.849   &0.842     & 0.753			&\\
+									&0.568     	&0.062     &0.744     	&0.568   &0.644    	    	&	>50K\\
+  								&0.938     	&0.432     &0.873     	&0.938   &0.904         	&$\leq$50K\\
+
+	Weighted Avg.   & 0.849    & 0.343     &0.842   		&0.849   &0.842     			&\\
 	\bottomrule
@@
 \centering
-\begin{tabular}{llllllll}
-	\toprule
-									&TP Rate   &FP Rate   &Precision   &Recall  &F-Measure &ROC Area  	&Class\\
+\begin{tabular}{lllllll}
+	\toprule
+									&TP Rate   &FP Rate   &Precision   &Recall  &F-Measure &Class\\
 	\midrule
-									&0.593     &0.088      &0.68      &0.593     &0.633    &0.884    		&>50K\\
-									&0.912     &0.407      &0.876     &0.912     &0.893    &0.884    		&$\leq$50K\\
-
-	Weighted Avg.   &0.835     &0.331      &0.829     &0.835     &0.831    &0.884 \\
+									&0.593     &0.088      &0.68      &0.593     &0.633    &>50K\\
+									&0.912     &0.407      &0.876     &0.912     &0.893    &$\leq$50K\\
+
+	Weighted Avg.   &0.835     &0.331      &0.829     &0.835     &0.831     \\
 	\bottomrule
@@
 
-\todo{todo marker}
+\begin{table}[h]
+\centering
+\begin{tabular}{llllllll}
+	\toprule
+									&TP Rate   	&FP Rate   &Precision &Recall  &F-Measure   &  Correctly classified\\
+	\midrule
+	Decision Tree		&  0.846   	& 0.285   	& 0.843   & 0.846  &  0.844 		& FEHLT\\
+  Random Forest		&  0.859   	& 0.333   	& 0.853   & 0.859  & 0.851     	& FEHLT\\
+  Naive Bayes		 	&  0.83    	& 0.186   	& 0.854   & 0.83   & 0.837     	& FEHLT\\
+  SVM							& 0.849    	& 0.343     &0.842   	&0.849   &0.842     	& 0.849\\
+  kNN							&0.835     	&0.331      &0.829    &0.835   &0.831				& 0.835\\
+	\bottomrule
+\end{tabular}
+\caption{Comparison of the different algorithms for the adult dataset}
+\end{table}
 
@@
 \centering
-\begin{tabular}{llllllll}
-	\toprule
-	&								TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure   &ROC Area  	&Class\\ 
+\begin{tabular}{lllllll}
+	\toprule
+	&								TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure     	&Class\\ 
 	\midrule
-									&0.583     	&0.355     &0.55      	&0.583   &0.566   		&0.628    	&1\\
-                 	&0.393     	&0.181     &0.389     	&0.393   &0.391      	&0.648    	&2\\
-                 	&0.444     	&0.252     &0.484     	&0.444   &0.463      	&0.614    	&3\\
-	Weighted Avg.   & 0.849    & 0.343      &0.842   &  0.849     &0.842     & 0.753			&\\
+									&0.583     	&0.355     &0.55      	&0.583   &0.566   		    	&1\\
+                 	&0.393     	&0.181     &0.389     	&0.393   &0.391      	    	&2\\
+                 	&0.444     	&0.252     &0.484     	&0.444   &0.463      	    	&3\\
+	Weighted Avg.   & 0.849    & 0.343      &0.842   &  0.849     &0.842     			&\\
 	\bottomrule
@@
 \centering
-\begin{tabular}{llllllll}
-	\toprule
-	&								TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure   &ROC Area  	&Class\\
+\begin{tabular}{lllllll}
+	\toprule
+	&								TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure     	&Class\\
 	\midrule
-									&0.569     	&0.351     &0.547     	&0.569   &0.558      	&0.65     	&1\\
-                 	&0.414     	&0.166     &0.422     	&0.414   &0.418      	&0.677    	&2\\
-                 	&0.438     	&0.279     &0.455     	&0.438   &0.447      	&0.603    	&3\\
-
-	Weighted Avg.   &0.489     	&0.284     &0.487     	&0.489   &0.488      	&0.64			&\\
+									&0.569     	&0.351     &0.547     	&0.569   &0.558      	     	&1\\
+                 	&0.414     	&0.166     &0.422     	&0.414   &0.418      	    	&2\\
+                 	&0.438     	&0.279     &0.455     	&0.438   &0.447      	    	&3\\
+
+	Weighted Avg.   &0.489     	&0.284     &0.487     	&0.489   &0.488      				&\\
 	\bottomrule
@@
 
-\todo{todo marker}
-
+\subsection{Comparison}
+
+\begin{table}[h]
+\centering
+\begin{tabular}{llllllll}
+	\toprule
+									&TP Rate   	&FP Rate   &Precision &Recall  &F-Measure   &  Correctly classified\\
+	\midrule
+	Decision Tree 	&  0.558   & 0.243   & 0.557     	& 0.558  & 0.557     & FEHLT\\
+  Random Forest		&  0.556   & 0.241   & 0.562     	& 0.556  & 0.556     & FEHLT\\
+  Naive Bayes		 	&  0.508   & 0.243   & 0.528     	& 0.508  & 0.513     & FEHLT\\
+  SVM							& 0.849    & 0.343   &0.842   		& 0.849  & 0.842 		 & 0.492\\
+  kNN							&0.489     &0.284    &0.487     	& 0.489  & 0.488 		 & 0.489\\
+	\bottomrule
+\end{tabular}
+\caption{Comparison of the different algorithms for the contraceptive method choice dataset}
+\end{table}
 \bibliography{references}

### 2012-11-25 paper.tex
toc

--- 2012-11-25 paper.tex
@@
 
-\todo[inline]{Choose 5 different classifiers, from at least 3 different types
-of learning algorithms
-
-Argue \& justify choice (part of grading...)
-
-Experiment with different parameter settings - And report on it !
-
-Compare results among classifiers and datasets
-
-Evaluate effect of pre-processing (e.g. different strategies
-for missing values)
-}
+\tableofcontents
 
@@
 \centering
-\begin{tabular}{lllllll}
-	\toprule
-									&TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure   &Class\\
+\begin{tabular}{llllllll}
+	\toprule
+									&TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure   &ROC Area  &Class\\
 	\midrule
-									&0         	&0         &0         	&0       &  0              &0\\
-  								&1         	&1         &0.959     	&1       &0.979             &1\\
-
-	Weighted Avg.		&0.959     	&0.959     &0.919   		&0.959   &0.939     		&\\
+									&0         	&0         &0         	&0       &  0         & 0.5      &0\\
+  								&1         	&1         &0.959     	&1       &0.979       &	0.5      &1\\
+
+	Weighted Avg.		&0.959     	&0.959     &0.919   		&0.959   &0.939     	&0.5	&\\
 	\bottomrule
@@
 \centering
-\begin{tabular}{lllllll}
-	\toprule
-									&TP Rate   	&FP Rate   &Precision &Recall  &F-Measure   &Class\\
+\begin{tabular}{llllllll}
+	\toprule
+									&TP Rate   	&FP Rate   &Precision &Recall  &F-Measure   &ROC Area  &Class\\
 	\midrule
-									&0         	&0         &0         &0       &  0         &0\\
-  								&1         	&1         &0.959     &1       &0.979       &1\\
-
-	Weighted Avg.		&0.959     	&0.959     &0.919   	&0.959   &0.939     					&\\
+									&0         	&0         &0         &0       &  0         & 0.5      &0\\
+  								&1         	&1         &0.959     &1       &0.979       &	0.5      &1\\
+
+	Weighted Avg.		&0.959     	&0.959     &0.919   	&0.959   &0.939     	&0.5				&\\
 	\bottomrule
@@
 \subsection{Comparison}
-
-\begin{table}[h]
-\centering
-\begin{tabular}{llllllll}
-	\toprule
-									&TP Rate   	&FP Rate   &Precision &Recall  &F-Measure   &  Correctly classified\\
-	\midrule
-	Decision Tree		\\
-  Random Forest		\\
-  Naive Bayes			&0.897  		& 0.244   	& 0.958   & 0.897  &0.92      	& \\
-  SVM							&0.959     	&0.959     	&0.919   	&0.959   &0.939     	&	0.959\\
-  kNN							&0.959     	&0.959     	&0.919   	&0.959   &0.939     	&	0.959\\
-	\bottomrule
-\end{tabular}
-\caption{Comparison of the different algorithms for the echocardiogram dataset}
-\end{table}
 
@@
                &  0.577   & 0.052   & 0.779     & 0.577  & 0.663     & >50K  \\   
-               &  0.948   & 0.423   & 0.876     & 0.948  & 0.911     & $\leq$50K \\   
+               &  0.948   & 0.423   & 0.876     & 0.948  & 0.911     & <=50K \\   
 Weighted Avg.  &  0.859   & 0.333   & 0.853     & 0.859  & 0.851     &       \\   \bottomrule
@@
 	\toprule
-									&TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure  	&Class\\
+									&TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure &ROC Area  	&Class\\
 	\midrule
-									&0.568     	&0.062     &0.744     	&0.568   &0.644    	    	&	>50K\\
-  								&0.938     	&0.432     &0.873     	&0.938   &0.904         	&$\leq$50K\\
-
-	Weighted Avg.   & 0.849    & 0.343     &0.842   		&0.849   &0.842     			&\\
+									&0.568     	&0.062     &0.744     	&0.568   &0.644    	&0.753    	&	>50K\\
+  								&0.938     	&0.432     &0.873     	&0.938   &0.904     &0.753    	&$\leq$50K\\
+
+	Weighted Avg.   & 0.849    & 0.343     &0.842   		&0.849   &0.842     & 0.753			&\\
 	\bottomrule
@@
 \centering
-\begin{tabular}{lllllll}
-	\toprule
-									&TP Rate   &FP Rate   &Precision   &Recall  &F-Measure &Class\\
+\begin{tabular}{llllllll}
+	\toprule
+									&TP Rate   &FP Rate   &Precision   &Recall  &F-Measure &ROC Area  	&Class\\
 	\midrule
-									&0.593     &0.088      &0.68      &0.593     &0.633    &>50K\\
-									&0.912     &0.407      &0.876     &0.912     &0.893    &$\leq$50K\\
-
-	Weighted Avg.   &0.835     &0.331      &0.829     &0.835     &0.831     \\
+									&0.593     &0.088      &0.68      &0.593     &0.633    &0.884    		&>50K\\
+									&0.912     &0.407      &0.876     &0.912     &0.893    &0.884    		&$\leq$50K\\
+
+	Weighted Avg.   &0.835     &0.331      &0.829     &0.835     &0.831    &0.884 \\
 	\bottomrule
@@
 
-\begin{table}[h]
-\centering
-\begin{tabular}{llllllll}
-	\toprule
-									&TP Rate   	&FP Rate   &Precision &Recall  &F-Measure   &  Correctly classified\\
-	\midrule
-	Decision Tree		&  0.846   	& 0.285   	& 0.843   & 0.846  &  0.844 		& FEHLT\\
-  Random Forest		&  0.859   	& 0.333   	& 0.853   & 0.859  & 0.851     	& FEHLT\\
-  Naive Bayes		 	&  0.83    	& 0.186   	& 0.854   & 0.83   & 0.837     	& FEHLT\\
-  SVM							& 0.849    	& 0.343     &0.842   	&0.849   &0.842     	& 0.849\\
-  kNN							&0.835     	&0.331      &0.829    &0.835   &0.831				& 0.835\\
-	\bottomrule
-\end{tabular}
-\caption{Comparison of the different algorithms for the adult dataset}
-\end{table}
+\todo{todo marker}
 
@@
 \centering
-\begin{tabular}{lllllll}
-	\toprule
-	&								TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure     	&Class\\ 
+\begin{tabular}{llllllll}
+	\toprule
+	&								TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure   &ROC Area  	&Class\\ 
 	\midrule
-									&0.583     	&0.355     &0.55      	&0.583   &0.566   		    	&1\\
-                 	&0.393     	&0.181     &0.389     	&0.393   &0.391      	    	&2\\
-                 	&0.444     	&0.252     &0.484     	&0.444   &0.463      	    	&3\\
-	Weighted Avg.   & 0.849    & 0.343      &0.842   &  0.849     &0.842     			&\\
+									&0.583     	&0.355     &0.55      	&0.583   &0.566   		&0.628    	&1\\
+                 	&0.393     	&0.181     &0.389     	&0.393   &0.391      	&0.648    	&2\\
+                 	&0.444     	&0.252     &0.484     	&0.444   &0.463      	&0.614    	&3\\
+	Weighted Avg.   & 0.849    & 0.343      &0.842   &  0.849     &0.842     & 0.753			&\\
 	\bottomrule
@@
 \centering
-\begin{tabular}{lllllll}
-	\toprule
-	&								TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure     	&Class\\
+\begin{tabular}{llllllll}
+	\toprule
+	&								TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure   &ROC Area  	&Class\\
 	\midrule
-									&0.569     	&0.351     &0.547     	&0.569   &0.558      	     	&1\\
-                 	&0.414     	&0.166     &0.422     	&0.414   &0.418      	    	&2\\
-                 	&0.438     	&0.279     &0.455     	&0.438   &0.447      	    	&3\\
-
-	Weighted Avg.   &0.489     	&0.284     &0.487     	&0.489   &0.488      				&\\
+									&0.569     	&0.351     &0.547     	&0.569   &0.558      	&0.65     	&1\\
+                 	&0.414     	&0.166     &0.422     	&0.414   &0.418      	&0.677    	&2\\
+                 	&0.438     	&0.279     &0.455     	&0.438   &0.447      	&0.603    	&3\\
+
+	Weighted Avg.   &0.489     	&0.284     &0.487     	&0.489   &0.488      	&0.64			&\\
 	\bottomrule
@@
 
-\subsection{Comparison}
-
-\begin{table}[h]
-\centering
-\begin{tabular}{llllllll}
-	\toprule
-									&TP Rate   	&FP Rate   &Precision &Recall  &F-Measure   &  Correctly classified\\
-	\midrule
-	Decision Tree 	&  0.558   & 0.243   & 0.557     	& 0.558  & 0.557     & FEHLT\\
-  Random Forest		&  0.556   & 0.241   & 0.562     	& 0.556  & 0.556     & FEHLT\\
-  Naive Bayes		 	&  0.508   & 0.243   & 0.528     	& 0.508  & 0.513     & FEHLT\\
-  SVM							& 0.849    & 0.343   &0.842   		& 0.849  & 0.842 		 & 0.492\\
-  kNN							&0.489     &0.284    &0.487     	& 0.489  & 0.488 		 & 0.489\\
-	\bottomrule
-\end{tabular}
-\caption{Comparison of the different algorithms for the contraceptive method choice dataset}
-\end{table}
+\todo{todo marker}
+
 \bibliography{references}

### 2012-11-25 paper.tex
added data in the tables

--- 2012-11-25 paper.tex
@@
 
+The table below shos a comparison of the different algorithms on the adult dataset.
+For an algorithm we present here only the best obtained result concerning the TP rate.
+Values are always the weighted average.
+
 \begin{table}[h]
@@
 	\midrule
-	Decision Tree		&  0.846   	& 0.285   	& 0.843   & 0.846  &  0.844 		& FEHLT\\
+	Decision Tree		& 0.86    	& 0.307   	& 0.854   & 0.86   & 0.855     & 		& FEHLT\\
   Random Forest		& 0.859   	& 0.333   	& 0.853   & 0.859  & 0.851     	& FEHLT\\
@@
 
-\subsection{Comparison}
+The table below shos a comparison of the different algorithms on the adult dataset.
+For an algorithm we present here only the best obtained result concerning the TP rate.
+Values are always the weighted average.
 

### 2012-11-25 paper.tex
asdf
added some percentage numbers of correctly classified instances

--- 2012-11-25 paper.tex
@@
 But if we configure the random forests with 6 features we can spot our first true positive in the confusion matrix (table \ref{tab:echo:rand:matrix3}).
-This gives us the best achievable accuracy that we can reach with decision based algorithms, shown in table \ref{tab:echo:rand:6f}.
+This gives us the best achievable accuracy that we can reach with decision based algorithms, shown in table \ref{tab:echo:rand:6f}. At this point, 96.8\% instances are correctly classified.
 
@@
 
-\begin{table}[h]
-\centering
-\begin{tabular}{llllllll}
-	\toprule
-									&TP Rate   	&FP Rate   &Precision &Recall  &F-Measure   &  Correctly classified\\
-	\midrule
-	Decision Tree		\\
-  Random Forest		\\
-  Naive Bayes			&0.897  		& 0.244   	& 0.958   & 0.897  &0.92      	& \\
-  SVM							&0.959     	&0.959     	&0.919   	&0.959   &0.939     	&	0.959\\
-  kNN							&0.959     	&0.959     	&0.919   	&0.959   &0.939     	&	0.959\\
-	\bottomrule
-\end{tabular}
-\caption{Comparison of the different algorithms for the echocardiogram dataset}
-\end{table}
-
 \todo{todo marker}
@@
 Smoothing the counts at the leaves with laplace has no effect.
-Either without mdl correction or without Subtree raising during pruning, in both cases the precision falls from 86\% to 85.5\%.
+Either without mdl correction or without Subtree raising during pruning, in both cases the precision falls from 86\% to
+85.5\%.
 If we rise the minimum of instances per leaves, e.g.
 to 20 instances and 50 instances, the precision falls from 86\% to 85\% and 84\% respectively.
+So we get the best results simply with the default configuration and classify 86.0\% instances correctly.
 
@@
 The default number of features for this data set with 13 attributes is $ log_2(13) + 1 \approx 5 $.
-Choosing 6 attributes brinchoose gs us a precision of 84,9\% and with 10 randomly chosen features the precision falls back to 84.5\%.
+Choosing 6 attributes gives us a precision of 84,9\% and with 10 randomly chosen features the precision falls back to 84.5\%.
 We can tease the most out of the random forest if we decrease the number of features a little.
@@
 But there is one small tweak from which the accuracy of the algorithm can benefit: If binary splits are used on nominal attributes, we get the best result for the J48 on this data set as seen in table \ref{tab:contra:dec:bin}.
+Eventually we have 95.88\% correctly
+classified instances with the J48 algorithm.
 

### 2012-11-25 paper.tex
comparisons

--- 2012-11-25 paper.tex
@@
 
+\tableofcontents
+
 \section{Preface}
@@
   Random Forest		\\
-  Naive Bayes			&0.897  		& 0.244   	& 0.958   & 0.897  &0.92      	& \\
+  Naive Bayes			&   0.897  & 0.244   & 0.958     & 0.897  & 0.92     \\
   SVM							&0.959     	&0.959     	&0.919   	&0.959   &0.939     	&	0.959\\
@@
 \end{table}
+
+%% 
+%%
+\begin{table*}[htb]\centering
+  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+      \toprule 
+              &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class    \\  \midrule
+              &   0.75   & 0.097   & 0.25      & 0.75   & 0.375     & 0 \\
+              &   0.903  & 0.25    & 0.988     & 0.903  & 0.944     & 1 \\
+Weighted Avg. &   0.897  & 0.244   & 0.958     & 0.897  & 0.92      &   \\ \bottomrule
+  \end{tabular*}
+\caption{Echocardiogram -- Naive Bayes with SMOTE filter} 
+\label{tab:echo:bayes:2d}
+\end{table*}
 
@@
   Random Forest		& 0.859   	& 0.333   	& 0.853   & 0.859  & 0.851     	& FEHLT\\
-  Naive Bayes		 	& 0.83    	& 0.186   	& 0.854   & 0.83   & 0.837     	& FEHLT\\
+  Naive Bayes		 	&  0.83    & 0.186   & 0.854     & 0.83   & 0.837     \\
   SVM							& 0.849    	& 0.343     &0.842   	&0.849   &0.842     	& 0.849\\
@@
   Random Forest		&  0.556   & 0.241   & 0.562     	& 0.556  & 0.556     & FEHLT\\
-  Naive Bayes		 	&  0.508   & 0.243   & 0.528     	& 0.508  & 0.513     & FEHLT\\
+  Naive Bayes		 	&  0.508   & 0.243   & 0.528     & 0.508  & 0.513 \\
   SVM							& 0.849    & 0.343   &0.842   		& 0.849  & 0.842 		 & 0.492\\

### 2012-11-25 paper.tex
asdf
removed column 'correctly classified'

@@
 	\toprule
-									&TP Rate   	&FP Rate   &Precision &Recall  &F-Measure   &  Correctly classified\\
+									&TP Rate   	&FP Rate   &Precision &Recall  &F-Measure \\
 	\midrule
@@
   Naive Bayes			&   0.897  & 0.244   & 0.958     & 0.897  & 0.92     \\
-  SVM							&0.959     	&0.959     	&0.919   	&0.959   &0.939     	&	0.959\\
-  kNN							&0.959     	&0.959     	&0.919   	&0.959   &0.939     	&	0.959\\
+  SVM							&0.959     	&0.959     	&0.919   	&0.959   &0.939    \\
+  kNN							&0.959     	&0.959     	&0.919   	&0.959   &0.939    \\
 	\bottomrule
@@
 \centering
-\begin{tabular}{llllllll}
-	\toprule
-									&TP Rate   	&FP Rate   &Precision &Recall  &F-Measure   &  Correctly classified\\
+\begin{tabular}{lllllll}
+	\toprule
+									&TP Rate   	&FP Rate   &Precision &Recall  &F-Measure  \\
 	\midrule
-	Decision Tree		& 0.86    	& 0.307   	& 0.854   & 0.86   & 0.855     & 		& FEHLT\\
-  Random Forest		& 0.859   	& 0.333   	& 0.853   & 0.859  & 0.851     	& FEHLT\\
+	Decision Tree		& 0.86    	& 0.307   	& 0.854   & 0.86   & 0.855     \\
+  Random Forest		& 0.859   	& 0.333   	& 0.853   & 0.859  & 0.851    \\
   Naive Bayes		 	&  0.83    & 0.186   & 0.854     & 0.83   & 0.837     \\
-  SVM							& 0.849    	& 0.343     &0.842   	&0.849   &0.842     	& 0.849\\
-  kNN							&0.835     	&0.331      &0.829    &0.835   &0.831				& 0.835\\
+  SVM							& 0.849    	& 0.343     &0.842   	&0.849   &0.842     \\
+  kNN							&0.835     	&0.331      &0.829    &0.835   &0.831			\\
 	\bottomrule
@@
                  	&0.444     	&0.252     &0.484     	&0.444   &0.463      	    	&3\\
-	Weighted Avg.   & 0.849    & 0.343      &0.842   &  0.849     &0.842     			&\\
+	Weighted Avg.   &0.492     	&0.28      &0.491     	&0.492   &0.491     			&\\
 	\bottomrule
@@
 \centering
-\begin{tabular}{llllllll}
-	\toprule
-									&TP Rate   	&FP Rate   &Precision &Recall  &F-Measure   &  Correctly classified\\
+\begin{tabular}{lllllll}
+	\toprule
+									&TP Rate   	&FP Rate   &Precision &Recall  &F-Measure   \\
 	\midrule
-	Decision Tree 	&  0.558   & 0.243   & 0.557     	& 0.558  & 0.557     & FEHLT\\
-  Random Forest		&  0.556   & 0.241   & 0.562     	& 0.556  & 0.556     & FEHLT\\
+	Decision Tree 	&  0.558   & 0.243   & 0.557     	& 0.558  & 0.557     \\
+  Random Forest		&  0.556   & 0.241   & 0.562     	& 0.556  & 0.556     \\
   Naive Bayes		 	&  0.508   & 0.243   & 0.528     	& 0.508  & 0.513 \\
-  SVM							& 0.849    & 0.343   &0.842   		& 0.849  & 0.842 		 & 0.492\\
-  kNN							&0.489     &0.284    &0.487     	& 0.489  & 0.488 		 & 0.489\\
+  SVM							& 0.492    &0.280     &0.491     	&0.492   & 0.491 		 \\
+  kNN							&0.489     &0.284    &0.487     	& 0.489  & 0.488 		 \\
 	\bottomrule

### 2012-11-25 paper.tex
bla

@@
 Smoothing the counts at the leaves with laplace has no effect.
-Either without mdl correction or without Subtree raising during pruning, in both cases the precision falls from 86\% to 85.5\%.
+Either without mdl correction or without Subtree raising during pruning, in both cases the precision falls from 86\% to
+85.5\%.
 If we rise the minimum of instances per leaves, e.g.
 to 20 instances and 50 instances, the precision falls from 86\% to 85\% and 84\% respectively.
-
-\subsection{Random Forest}
+So we get the best results simply with the default configuration and classify 86.0\% instances correctly.
+
+\subsubsection{Random Forest}
 
@@
 
-\subsection{Random Forest}
+\subsubsection{Random Forest}
 Random Forests perform even worse than the J48 with default settings. We can see in table \ref{tab:contra:rand:default} that only the half of all classified instances are classified correctly.
@@
 
-\subsection{Naive Bayes}
-
-\paragraph{}The input data set does perfectly fit for Naive Bayes.
-But the classifier is not able to correctly predict the instances.
-As shown in tables \ref{tab:contra:bayes:1d} and \ref{tab:contra:bayes:1c} it gains a 50:50 chance of a TP prediction.
-
-\begin{table*}[htb]\centering
-  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-      \toprule 
-               &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class \\   \midrule
-               &  0.52    & 0.223   & 0.635     & 0.52   & 0.572     & 1  \\  
-               &  0.526   & 0.245   & 0.385     & 0.526  & 0.445     & 2  \\
-               &  0.481   & 0.268   & 0.488     & 0.481  & 0.485     & 3  \\
-Weighted Avg.  &  0.508   & 0.243   & 0.528     & 0.508  & 0.513     &    \\   \bottomrule
-    \end{tabular*}
-\caption{Naive Bayes} 
-\label{tab:contra:bayes:1d}
-\end{table*}
-
-\begin{table}[h]
-\centering
-\begin{tabular}{|l|lll|}
-\hline
-              &   a &   b &   c  \\
-\hline
- a = 1        & 327 & 141 & 161  \\
- b = 2        &  61 & 175 &  97  \\
- c = 3        & 127 & 138 & 246  \\
-\hline
-\end{tabular}
-\caption{Confusion matrix for Naive Bayes}
-\label{tab:contra:bayes:1c}
-\end{table}
-
 \subsection{Support Vector Machine}
@@
 \caption{Confusion matrix for kNN}
+\end{table}
+
+\subsection{Naive Bayes}
+
+\paragraph{}The input data set does perfectly fit for Naive Bayes.
+But the classifier is not able to correctly predict the instances.
+As shown in tables \ref{tab:contra:bayes:1d} and \ref{tab:contra:bayes:1c} it gains a 50:50 chance of a TP prediction.
+
+\begin{table*}[htb]\centering
+  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
+      \toprule 
+               &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class \\   \midrule
+               &  0.52    & 0.223   & 0.635     & 0.52   & 0.572     & 1  \\  
+               &  0.526   & 0.245   & 0.385     & 0.526  & 0.445     & 2  \\
+               &  0.481   & 0.268   & 0.488     & 0.481  & 0.485     & 3  \\
+Weighted Avg.  &  0.508   & 0.243   & 0.528     & 0.508  & 0.513     &    \\   \bottomrule
+    \end{tabular*}
+\caption{Naive Bayes} 
+\label{tab:contra:bayes:1d}
+\end{table*}
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|lll|}
+\hline
+              &   a &   b &   c  \\
+\hline
+ a = 1        & 327 & 141 & 161  \\
+ b = 2        &  61 & 175 &  97  \\
+ c = 3        & 127 & 138 & 246  \\
+\hline
+\end{tabular}
+\caption{Confusion matrix for Naive Bayes}
+\label{tab:contra:bayes:1c}
 \end{table}

### 2012-11-25 paper.tex
asdf
bla

--- 2012-11-25 paper.tex
@@
 
-\paragraph{}It should be clear that the variables ``survival'' and ``still-alive'' should not be used for classification since they constitute the final prediction attribute.
+%\paragraph{}It should be clear that the variables ``survival'' and ``still-alive'' should not be used for classification since they constitute the final prediction attribute.
 Listing \ref{lst:silly-tree} shows what happens if a decision tree can take these attributes into account.
@@
 Pretty accurate, right?
-The only remarkable thing is, that the J48 implementation under default settings seems to enforce a certain depth of the tree, even if the splitting on the same attribute is redundant since $ x > 11 \rightarrow x > 12 $.
-
-\begin{lstlisting}[caption={Decision tree on 100\% dependant attributes}, label=lst:silly-tree, float=htb]
-J48 pruned tree
-------------------
-    survival <= 12
-    |   survival <= 11: 0 (4.0)
-    |   survival > 11: 1 (4.0)
-    survival > 12: 1 (87.0)
-\end{lstlisting}
+%The only remarkable thing is, that the J48 implementation under default settings seems to enforce a certain depth of the tree, even if the splitting on the same attribute is redundant since $ x > 11 \rightarrow x > 12 $.
+%
+%\begin{lstlisting}[caption={Decision tree on 100\% dependant attributes}, label=lst:silly-tree, float=htb]
+%J48 pruned tree
+%------------------
+%    survival <= 12
+%    |   survival <= 11: 0 (4.0)
+%    |   survival > 11: 1 (4.0)
+%    survival > 12: 1 (87.0)
+%\end{lstlisting}
 
@@
 
-\subsection{Random Forest}
+\subsubsection{Random Forest}
 

### 2012-11-25 paper.tex
added comparison lines

--- 2012-11-25 paper.tex
@@
 	\midrule
-	Decision Tree		\\
-  Random Forest		\\
+	Decision Tree &  0.959   & 0.959   & 0.919     & 0.959  & 0.939     &    0.968     \\ 
+  Random Forest &		 0.948   &  0.72    &  0.942     &  0.948  &  0.945             \\ 
   Naive Bayes			&   0.897  & 0.244   & 0.958     & 0.897  & 0.92     \\

### 2012-11-25 paper.tex
hopefully all comparison tables filled?

--- 2012-11-25 paper.tex
@@
 	\midrule
-	Decision Tree		\\
-  Random Forest		\\
+	Decision Tree &  0.959   & 0.959   & 0.919     & 0.959  & 0.939           \\ 
+  Random Forest &		 0.948   &  0.72    &  0.942     &  0.948  &  0.945             \\ 
   Naive Bayes			&   0.897  & 0.244   & 0.958     & 0.897  & 0.92     \\
@@
 \end{table}
-
-%% 
-%%
-\begin{table*}[htb]\centering
-  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-      \toprule 
-              &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class    \\  \midrule
-              &   0.75   & 0.097   & 0.25      & 0.75   & 0.375     & 0 \\
-              &   0.903  & 0.25    & 0.988     & 0.903  & 0.944     & 1 \\
-Weighted Avg. &   0.897  & 0.244   & 0.958     & 0.897  & 0.92      &   \\ \bottomrule
-  \end{tabular*}
-\caption{Echocardiogram -- Naive Bayes with SMOTE filter} 
-\label{tab:echo:bayes:2d}
-\end{table*}
-
-\todo{todo marker}
 

### 2012-11-25 paper.tex
bla

@@
 But if we configure the random forests with 6 features we can spot our first true positive in the confusion matrix (table \ref{tab:echo:rand:matrix3}).
-This gives us the best achievable accuracy that we can reach with decision based algorithms, shown in table \ref{tab:echo:rand:6f}. At this point, 96.8\% instances are correctly classified.
+This gives us the best achievable accuracy that we can reach with decision based algorithms, shown in table \ref{tab:echo:rand:6f}.
 
@@
 \end{tabular}
-\caption{Confusion matrix for kNN}
+\caption{Confusion matrix for Naive Bayes}
 \label{tab:echo:bayes:1c}
@@
 	\midrule
-	Decision Tree &  0.959   & 0.959   & 0.919     & 0.959  & 0.939           \\ 
-  Random Forest &		 0.948   &  0.72    &  0.942     &  0.948  &  0.945             \\ 
+	Decision Tree		\\
+  Random Forest		\\
   Naive Bayes			&   0.897  & 0.244   & 0.958     & 0.897  & 0.92     \\
@@
 Smoothing the counts at the leaves with laplace has no effect.
-Either without mdl correction or without Subtree raising during pruning, in both cases the precision falls from 86\% to
-85.5\%.
+Either without mdl correction or without Subtree raising during pruning, in both cases the precision falls from 86\% to 85.5\%.
 If we rise the minimum of instances per leaves, e.g.
 to 20 instances and 50 instances, the precision falls from 86\% to 85\% and 84\% respectively.
-So we get the best results simply with the default configuration and classify 86.0\% instances correctly.
 
@@
 The default number of features for this data set with 13 attributes is $ log_2(13) + 1 \approx 5 $.
-Choosing 6 attributes gives us a precision of 84,9\% and with 10 randomly chosen features the precision falls back to 84.5\%.
+Choosing 6 attributes brinchoose gs us a precision of 84,9\% and with 10 randomly chosen features the precision falls back to 84.5\%.
 We can tease the most out of the random forest if we decrease the number of features a little.
@@
 But there is one small tweak from which the accuracy of the algorithm can benefit: If binary splits are used on nominal attributes, we get the best result for the J48 on this data set as seen in table \ref{tab:contra:dec:bin}.
-Eventually we have 95.88\% correctly
-classified instances with the J48 algorithm.
 

### 2012-11-25 paper.tex
asdf

@@
 
-\begin{table*}[htb]\centering
-    \begin{tabular}{@{}ll|ll@{}} 
+\begin{table}[htb]\centering
+    \begin{tabular}{|ll|ll|} 
 \multicolumn{2}{c}{\phantom{bla}}  & \multicolumn{2}{c}{Classified as} \\  
@@
 \end{tabular}
-\caption{Confusion matrix for Naive Bayes}
+\caption{Confusion matrix for kNN}
 \label{tab:echo:bayes:1c}

### 2012-11-25 paper.tex
tables floatbarrier

--- 2012-11-25 paper.tex
@@
 \end{table}
-
+\FloatBarrier
 \vspace{6pt}
@@
 \end{table}
+\FloatBarrier
 
@@
 \end{table}
-
+\FloatBarrier
 \paragraph{}Also the detailed accuracy table \ref{tab:echo:bayes:2d} shows that this approach gained a high recall with a low precision for alive-at-1 = 0 and both high precision and recall for alive-at-1 = 1.
@@
 \end{table}
-
+\FloatBarrier
 \paragraph{}By using different discretizers and also do not seperately discretize in the preprocessing section of weka but instead let the NaiveBayers classifier use it's own discretizers namely KernelEstimator or SupervisedDiscretizer the outcome of precision and recall differ a bit by 0.05 to 0.1 higher and lower.
@@
 \end{table}
-\bibliography{references}
+\FloatBarrier\bibliography{references}
 

### 2012-11-27 exercise2/paper.tex
added folder for different exercices 1 and 2

--- 2012-11-25 paper.tex
@@
 \horrule{0.5pt} \\[0.4cm] % Thin top horizontal rule
-\huge Experiments in Machine Learning \\ % The assignment title
+\huge Experiments in Machine Learning 2 \\ % The assignment title
 \horrule{2pt} \\[0.5cm] % Thick bottom horizontal rule
@@
 
-\paragraph{}For the first exercise 3 data sets were chosen on which five classification algorithms were tested to practice basic machine learning tasks and see how these algorithms behave with their various settings.
-For the elaboration of this paper the Machine Learning Software Weka was used.
-It offered all necessary algorithms.
+\paragraph{}This is a template.
 
-\paragraph{}As data sets we chose as follows:
-
-\begin{itemize}
-\item Echo Cardiogram, a data set about heart attacks and if a patient survived one year after the attack.
-\item Adult / Census Income, a data set about people and their basic information and if they earn more than 50k a year.
-\item Contraceptive Method Choice, a data set about women and their environment and basic information and what contraceptive method they use.
-\end{itemize}
-
-\paragraph{}Five classification algorithms were chosen that we used on all data sets. Each chapter will also dicuss how the classifiers perform an the data and why.
-
-% ------------------------------------
-% E C H O   C A R D I O    S E T
-% ------------------------------------
-
-\section{Echo Cardiogram data}
-
-\paragraph{}The Echocardiogram data set originally consists of 132 entries with 13 attributes respectively.
-The entries of the dataset refer to patients that suffered from a heart attack at some point in the past \cite{uci-repo}.
-The prediction task is to determine from several different attributes if the patient will survive at least one year, beginning from the day when the heart attack occured.
-The dataset contains missing values and there's a lot of other preprocessing steps (described later) which have to be done.
-The dataset is very small, especially after preprocessing and a further problem is that the number of patients which died within one year after their heart attack is very small (4) which makes the classification very hard.
-It is nevertheless an interesting question wether some classification algorithm can perform well on that data.
-Since it is very small we performed a 3-fold-cross-validation on this dataset.
+\section{Data set section}
 
@@
 
-\paragraph{}The patient's attribute ``name'' was replaced by a dummy for all entries.
-For this reason, this attribute could be entirely removed.
-Other unnecessary attributes were ``group'' and ``mult'' which were also (as explained in the description) just meaningless and could thus be ignored \cite{uci-repo}.
-There were also two attributes ``wall-motion-score'' and ``wall-motion-index'' where the description of the dataset said that one should only use the ``wall-motion-index'' for classification issues and so ``wall-motion-score'' was also removed.
+\paragraph{}Write about how we preprocessed the data.
 
-\paragraph{}The actual prediction attribute ``alive-at-1'' (denoting that a patient lived at least one year after having a heart attack) had to be inferred from two other attributes, namely ``survival'' (number of months that the test-person survived after the heart attack) and ``still-alive'' (a binary variable denoting if a patient is still alive at the time when the data was recorded).
-It is possible that the ``survival'' variable has a value lower than 12 but the ``still-alive'' variable is set to true, meaning that the test-person still lived.
-This is the case if the data for a test-person was recorded in the last 12 months and so this data cannot be used for the classification and was thus removed.
+\subsection{Classifier}
 
-%\paragraph{}It should be clear that the variables ``survival'' and ``still-alive'' should not be used for classification since they constitute the final prediction attribute.
-Listing \ref{lst:silly-tree} shows what happens if a decision tree can take these attributes into account.
-It states that if the patient survived $ <= 12 $ months, he survived at least 12 month.
-Pretty accurate, right?
-%The only remarkable thing is, that the J48 implementation under default settings seems to enforce a certain depth of the tree, even if the splitting on the same attribute is redundant since $ x > 11 \rightarrow x > 12 $.
-%
-%\begin{lstlisting}[caption={Decision tree on 100\% dependant attributes}, label=lst:silly-tree, float=htb]
-%J48 pruned tree
-%------------------
-%    survival <= 12
-%    |   survival <= 11: 0 (4.0)
-%    |   survival > 11: 1 (4.0)
-%    survival > 12: 1 (87.0)
-%\end{lstlisting}
+=There is a beautiful confusion matrix:
 
-\paragraph{}The last preprocessing step was the replacement of missing values (data imputation).
-Most of the data was complete but there were still some data records having missing values which we replaced by the mean values of the specific variables.
-One data record did contain completeley useless data and was thus removed (see in listing \ref{lst:crappy-entry}).
-
-\begin{lstlisting}[escapechar=\%, float=htb, caption={76 years old patient without useful information}, label=lst:crappy-entry]
-20,1,59,0,0.030,21.300,6.290,17,1.310,0.928,name,2,0
-0.25,1,63,1,?,?,?,23,2.300,0.714,name,2,1
-%\Hilight% ?,?,?,77,?,?,?,?,?,2,?,name,2,?
-2,1,56,1,0.040,14,5,?,?,?,name,2,1
-7,1,61,1,0.270,?,?,9,1.500,0.428,name,2,1
-\end{lstlisting}
-
-\paragraph{}
-
-%So long, these preprocessing tasks are just removing attributes (column-wise) or removing certain entries (row-wise), depending on their values.
-%
-%Unfortunately, there is another peculiarity with the actual prediction attribute ``alive-at-1'', here is the official description:
-%\begin{quotation} ``Boolean-valued.
-Derived from the first two attributes.
-0 means patient was either dead after 1 year or had been followed for less than 1 year.
-1 means patient was alive at 1 year.''\cite{uci-repo} \end{quotation}
-%De facto, the values don't fit to this description.
-The attribute is $1$ if the patient was followed less than one year and $0$ otherwise.
-Very astonishing: Even if the attribute is supposed to be a derived attribute, there are several \texttt{null} values, making the corresponding values useless for training data.
-%
-%We thus inferred the prediction attribute ``alive-at-1'' as explained in the description of the whole dataset: For all patients where the months of survival are greater than or equal to 12 we set the ``alive-at-1''-attribute to 1, otherwise we set it to 0.
-
-%The only thing which remained then was to a
-
-\subsection{Decision Tree}
-
-With default settings J48 Decision Tree seems to show a reasonable tree to some extent as seen in listing \ref{lst:echo:dec:default}.
-
-\begin{lstlisting}[label=lst:echo:dec:default, caption={Vizualization of Decision Tree on echocardiogram data -- Default Settings}]
-J48 pruned tree
-------------------
-epss <= 13100: 1 (90.63/1.95)
-epss > 13100
-|   fractional-shortening <= 190: 1 (2.25/0.05)
-|   fractional-shortening > 190: 0 (2.11/0.11)
-\end{lstlisting}
-
-\paragraph{}The documentation of the data set explains the two attributes as follows:
-
-\begin{quote}
-``epss -- E-point septal separation, another measure of contractility. Larger numbers are increasingly abnormal.''
-\end{quote}
-
-\begin{quote}
-`` fractional-shortening -- a measure of contracility around the heart lower numbers are increasingly abnormal''
-\end{quote}
-
-\paragraph{}Only the rule for the fractional shortening should be the other way round, because a low fractional shortening is a rather pathological sign.
-Although the decision tree has a high precision of almost 95\% it performs very imprecise, because none of the patients who died after one year are correctly classified:
-
-\begin{table}[htb]\centering
-    \begin{tabular}{|ll|ll|} 
+\begin{table*}[htb]\centering
+    \begin{tabular}{ll|ll} 
 \multicolumn{2}{c}{\phantom{bla}}  & \multicolumn{2}{c}{Classified as} \\  
@@
 
-Using an unpruned tree doesn't improve the results:
+And a listing:
+
 \begin{lstlisting}
@@
 \end{lstlisting}
-Here, the decision tree assumes a pericardial effusion (fluid around the heart) to be a healthy characteristic, but it is not.
-Accordingly, the precision of the J48 decreases from 95\% to 88\%.
-Using a higher confidence factor instead of a totally unpruned tree has no effect on the quality or the classification distribution.
 
-\paragraph{Other parameter}
-Neither the laplace correctin to smooth the count of the leaves nor reduced error pruning has any effect on the quality of result.
-Without mdl correction the precision decreases a little to 91.9\%.
-Raising the minimum of instances per leaf has no effect either, which is no suprise since the data set is rather small.
-We couldn't get a better result with the J48 algorithm than to classify all patients to survive at least one year.
-The detailed accuracy for this simple, single-sided classification can be seen in table
-\ref{tab:echo:dec:best}.
+And look at this performance/accuracy table:
 
@@
 
-\subsubsection{Random Forest}
-Let's see if a random forest in place of a simple decision tree can make better result out of the echocardiogram data set.
-With default settings, i.e.
-default number of features $ log_2(7) + 1 \approx 4 $, and 10 random trees with unlimited tree depth, the random forest shows us a rather disappointing result as shown in the confusion matrix in table \ref{tab:echo:rand:matrix1} and the accuracy details in table \ref{tab:echo:rand:default}:
+\subsubsection{Another Classifier}
 
-\begin{table*}[htb]\centering
-    \begin{tabular}{@{}ll|ll@{}} 
-\multicolumn{2}{c}{\phantom{bla}}  & \multicolumn{2}{c}{Classified as} \\  
-                   & \phantom{aa}        &  0  &  1  \\ \cmidrule{2-4}
-\multirow{2}{*}{Instances}         &  0  &  0  &  4  \\
-                                   &  1  &  2  &  91 \\ 
-\end{tabular}
-\caption{Confusion matrix for Random Forest}
-\label{tab:echo:rand:matrix1}
-\end{table*}
+Look what I can predict!
 
-\begin{table*}[htb]\centering
-    \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-        \toprule 
-              &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class \\ \midrule
-              &  0       & 0.022   & 0         & 0      & 0         & 0     \\
-              &  0.978   & 1       & 0.958     & 0.978  & 0.968     & 1     \\
-Weighted Avg. &  0.938   & 0.96    & 0.918     & 0.938  & 0.928     &       \\ \bottomrule
-    \end{tabular*}
-\caption{Random Forest -- Default Settings} 
-\label{tab:echo:rand:default}
-\end{table*}
-\FloatBarrier
-
-\paragraph{Number of Random Trees}
-Reducing the number of the trees continuously decreases the quality of the model, tables \ref{tab:echo:rand:5t} and \ref{tab:echo:rand:1t} show the regression of quality.
-This method increases the number of false positives but doesn't catch any true positives.
-\begin{table*}[htb]\centering
-    \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-        \toprule 
-               &  TP Rate & FP Rate & Precision & Recall & F-Measure &   Class  \\ \midrule
-               &  0       & 0.032   & 0         & 0      & 0         &   0      \\
-               &  0.968   & 1       & 0.957     & 0.968  & 0.963     &   1      \\     
-Weighted Avg.  &  0.928   & 0.96    & 0.918     & 0.928  & 0.923     &          \\ \bottomrule
-    \end{tabular*}
-\caption{Random Forest -- 5 Random Trees} 
-\label{tab:echo:rand:5t}
-\end{table*}
-\FloatBarrier
-
-\begin{table*}[htb]\centering
-    \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-        \toprule 
-               &  TP Rate & FP Rate & Precision & Recall & F-Measure &   Class \\ \midrule
-               &  0       & 0.054   & 0         & 0      & 0         &   0     \\ 
-               &  0.946   & 1       & 0.957     & 0.946  & 0.951     &   1     \\ 
-Weighted Avg.  &  0.907   & 0.961   & 0.917     & 0.907  & 0.912     &         \\ \bottomrule
-    \end{tabular*}
-\caption{Random Forest -- 1 Random Tree} 
-\label{tab:echo:rand:1t}
-\end{table*}
-\FloatBarrier
-\paragraph{Limiting the maximum depth of the trees}
-Limitations in the depths of the random trees has also almost no effect on the accuracy of random forests on this data set.
-The only change to the outcome can be observed as soon the depth of the trees is limited to 6 (see table \ref{tab:echo:rand:matrix2}).
-Only one false positive has gone. Not really a remarkable change in the quality. This does not change if the tree-depth is further reduced.
-\begin{table*}[htb]\centering
-    \begin{tabular}{@{}ll|ll@{}} 
-\multicolumn{2}{c}{\phantom{bla}}  & \multicolumn{2}{c}{Classified as} \\  
-                   & \phantom{aa}        &  0  &  1  \\ \cmidrule{2-4}
-\multirow{2}{*}{Instances}         &  0  & 0 & 4 \\
-                                   &  1  & 1 & 92\\ 
-\end{tabular}
-\caption{Confusion matrix for Random Forest, limited depth}
-\label{tab:echo:rand:matrix2}
-\end{table*}
-
-\paragraph{Increasing the number of features}
-Increasing the number of features chosen by each random tree seems to has also a low effect on the quality, because until a certain border of 5 chosen features the outcome is exactly the same as it was with default settings.
-But if we configure the random forests with 6 features we can spot our first true positive in the confusion matrix (table \ref{tab:echo:rand:matrix3}).
-This gives us the best achievable accuracy that we can reach with decision based algorithms, shown in table \ref{tab:echo:rand:6f}.
-
-\begin{table*}[htb]\centering
-    \begin{tabular}{@{}ll|ll@{}} 
-\multicolumn{2}{c}{\phantom{bla}}  & \multicolumn{2}{c}{Classified as} \\  
-                   & \phantom{aa}        &  0  &  1  \\ \cmidrule{2-4}
-\multirow{2}{*}{Instances}         &  0  & 1 & 3 \\
-                                   &  1  & 2 & 91\\ 
-\end{tabular}
-\caption{Confusion matrix for Random Forest, increased number of features}
-\label{tab:echo:rand:matrix3}
-\end{table*}
-
-\begin{table*}[htb]\centering
-    \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-        \toprule 
-               &  TP Rate &  FP Rate &  Precision &  Recall &  F-Measure &  Class \\ \midrule
-               &  0.25    &  0.022   &  0.333     &  0.25   &  0.286     &  0     \\
-               &  0.978   &  0.75    &  0.968     &  0.978  &  0.973     &  1     \\
-Weighted Avg.  &  0.948   &  0.72    &  0.942     &  0.948  &  0.945     &        \\ \bottomrule
-    \end{tabular*}
-\caption{Random Forest -- 6 randomly chosen features} 
-\label{tab:echo:rand:6f}
-\end{table*}
-\FloatBarrier
-
-\subsection{Support Vector Machine}
-
-\paragraph{}We used the support vector machine implementation called `SMO' in WEKA.
-Since a support vector machine uses the numeric coordinates of the data records to do the classification we tried to normalize the data in advance to obtain better results.
-Since this normalization did not have any impact on the results (we tried this not only on this dataset) we followed that the support vector machine algorithm of the WEKA-toolkit automatically normalizes all the numeric values.
-
-\paragraph{}The build of the model took not even a second since the dataset is that small.
-The same holds for the evaluation.
-We tried all the different kernel-functions for the classification but the result didn't change.
-The following table shows the results of the evaluation and if you take a close look on the results, you can see that all data records were classified to the same class which means that the SVM algorithm wasn't able to cope with the small number of data records which should be classified with value 0:
-
-\begin{table}[h]
-\centering
-\begin{tabular}{lllllll}
-	\toprule
-									&TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure   &Class\\
-	\midrule
-									&0         	&0         &0         	&0       &  0              &0\\
-  								&1         	&1         &0.959     	&1       &0.979             &1\\
-
-	Weighted Avg.		&0.959     	&0.959     &0.919   		&0.959   &0.939     		&\\
-	\bottomrule
-\end{tabular}
-\caption{Results of the SVM-classification for the echocardiogram dataset}
-\end{table}
-
-\vspace{6pt}
-
-\begin{table}[h]
-\centering
-\begin{tabular}{|l|ll|}
-\hline
- & a &	b \\
-\hline
-a=0 & 0 & 4 \\
-b=1 & 0 & 93\\
-\hline
-\end{tabular}
-\caption{Confusion matrix for SVM}
-\end{table}
-
-\subsection{k-Nearest Neighbour}
-
-\paragraph{}As with the SVM-algorithm there was no normalization necessary (we tested this on different datasets).
-In WEKA there are different implementations of the kNN-algorithm available and we chose (for all our datasets) the brute force kNN-algorithm since this is the classic implementation of the kNN-algorithm where we were interested in the running time.
-We tried different parameters concerning the number of neighbours which are considered for classification.
-Changes on that parameter didn't make any differences.
-As with the SVM, the kNN-algorithm was not able to cope with the small number of data records which should be classified with value 0.
-The creation of the model doesn't take any time with a kNN-algorithm since there is no model which would be built: kNN is a lazy classification algorithm.
-The evaluation of the model took not even a second.
-In the following are the results of the classification.
-The results are exactly the same as for the SVM.
-
-\begin{table}[h]
-\centering
-\begin{tabular}{lllllll}
-	\toprule
-									&TP Rate   	&FP Rate   &Precision &Recall  &F-Measure   &Class\\
-	\midrule
-									&0         	&0         &0         &0       &  0         &0\\
-  								&1         	&1         &0.959     &1       &0.979       &1\\
-
-	Weighted Avg.		&0.959     	&0.959     &0.919   	&0.959   &0.939     					&\\
-	\bottomrule
-\end{tabular}
-\caption{Results of the kNN-classification for the echocardiogram dataset}
-\end{table}
-
-\vspace{6pt}
-
-\begin{table}[h]
-\centering
-\begin{tabular}{|l|ll|}
-\hline
- & a &	b \\
-\hline
-a=0 & 0 & 4 \\
-b=1 & 0 & 93\\
-\hline
-\end{tabular}
-\caption{Confusion matrix for kNN}
-\end{table}
-
-\subsection{Naive Bayes}
-
-\paragraph{}For the fist try, the unsupervised discretizer was used. But the result was bad. The classifier could predict one single instance for for patients that did not survive one year.
-
-\paragraph{}For the second try we prepare the data set with the PKIDiscretizer.
-The description says it "Discretizes numeric attributes using equal frequency binning, where the number of bins is equal to the square root of the number of non-missing values." But as we can see in \ref{tab:echo:bayes:1c} this discretizer does not perform better.
-
-\begin{table}[h]
-\centering
-\begin{tabular}{|l|ll|}
-\hline
- & a &	b \\
-\hline
-a=0 & 0 & 4 \\
-b=1 & 1 & 92\\
-\hline
-\end{tabular}
-\caption{Confusion matrix for kNN}
-\label{tab:echo:bayes:1c}
-\end{table}
-
-\paragraph{}The log output of Naive Bayes shows that each bin has 1-2 negative (alive-at-1 = 0) instances. This means that each bin has the same low possibility for TP-prediction.
-
-\paragraph{}So we have to boost the instances we want to predict.
-For that the SMOTE filter is used.
-It produces more instances of a given class index by looking at near neighbours with the same class value.
-The filter was applied several times, until the number of instances that have alive-at-1 = 0 1.2 times more than the alive-at-1 = 1 instances (128:93).
-The SMOTE filter used 3 nearest neighbours and produced 100\% new instances in each run.
-In this way we raised the numbers of instances with an outcome we want to predict, as for this training set it is better to predict if a patient is at risk to not survive one year.
-
-\paragraph{}We have to supply a separate set for validation as we kind of destroy the data set when generating new instances.
-For that the original training set was taken.
-Although it is not good to validate a classifier with the same set as the training set, we can use this approach because our training set does not have a lot in common with the original set as we added a lot of generated instances.
-
-\paragraph{}The confusion matrix \ref{tab:echo:bayes:2c} shows, that we predicted three out of four correct instances and also had more errors when predicting alive-at-1 = 0 than  alive-at-1 = 1.
-
-\begin{table}[h]
-\centering
-\begin{tabular}{|l|ll|}
-\hline
- & a &	b \\
-\hline
-a=0 & 3 & 1 \\
-b=1 & 9 & 84\\
-\hline
-\end{tabular}
-\caption{Confusion matrix for kNN}
-\label{tab:echo:bayes:2c}
-\end{table}
-
-\paragraph{}Also the detailed accuracy table \ref{tab:echo:bayes:2d} shows that this approach gained a high recall with a low precision for alive-at-1 = 0 and both high precision and recall for alive-at-1 = 1.
-
-\begin{table*}[htb]\centering
-  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-      \toprule 
-              &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class    \\  \midrule
-              &   0.75   & 0.097   & 0.25      & 0.75   & 0.375     & 0 \\
-              &   0.903  & 0.25    & 0.988     & 0.903  & 0.944     & 1 \\
-Weighted Avg. &   0.897  & 0.244   & 0.958     & 0.897  & 0.92      &   \\ \bottomrule
-  \end{tabular*}
-\caption{Naive Bayes with SMOTE filter} 
-\label{tab:echo:bayes:2d}
-\end{table*}
-
-\subsection{Comparison}
-
-The table below shos a comparison of the different algorithms on the adult dataset.
-For an algorithm we present here only the best obtained result concerning the TP rate.
-Values are always the weighted average.
-
-\begin{table}[h]
-\centering
-\begin{tabular}{llllllll}
-	\toprule
-									&TP Rate   	&FP Rate   &Precision &Recall  &F-Measure \\
-	\midrule
-	Decision Tree		\\
-  Random Forest		\\
-  Naive Bayes			&   0.897  & 0.244   & 0.958     & 0.897  & 0.92     \\
-  SVM							&0.959     	&0.959     	&0.919   	&0.959   &0.939    \\
-  kNN							&0.959     	&0.959     	&0.919   	&0.959   &0.939    \\
-	\bottomrule
-\end{tabular}
-\caption{Comparison of the different algorithms for the echocardiogram dataset}
-\end{table}
-
-% ------------------------------------
-% A D U L T   D A T A    S E T
-% ------------------------------------
-
-\section{Adult data}
-
-\paragraph{}The Adult or Census Income data sets contains 48842 instances of information about persons and the task is to predict if a person earns more than 50.000 dollar a year.
-The data set has 14 attributes and missing values.
-
-\paragraph{}The instances origin from the 1994 Census Database and were extracted by Barry Becker.
-For having a clean data set, only records with age above 16 with minimium one working hour per week have been selected from the database.
-We performed a 10-fold-cross-validation on this dataset.
-
-\subsection{Preprocessing}
-
-\paragraph{}The ``Adult'' dataset has missing values which we all replaced by the mean value of the respective attribute.
-We did no further preprocessing.
-As for the Echocardiogram dataset there was no normalization neccessary for the SVM or the kNN algorithms.
-
-\subsection{Decision Tree}
-\paragraph{}Using a decision tree yields a precision about 86\%. Varying the parameters for this algorithm influences the quality of the results only slightly, as the following examples illustrate.
-
-\paragraph{}First of all, with the default settings of the J48 implementation of weka, i.e.
-using pruning with a confidence factor of 0.25 together with mdl correction and subtree raising and at least 2 instances per leaf of the tree, we get a summary of the quality as shown in table \ref{tab:adult:dec:default}:
-The decision tree that was used for this run has a size of 413 and has 311 leaves.
-
-\begin{table*}[htb]\centering
-    \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-        \toprule 
-               & TP Rate & FP Rate & Precision & Recall &  F-Measure &  Class \\ \midrule  
-               & 0.615   & 0.062   & 0.759     & 0.615  &  0.68      &  >50K  \\  
-               & 0.938   & 0.385   & 0.885     & 0.938  &  0.911     &  <=50K \\  
-Weighted Avg.  & 0.86    & 0.307   & 0.854     & 0.86   &  0.855     &        \\  \bottomrule 
-    \end{tabular*}
-\caption{Decision Tree -- Default Settings} 
-\label{tab:adult:dec:default}
-\end{table*}
-\FloatBarrier
-
-\paragraph{Raising the confidence factor}
-Subsequently increasing the confidence factor, which result in less pruning, leads to a continuous worse quality.
-If we set the confidence factor to 0.5 we get an outcome as shown in table \ref{tab:adult:dec:cf5} and if we set it to 0.75 the quality gets even worse as seen in \ref{tab:adult:dec:cf75}.
-\begin{table*}[htb]\centering
-    \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-        \toprule 
-              &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class    \\  \midrule
-              &  0.632   & 0.071   & 0.738     & 0.632  & 0.681     & >50K     \\ 
-              &  0.929   & 0.368   & 0.888     & 0.929  & 0.908     & <=50K    \\ 
-Weighted Avg. &  0.857   & 0.296   & 0.852     & 0.857  & 0.854     &          \\  \bottomrule
-    \end{tabular*}
-\caption{Decision Tree -- Confidence Factor of 0.5} 
-\label{tab:adult:dec:cf5}
-\end{table*}
-\FloatBarrier
-
-\begin{table*}[htb]\centering
-    \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-        \toprule 
-              &  TP Rate & FP Rate & Precision & Recall &  F-Measure &  Class  \\ \midrule        
-              &  0.655   & 0.094   & 0.689     & 0.655  &  0.672     &  >50K   \\          
-              &  0.906   & 0.345   & 0.892     & 0.906  &  0.899     &  <=50K  \\          
-Weighted Avg. &  0.846   & 0.285   & 0.843     & 0.846  &  0.844     &         \\ \bottomrule          
-    \end{tabular*}
-\caption{Decision Tree -- Confidence Factor of 0.75} 
-\label{tab:adult:dec:cf75}
-\end{table*}
-\FloatBarrier
-
-\paragraph{Other parameters}
-Varying other parameters has a similar effect like the one mentioned above.
-Either the quality of the J48 algorithm decreases a little or it stays the same, thus never exceeding an overall quality of 86\%.
-To sum it up: Using an unpruned tree yields the same quality as a tree with a confidence factor of 0.75.
-Smoothing the counts at the leaves with laplace has no effect.
-Either without mdl correction or without Subtree raising during pruning, in both cases the precision falls from 86\% to 85.5\%.
-If we rise the minimum of instances per leaves, e.g.
-to 20 instances and 50 instances, the precision falls from 86\% to 85\% and 84\% respectively.
-
-\subsubsection{Random Forest}
-
-With default settings, a random forest gains an accuracy on the census income data set that can be seen in table \ref{tab:adult:rand:default}, which is surprisingly worse than the J48 algorithm.
-\begin{table*}[htb]\centering
-  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-      \toprule 
-              &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class    \\  \midrule
-              &  0.615   & 0.108   & 0.644     & 0.615  & 0.629     & >50K     \\     
-              &  0.892   & 0.385   & 0.88      & 0.892  & 0.886     & <=50K    \\ 
-Weighted Avg. &  0.825   & 0.318   & 0.823     & 0.825  & 0.824     &          \\ \bottomrule
-    \end{tabular*}
-\caption{Random Forest -- Default Settings} 
-\label{tab:adult:rand:default}
-\end{table*}
-\FloatBarrier
-
-\paragraph{Number of Random Trees}
-As expected, reducing the number of random trees reduces the quality, with 5 random trees the precision lays at 82.6\% with only 3 random trees it lays at 82.3\%.
-Interestingly, the quality gets higher if we take even more than 10 random trees into account: With 15 trees, the precision is at 83.2\% and with 25 random trees it is at 83.3\%.
-
-\paragraph{Limiting the maximum depth of the trees}
-We have a sign for overfitting under default settings: Reducing the maximum depth has a positive effect on the quality.
-From size 20 over 10 until the size of 7 the precision of the randomeforest rises until 85\%.
-Reducing the depth even further lets the precision fall again.
-Table \ref{tab:adult:rand:6d} shows the detailed accuracy for a random forest of seven trees which can grow only to a height of 6:
-
-\begin{table*}[htb]\centering
-  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-      \toprule 
-               &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class \\ \midrule
-               &  0.591   & 0.057   & 0.766     & 0.591  & 0.667     & >50K  \\
-               &  0.943   & 0.409   & 0.879     & 0.943  & 0.91      & <=50K \\
-Weighted Avg.  &  0.858   & 0.324   & 0.852     & 0.858  & 0.851     &       \\ \bottomrule
-    \end{tabular*}
-\caption{Random Forest -- 7 Random Trees each not more than 6 branches high} 
-\label{tab:adult:rand:6d}
-\end{table*}
-\FloatBarrier
-
-\paragraph{Number of features}
-Increasing the number of features doesn't support the quality at all.
-The default number of features for this data set with 13 attributes is $ log_2(13) + 1 \approx 5 $.
-Choosing 6 attributes brinchoose gs us a precision of 84,9\% and with 10 randomly chosen features the precision falls back to 84.5\%.
-We can tease the most out of the random forest if we decrease the number of features a little.
-To be concise, we get the best achievable result, if we take only 3 randomly chosen
-features. Table \ref{tab:adult:rand:3f} shows these results which are of similar accuracy like the best of the J48 algorithm.
-
-\begin{table*}[htb]\centering
-  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-      \toprule 
-               &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class \\   \midrule
-               &  0.577   & 0.052   & 0.779     & 0.577  & 0.663     & >50K  \\   
-               &  0.948   & 0.423   & 0.876     & 0.948  & 0.911     & $\leq$50K \\   
-Weighted Avg.  &  0.859   & 0.333   & 0.853     & 0.859  & 0.851     &       \\   \bottomrule
-    \end{tabular*}
-\caption{Random Forest -- 3 Random Features} 
-\label{tab:adult:rand:3f}
-\end{table*}
-\FloatBarrier
-
-\subsection{Support Vector Machine}
-
-\paragraph{}From the different kernel functions for the SVM-algorithm we used the PolyKernel on this dataset.
-As with the dataset before the normalization of the numeric data had no inpact on the validation results.
-One interesting aspect is that the creation of the model and the 10-fold-cross-validation took 135 minutes since the dataset is big.
-The SVM-algorithm was able to correctly classify 84,9\% of the instances.
-The following two tables show further validation results including a confusion matrix:
-
-\begin{table}[h]
-\centering
-\begin{tabular}{llllllll}
-	\toprule
-									&TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure  	&Class\\
-	\midrule
-									&0.568     	&0.062     &0.744     	&0.568   &0.644    	    	&	>50K\\
-  								&0.938     	&0.432     &0.873     	&0.938   &0.904         	&$\leq$50K\\
-
-	Weighted Avg.   & 0.849    & 0.343     &0.842   		&0.849   &0.842     			&\\
-	\bottomrule
-\end{tabular}
-\caption{Results of the SVM-classification for the adult dataset}
-\end{table}
-
-\vspace{6pt}
-
-\begin{table}[h]
-\centering
-\begin{tabular}{|l|ll|}
-\hline
-              &    a &     b \\
-\hline
- a = >50k      & 4454 &  3387 \\
- b = $\leq$50k & 1531 & 23189\\
-\hline
-\end{tabular}
-\caption{Confusion matrix for SVM}
-\end{table}
-
-\subsection{Naive Bayes}
-
-\paragraph{}As further preprocessing for Naive Bayes it is important to dicretize the numerical attributes.
-The PKIDescretizer was used again as it produces better bin sizes than the standard discretizer without any configuration.
-After the first run with the Simple Naive Bayes algorithm we gain the result shown in table \ref{tab:adult:bayes:1d} and a confusion matrix \ref{tab:adult:bayes:1c}.
-
-\begin{table*}[htb]\centering
-  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-      \toprule 
-               &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class \\   \midrule
-               &  0.806   & 0.163   & 0.611     & 0.806  & 0.695     & >50K  \\
-               &  0.837   & 0.194   & 0.932     & 0.837  & 0.882     & <=50K \\
-Weighted Avg.  &  0.83    & 0.186   & 0.854     & 0.83   & 0.837     &       \\   \bottomrule
-    \end{tabular*}
-\caption{Naive Bayes -- preprocessed with PKIDiscretizer} 
-\label{tab:adult:bayes:1d}
-\end{table*}
-
-\begin{table}[h]
-\centering
-\begin{tabular}{|l|ll|}
-\hline
-              &    a &     b \\
-\hline
- a = >50k      & 6321 &1520   \\
- b = $\leq$50k & 4025 &20695 \\
-\hline
-\end{tabular}
-\caption{Confusion matrix for Naive Bayes}
-\label{tab:adult:bayes:1c}
-\end{table}
-
-\paragraph{}By using different discretizers and also do not seperately discretize in the preprocessing section of weka but instead let the NaiveBayers classifier use it's own discretizers namely KernelEstimator or SupervisedDiscretizer the outcome of precision and recall differ a bit by 0.05 to 0.1 higher and lower.
-It does obviously do not really make a difference.
-
-\subsection{k-Nearest Neighbour}
-
-\paragraph{}We tested the k-nearest neighbour algorithm with different values for $k$.
-For $k$=1 (nearest neighbour) the algorithm was able to correctly classify 79,4\% of the instances.
-We tested the algorithm also with values 10,40 and 100 for $k$ and obtained the best results with $k=40$, where 83,48\% of the instances were correctly classified.
-Again the normalization of the input had no influence on the results.
-The evaluation of this algorithm took 120 seconds.
-In the following tables we present further results of the evaluation with $k=40$ including a confusion matrix.
-
-\begin{table}[h]
-\centering
-\begin{tabular}{lllllll}
-	\toprule
-									&TP Rate   &FP Rate   &Precision   &Recall  &F-Measure &Class\\
-	\midrule
-									&0.593     &0.088      &0.68      &0.593     &0.633    &>50K\\
-									&0.912     &0.407      &0.876     &0.912     &0.893    &$\leq$50K\\
-
-	Weighted Avg.   &0.835     &0.331      &0.829     &0.835     &0.831     \\
-	\bottomrule
-\end{tabular}
-\caption{Results of the kNN-classification for the adult dataset}
-\end{table}
-
-\vspace{6pt}
-
-\begin{table}[h]
-\centering
-\begin{tabular}{|l|ll|}
-	\hline
-	 &              a    & b     \\
-	\hline
-    a = >50K      &	4647 & 3194  \\
-    b = $\leq$50K & 2186 & 22534 \\
-    \hline
-\end{tabular}
-\caption{Confusion matrix for kNN}
-\end{table}
-
-\subsection{Comparison}
-
-The table below shos a comparison of the different algorithms on the adult dataset.
-For an algorithm we present here only the best obtained result concerning the TP rate.
-Values are always the weighted average.
-
-\begin{table}[h]
-\centering
-\begin{tabular}{lllllll}
-	\toprule
-									&TP Rate   	&FP Rate   &Precision &Recall  &F-Measure  \\
-	\midrule
-	Decision Tree		& 0.86    	& 0.307   	& 0.854   & 0.86   & 0.855     \\
-  Random Forest		& 0.859   	& 0.333   	& 0.853   & 0.859  & 0.851    \\
-  Naive Bayes		 	&  0.83    & 0.186   & 0.854     & 0.83   & 0.837     \\
-  SVM							& 0.849    	& 0.343     &0.842   	&0.849   &0.842     \\
-  kNN							&0.835     	&0.331      &0.829    &0.835   &0.831			\\
-	\bottomrule
-\end{tabular}
-\caption{Comparison of the different algorithms for the adult dataset}
-\end{table}
-
-% ------------------------------------
-% C O N T R A S E P T I V E    S E T
-% ------------------------------------
-
-\section{Contraceptive Method Choice Data}
-
-\paragraph{}The Contraceptive Method Choice Data Set consists of information about the demographic and socio-economic characteristics of married women.
-There are 1473 instances with 9 different attributes.
-Examples of these attributes are the wife's age, husbands education, wife's religion, etc.
-The task is to predict which kind of contraceptive method a woman uses.
-These are categorized as no use, short term or long term.
-We performed a 10-fold-cross-validation on this dataset.
-
-\subsection{Preprocessing}
-
-\paragraph{}The values of the data sets's attributes are already in a numerical and categorical form, whereby the categories are encoded in an ordinal form.
-These go from 1 to 4 where 1 means low/bad and 4 means high/good.
-The number-of-children attribute is numeric but the values range from 0 to 16 so that a preprocessing is not necessary as one can think of 15 classes (15 distinct values).
-This set has no missing values and the attributes do not depend on each other.
-So we can use the data as is.
-
-\subsection{Decision Tree}
-
-With default settings, the J48 algorithm comes off very badly, in table \ref{tab:contra:dec:default} one can see a that it classifies about the half of all instances correctly.
-It must be apologetically supplemented that the task includes more distinct values, three instead of two nominals, to be predicted.
-
-\begin{table*}[htb]\centering
-  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-      \toprule 
-               &  TP Rate & FP Rate & Precision & Recall & F-Measure &  Class \\ \midrule 
-               &  0.609   & 0.287   & 0.613     & 0.609  & 0.611     &  1     \\  
-               &  0.378   & 0.138   & 0.445     & 0.378  & 0.409     &  2     \\      
-               &  0.507   & 0.318   & 0.458     & 0.507  & 0.481     &  3     \\      
-Weighted Avg.  &  0.521   & 0.264   & 0.521     & 0.521  & 0.52      &        \\ \bottomrule     
-    \end{tabular*}
-\caption{Decision Tree on Contraceptive Data Set -- Default Settings} 
-\label{tab:contra:dec:default}
-\end{table*}
-\FloatBarrier
-
-\paragraph{Raising the confidence factor}
-As expected, raising the confidence factor worsens the result.
-For a series of a factor 0.25, 0.5, 0.25 and a totally unpruned tree, the quality decreases from originally 0.52\% to 50.4\%, 48.6\% and finally 48.7\%.
-The used decision tree grows from a size of 157 leaves and a maximum depth of 263 to a huge tree with 417 leaves and 655 branches.
-As said, this is a definitely wrong direction of optimizing.
-
-\paragraph{Raising the minimum of instances per leaf}
-We can notice a certain improvement to the outcome if we ensure that every leaf of the decision tree has to contain more instances, thereby preventing the tree from deep nesting.
-We can make the most out of it if we set the minimum to 23 leaves, the yield can be seen in table \ref{tab:contra:dec:23l}.
-This tree has only 31 leaves and 49 branches.
-Further enhancing of this parameter cuts the amount of correctly classified instances again.
-
-\begin{table*}[htb]\centering
-  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-      \toprule 
-              &   TP Rate & FP Rate & Precision & Recall & F-Measure &  Class \\ \midrule      
-              &   0.595   & 0.231   & 0.657     & 0.595  & 0.624     &  1     \\      
-              &   0.405   & 0.135   & 0.467     & 0.405  & 0.434     &  2     \\      
-              &   0.575   & 0.334   & 0.478     & 0.575  & 0.522     &  3     \\      
-Weighted Avg. &   0.545   & 0.245   & 0.552     & 0.545  & 0.546     &        \\ \bottomrule     
-    \end{tabular*}
-\caption{Decision Tree on Contraceptive Data Set -- at least 23 instances per leaf} 
-\label{tab:contra:dec:23l}
-\end{table*}
-\FloatBarrier
-\paragraph{Other parameters}
-Neither laplace correction nor mdl correction has any positive effect on the output of the algorithm on this data set.
-The absence of mdl correction make the outcome a bit worse, it brings the precision down to 0.551\%.
-But there is one small tweak from which the accuracy of the algorithm can benefit: If binary splits are used on nominal attributes, we get the best result for the J48 on this data set as seen in table \ref{tab:contra:dec:bin}.
-
-\begin{table*}[htb]\centering
-  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-      \toprule 
-               &  TP Rate & FP Rate & Precision & Recall & F-Measure &  Class \\ \midrule   
-               &  0.649   & 0.263   & 0.648     & 0.649  & 0.648     &  1     \\    
-               &  0.402   & 0.13    & 0.475     & 0.402  & 0.436     &  2     \\        
-               &  0.548   & 0.292   & 0.499     & 0.548  & 0.522     &  3     \\    
-Weighted Avg.  &  0.558   & 0.243   & 0.557     & 0.558  & 0.557     &        \\ \bottomrule 
-    \end{tabular*}
-\caption{Decision Tree on Contraceptive Data Set -- using binary splits} 
-\label{tab:contra:dec:bin}
-\end{table*}
-\FloatBarrier
-
-\subsubsection{Random Forest}
-Random Forests perform even worse than the J48 with default settings. We can see in table \ref{tab:contra:rand:default} that only the half of all classified instances are classified correctly.
-
-\begin{table*}[htb]\centering
-  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-      \toprule 
-               &  TP Rate & FP Rate & Precision & Recall & F-Measure  & Class  \\  \midrule      
-               &  0.631   & 0.328   & 0.589     & 0.631  & 0.609      & 1      \\       
-               &  0.36    & 0.16    & 0.397     & 0.36   & 0.378      & 2      \\       
-               &  0.456   & 0.274   & 0.469     & 0.456  & 0.462      & 3      \\       
-Weighted Avg.  &  0.509   & 0.271   & 0.504     & 0.509  & 0.506      &        \\  \bottomrule     
-    \end{tabular*}
-\caption{Random Forests on Contraceptive Data Set -- Default Settings} 
-\label{tab:contra:rand:default}
-\end{table*}
-\FloatBarrier
-
-\paragraph{More but shorter trees}
-Random forests perform best on the data set with a maximum depth of 5 branches.
-After this improvement we can lift the precision up to 54.4\%.
-Furthermore, our random forest benefits from adding five more random trees so we can enhance the precision to  54.7\%.
-\paragraph{More Features}
-The most significant improvement can then observed after increasing the number of randomly chosen features.
-By default, the number of chosen features for this data set lays about 4 ( $ log_2(10) + 1 $ ).
-Choosing two more of them results in the best accuracy for random forests on this data set, as shown in table \ref{tab:contra:rand:6f}.
-
-\begin{table*}[htb]\centering
-  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-      \toprule 
-               &  TP Rate & FP Rate & Precision & Recall & F-Measure &  Class \\  \midrule     
-               &  0.604   & 0.231   & 0.661     & 0.604  & 0.631     &  1     \\  
-               &  0.408   & 0.127   & 0.484     & 0.408  & 0.443     &  2     \\  
-               &  0.593   & 0.326   & 0.491     & 0.593  & 0.537     &  3     \\  
-Weighted Avg.  &  0.556   & 0.241   & 0.562     & 0.556  & 0.556     &        \\  \bottomrule 
-    \end{tabular*}
-\caption{Random Forests on Contraceptive Data Set -- with at least 6 chosen features} 
-\label{tab:contra:rand:6f}
-\end{table*}
-\FloatBarrier
-
-\subsection{Support Vector Machine}
-
-\paragraph{}We tested different kernel functions on this dataset and obtained the best results with the NormalizedPolyKernel.
-The whole process of the evaluation, including the building of the models, took 28 seconds.
-Normalization of the data had again no influence on the results.
-The SVM algorithm was able to correctly classify 49,2\% of the instances.
-The following tables show further results of the evaluation with the NormalizedPolyKernel including a confusion matrix.
-
-\begin{table}[htb]
-\centering
-\begin{tabular}{lllllll}
-	\toprule
-	&								TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure     	&Class\\ 
-	\midrule
-									&0.583     	&0.355     &0.55      	&0.583   &0.566   		    	&1\\
-                 	&0.393     	&0.181     &0.389     	&0.393   &0.391      	    	&2\\
-                 	&0.444     	&0.252     &0.484     	&0.444   &0.463      	    	&3\\
-	Weighted Avg.   &0.492     	&0.28      &0.491     	&0.492   &0.491     			&\\
-	\bottomrule
-\end{tabular}
-\caption{Results of the SVM-classification for the contraception dataset}
-\end{table}
-
-\vspace{6pt}
-
-\begin{table}[h]
-\centering
-\begin{tabular}{|l|lll|}
-	\hline
-	      &a   &b   &c \\
-	\hline
-	a = 1 &367 &105 &157 \\
- 	b = 2 &117 &131 &85 \\
- 	c = 3 &183 &101 &227 \\
-  \hline
-\end{tabular}
-\caption{Confusion matrix for SVM}
-\end{table}
-
-\subsection{k-Nearest Neighbour}
-
-\paragraph{}We tested the kNN-algorithm with a lot of different values for $k$ and obtained the best results for $k=7$ where 48,88\% of the instances were correctly classified.
-The evaluation took not even a second.
-As with the other datasets the normalization of the input had no influence on the results.
-In the following tables we present further results of the evaluation with $k=7$ including a confusion matrix.
-
-\begin{table}[h]
-\centering
-\begin{tabular}{lllllll}
-	\toprule
-	&								TP Rate   	&FP Rate   &Precision   &Recall  &F-Measure     	&Class\\
-	\midrule
-									&0.569     	&0.351     &0.547     	&0.569   &0.558      	     	&1\\
-                 	&0.414     	&0.166     &0.422     	&0.414   &0.418      	    	&2\\
-                 	&0.438     	&0.279     &0.455     	&0.438   &0.447      	    	&3\\
-
-	Weighted Avg.   &0.489     	&0.284     &0.487     	&0.489   &0.488      				&\\
-	\bottomrule
-\end{tabular}
-\caption{Results of the kNN-classification for the contraception dataset}
-\end{table}
-
-\vspace{6pt}
-
-\begin{table}[h]
-\centering
-\begin{tabular}{|l|lll|}
-	\hline
-	      &a   &b   &c \\
-	\hline
-    a = 1 &358 &99  &172\\
-    b = 2 &99  &138 &96 \\
- 	c = 3 &197 &90  &224 \\
-  \hline
-\end{tabular}
-\caption{Confusion matrix for kNN}
-\end{table}
-
-\subsection{Naive Bayes}
-
-\paragraph{}The input data set does perfectly fit for Naive Bayes.
-But the classifier is not able to correctly predict the instances.
-As shown in tables \ref{tab:contra:bayes:1d} and \ref{tab:contra:bayes:1c} it gains a 50:50 chance of a TP prediction.
-
-\begin{table*}[htb]\centering
-  \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-      \toprule 
-               &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class \\   \midrule
-               &  0.52    & 0.223   & 0.635     & 0.52   & 0.572     & 1  \\  
-               &  0.526   & 0.245   & 0.385     & 0.526  & 0.445     & 2  \\
-               &  0.481   & 0.268   & 0.488     & 0.481  & 0.485     & 3  \\
-Weighted Avg.  &  0.508   & 0.243   & 0.528     & 0.508  & 0.513     &    \\   \bottomrule
-    \end{tabular*}
-\caption{Naive Bayes} 
-\label{tab:contra:bayes:1d}
-\end{table*}
-
-\begin{table}[h]
-\centering
-\begin{tabular}{|l|lll|}
-\hline
-              &   a &   b &   c  \\
-\hline
- a = 1        & 327 & 141 & 161  \\
- b = 2        &  61 & 175 &  97  \\
- c = 3        & 127 & 138 & 246  \\
-\hline
-\end{tabular}
-\caption{Confusion matrix for Naive Bayes}
-\label{tab:contra:bayes:1c}
-\end{table}
-
-\subsection{Comparison}
-
-The table below shos a comparison of the different algorithms on the adult dataset.
-For an algorithm we present here only the best obtained result concerning the TP rate.
-Values are always the weighted average.
-
-\begin{table}[h]
-\centering
-\begin{tabular}{lllllll}
-	\toprule
-									&TP Rate   	&FP Rate   &Precision &Recall  &F-Measure   \\
-	\midrule
-	Decision Tree 	&  0.558   & 0.243   & 0.557     	& 0.558  & 0.557     \\
-  Random Forest		&  0.556   & 0.241   & 0.562     	& 0.556  & 0.556     \\
-  Naive Bayes		 	&  0.508   & 0.243   & 0.528     	& 0.508  & 0.513 \\
-  SVM							& 0.492    &0.280     &0.491     	&0.492   & 0.491 		 \\
-  kNN							&0.489     &0.284    &0.487     	& 0.489  & 0.488 		 \\
-	\bottomrule
-\end{tabular}
-\caption{Comparison of the different algorithms for the contraceptive method choice dataset}
-\end{table}
 \bibliography{references}

### 2013-01-02 exercise3/BayesianNetwork/src/main/java/at/ac/tuwien/machine_learning/SillyClassifier.java
added R script to load our java classes in RWeka     importing our SillyClassifier works     prediction with our classifier works

@@ -1,59 +1,135 @@
 package at.ac.tuwien.machine_learning;
 
+import java.util.Enumeration;
 
+import weka.classifiers.AbstractClassifier;
+import weka.core.Attribute;
 import weka.core.Capabilities;
 import weka.core.Capabilities.Capability;
-import weka.classifiers.Classifier;
-import weka.core.Instances;
 import weka.core.Instance;
+import weka.core.Instances;
+import weka.core.WeightedInstancesHandler;
 
-public class SillyClassifier extends Classifier {
+public class SillyClassifier extends AbstractClassifier implements WeightedInstancesHandler {
 
-	
-	private static final long serialVersionUID = 4148318713741359627L;
-	
+	/**
+	 * 
+	 */
+	private static final long serialVersionUID = -4598198166434716177L;
+
+	private Attribute m_Class;
+
 	private double m_ClassValue;
-	
+
 	public String globalInfo() {
-	    return "Class for building and using a very silly classifier. Simply predicts the value of a class attribute.";	    
-	  }
-	
+		return "Class for building and using a very silly classifier. Permanently classifies new instances with the first class attribute of the training data.";
+	}
+
+	/**
+	 * Returns default capabilities of the classifier.
+	 * 
+	 * @return the capabilities of this classifier
+	 */
 	public Capabilities getCapabilities() {
-	    Capabilities result = super.getCapabilities();
-	    result.disableAll();
+		Capabilities result = super.getCapabilities();
+		result.disableAll();
 
-	    // attributes
-	    result.enable(Capability.NOMINAL_ATTRIBUTES);
-	    result.enable(Capability.NUMERIC_ATTRIBUTES);
-	    result.enable(Capability.DATE_ATTRIBUTES);
-	    result.enable(Capability.MISSING_VALUES);
-	 
-	    // class
-	    result.enable(Capability.NOMINAL_CLASS);
-	    result.enable(Capability.MISSING_CLASS_VALUES);
+		// attributes
+		result.enable(Capability.NOMINAL_ATTRIBUTES);
+		result.enable(Capability.NUMERIC_ATTRIBUTES);
+		result.enable(Capability.DATE_ATTRIBUTES);
+		result.enable(Capability.STRING_ATTRIBUTES);
+		result.enable(Capability.RELATIONAL_ATTRIBUTES);
+		result.enable(Capability.MISSING_VALUES);
 
-	    // instances
-	    result.setMinimumNumberInstances(0);
-	    
-	    return result;
-	  }
+		// class
[…]

### 2013-01-02 exercise3/BayesianNetwork/testOwnWekaClassifier.R

@@ -1,0 +1,14 @@
+#!/usr/bin/Rscript
+require("RWeka")
+require("rJava")
+.jaddClassPath("./target/machine_learning-0.0.1-SNAPSHOT.jar")
+
+classifier <- make_Weka_classifier("at/ac/tuwien/machine_learning/SillyClassifier")
+classifier 
+data(iris)
+model = classifier(Species ~., data=iris)
+model
+
+result = table(iris$Species, predict(model))
+result
+

### 2013-01-03 exercise3/BayesianNetwork/bayesNetTest.R
added R-Script to test normal execution of a bayesian network classifier

@@ -1,0 +1,10 @@
+#!/usr/bin/Rscript
+require("RWeka")
+
+bayesNet <- make_Weka_classifier("weka/classifiers/bayes/BayesNet")
+
+res = bayesNet(Species~., data=iris, control = Weka_control(D="",  Q = "weka.classifiers.bayes.net.search.local.HillClimber", "--", P=1, S="BAYES", E="weka.classifiers.bayes.net.estimate.SimpleEstimator", "--", A=0.5))
+predict(res)
+
+
+

### 2013-01-03 exercise2/paper.tex
added some lines in paper

@@
 
-\paragraph{}This is a template.
+\paragraph{}This assignment is about extending a Bayes Network Classifier by adding a custom search algorithm.
 
-\section{Data set section}
+\section{Setup}
+
+Used stuff: R.
+RWeka -- an interface in R for Weka.
+implement a searcher by extending a java class from the weka library/framework.
+running a R script that starts the weka BayesNet-classifier with our custom searcher.
 

### 2013-01-03 exercise2/R/testOwnWekaClassifier.R
reorderd R-scripts to own folder

@@ -1,7 +1,7 @@
 #!/usr/bin/Rscript
 require("RWeka")
 require("rJava")
-.jaddClassPath("./target/machine_learning-0.0.1-SNAPSHOT.jar")
+.jaddClassPath("../BayesianNetwork/target/machine_learning-0.0.1-SNAPSHOT.jar")
 
 classifier <- make_Weka_classifier("at/ac/tuwien/machine_learning/SillyClassifier")
 classifier 

### 2013-01-03 exercise2/R/bayesNetTest.R
bayesNetTest.R now uses our own UltraSearcher which is nothing more than a copy of HillClimber (at the moment!)

@@ -1,9 +1,12 @@
 #!/usr/bin/Rscript
 require("RWeka")
+require("rJava")
+
+.jaddClassPath("../BayesianNetwork/target/machine_learning-0.0.1-SNAPSHOT.jar")
 
 bayesNet <- make_Weka_classifier("weka/classifiers/bayes/BayesNet")
 
-res = bayesNet(Species~., data=iris, control = Weka_control(D="",  Q = "weka.classifiers.bayes.net.search.local.HillClimber", "--", P=1, S="BAYES", E="weka.classifiers.bayes.net.estimate.SimpleEstimator", "--", A=0.5))
+res = bayesNet(Species~., data=iris, control = Weka_control(D="",  Q = "at.ac.tuwien.machine_learning.UltraSearcher", "--", P=1, S="BAYES", E="weka.classifiers.bayes.net.estimate.SimpleEstimator", "--", A=0.5))
 predict(res)
 
 

### 2013-01-03 exercise2/BayesianNetwork/src/main/java/at/ac/tuwien/machine_learning/UltraSearcher.java
added confusion matrix to output

@@ -1,0 +1,665 @@
+/*
+ *   This program is free software: you can redistribute it and/or modify
+ *   it under the terms of the GNU General Public License as published by
+ *   the Free Software Foundation, either version 3 of the License, or
+ *   (at your option) any later version.
+ *
+ *   This program is distributed in the hope that it will be useful,
+ *   but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *   MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *   GNU General Public License for more details.
+ *
+ *   You should have received a copy of the GNU General Public License
+ *   along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+/*
+ * HillClimber.java
+ * Copyright (C) 2004-2012 University of Waikato, Hamilton, New Zealand
+ * 
+ */
+ 
+package at.ac.tuwien.machine_learning;
+
+import java.io.Serializable;
+import java.util.Enumeration;
+import java.util.Vector;
+
+import weka.classifiers.bayes.BayesNet;
+import weka.classifiers.bayes.net.ParentSet;
+import weka.classifiers.bayes.net.search.local.LocalScoreSearchAlgorithm;
+import weka.core.Instances;
+import weka.core.Option;
+import weka.core.RevisionHandler;
+import weka.core.RevisionUtils;
+import weka.core.Utils;
+
+/** 
+ <!-- globalinfo-start -->
+ * This Bayes Network learning algorithm uses a hill climbing algorithm adding, deleting and reversing arcs. The search is not restricted by an order on the variables (unlike K2). The difference with B and B2 is that this hill climber also considers arrows part of the naive Bayes structure for deletion.
+ * <p/>
+ <!-- globalinfo-end -->
+ *
+ <!-- options-start -->
+ * Valid options are: <p/>
+ * 
+ * <pre> -P &lt;nr of parents&gt;
+ *  Maximum number of parents</pre>
+ * 
+ * <pre> -R
+ *  Use arc reversal operation.
+ *  (default false)</pre>
+ * 
+ * <pre> -N
+ *  Initial structure is empty (instead of Naive Bayes)</pre>
+ * 
+ * <pre> -mbc
+ *  Applies a Markov Blanket correction to the network structure, 
+ *  after a network structure is learned. This ensures that all 
[…]

### 2013-01-03 exercise2/paper/paper.tex
moved and edited paper

--- 2013-01-03 exercise2/paper.tex
@@
 
+\section{TODO}
+
+\begin{itemize}
+\item Structure for paper
+\item Implementation of cutsom bn-searcher
+\item documentation of cutsom bn-searcher
+\item comparison and interpretation of the results
+\item description of the used R and Weka features and commands
+\item description of the setup/infrastructure (weka, custom searcher, R, runtime, comparison, data set, ...)
+\item R script that runs the custom searcher
+\item R script that runs a default searcher
+\item find and describe a custom search algorithm we want to uses
+\end{itemize}
+
 \section{Preface}
@@
 running a R script that starts the weka BayesNet-classifier with our custom searcher.
+
+\subsection{RWeka}
+
+RWeka is a interface in R that wraps the Weka framework and has a really good documentation.
+
+Some good stuff to know:
+
+\texttt{WOW("at.ac.tuwien.machine\_learning.UltraSearcher")} WOW is a function in R that reads and prints all possible options of a weka class.
+This can be done for classifiers, filters, searchers and all other types in weka.
 

### 2013-01-05 exercise2/paper/paper.tex
paper

@@
 
-\subsection{Preprocessing}
+\subsection{Weka Javacode}
 
-\paragraph{}Write about how we preprocessed the data.
+\paragraph{initStructure()} of class BayesNet initializes the Bayesnetwork. It creates an adjacency matrix where each attribute has an array of parents. These arrays are empty after initialization.
 
-\subsection{Classifier}
-
-=There is a beautiful confusion matrix:
-
-\begin{table*}[htb]\centering
-    \begin{tabular}{ll|ll} 
-\multicolumn{2}{c}{\phantom{bla}}  & \multicolumn{2}{c}{Classified as} \\  
-                        & \phantom{aa}  &  0  &  1  \\ \cmidrule{2-4}
-\multirow{2}{*}{Instances}     &  0  &  0  &  4  \\
-                                   &  1  &  1  &  90 \\ 
-\end{tabular}
-\caption{Confusion matrix for Decision tree}
-\end{table*}
-
-And a listing:
+\paragraph{} The default SearchAlgorithm.class initializes the network like a Naive Bayes network. This shows how we can use the adjacency matrix of BayesNet:
 
 \begin{lstlisting}
-J48 unpruned tree
-pericardial-effusion = 0
-|   epss <= 13100: 1 (76.62/1.95)
-|   epss > 13100
-|   |   epss <= 15600: 0 (2.19/0.16)
-|   |   epss > 15600: 1 (2.19/0.03)
-pericardial-effusion = 1: 1 (14.0)
+...
+// initialize parent sets to have arrow from classifier node to
+// each of the other nodes
+for (int iAttribute = 0; iAttribute < instances.numAttributes(); iAttribute++) {
+    if (iAttribute != iClass) {
+        bayesNet.getParentSet(iAttribute).addParent(iClass, instances);
+    }
+}
+...
 \end{lstlisting}
 
-And look at this performance/accuracy table:
-
-\begin{table*}[htb]\centering
-    \begin{tabular*}{\columnwidth}{@{}lllllll@{}}
-        \toprule 
-               &  TP Rate & FP Rate & Precision & Recall & F-Measure &  Class  \\ \midrule
-               &  0       & 0       & 0         & 0      & 0         &  0      \\     
-               &  1       & 1       & 0.959     & 1      & 0.979     &  1      \\     
-Weighted Avg.  &  0.959   & 0.959   & 0.919     & 0.959  & 0.939     &         \\ \bottomrule
-    \end{tabular*}
-\caption{Decision Tree -- best achievable results} 
-\label{tab:echo:dec:best}
-\end{table*}
-\FloatBarrier
-
-\subsubsection{Another Classifier}
-
-Look what I can predict!
+\paragraph{} Where \texttt{BayesNet} is the classifier that holds the network datastructure and \texttt{getParentSet(n)} returns the \texttt{nth} row in the adjacency matrix.
 

### 2013-01-05 exercise2/BayesianNetwork/src/main/java/at/ac/tuwien/machine_learning/SillyClassifier.java
added first version of iterAlgo

@@ -1,135 +1,0 @@
-package at.ac.tuwien.machine_learning;
-
-import java.util.Enumeration;
-
-import weka.classifiers.AbstractClassifier;
-import weka.core.Attribute;
-import weka.core.Capabilities;
-import weka.core.Capabilities.Capability;
-import weka.core.Instance;
-import weka.core.Instances;
-import weka.core.WeightedInstancesHandler;
-
-public class SillyClassifier extends AbstractClassifier implements WeightedInstancesHandler {
-
-	/**
-	 * 
-	 */
-	private static final long serialVersionUID = -4598198166434716177L;
-
-	private Attribute m_Class;
-
-	private double m_ClassValue;
-
-	public String globalInfo() {
-		return "Class for building and using a very silly classifier. Permanently classifies new instances with the first class attribute of the training data.";
-	}
-
-	/**
-	 * Returns default capabilities of the classifier.
-	 * 
-	 * @return the capabilities of this classifier
-	 */
-	public Capabilities getCapabilities() {
-		Capabilities result = super.getCapabilities();
-		result.disableAll();
-
-		// attributes
-		result.enable(Capability.NOMINAL_ATTRIBUTES);
-		result.enable(Capability.NUMERIC_ATTRIBUTES);
-		result.enable(Capability.DATE_ATTRIBUTES);
-		result.enable(Capability.STRING_ATTRIBUTES);
-		result.enable(Capability.RELATIONAL_ATTRIBUTES);
-		result.enable(Capability.MISSING_VALUES);
-
-		// class
-		result.enable(Capability.NOMINAL_CLASS);
-		result.enable(Capability.NUMERIC_CLASS);
-		result.enable(Capability.DATE_CLASS);
-		result.enable(Capability.MISSING_CLASS_VALUES);
-
-		// instances
-		result.setMinimumNumberInstances(0);
-
-		return result;
-	}
-
-	/**
-	 * Generates the classifier.
-	 * 
-	 * @param instances
-	 *            set of instances serving as training data
-	 * @throws Exception
-	 *             if the classifier has not been generated successfully
-	 */
-	public void buildClassifier(Instances instances) throws Exception {
-		// can classifier handle the data?
-		getCapabilities().testWithFail(instances);
-
-		// remove instances with missing class
-		instances = new Instances(instances);
-		instances.deleteWithMissingClass();
-
-		m_Class = instances.classAttribute();
-		m_ClassValue = 0;
-
[…]

### 2013-01-05 exercise2/BayesianNetwork/src/main/java/weka/classifiers/bayes/net/search/global/IteratedLocalSearch.java

@@ -1,33 +1,35 @@
+package weka.classifiers.bayes.net.search.global;
+
 /*
- *   This program is free software: you can redistribute it and/or modify
- *   it under the terms of the GNU General Public License as published by
- *   the Free Software Foundation, either version 3 of the License, or
- *   (at your option) any later version.
- *
- *   This program is distributed in the hope that it will be useful,
- *   but WITHOUT ANY WARRANTY; without even the implied warranty of
- *   MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- *   GNU General Public License for more details.
- *
- *   You should have received a copy of the GNU General Public License
- *   along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ * This program is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU General Public License as published by
+ * the Free Software Foundation; either version 2 of the License, or
+ * (at your option) any later version.
+ * 
+ * This program is distributed in the hope that it will be useful,
+ * but WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ * GNU General Public License for more details.
+ * 
+ * You should have received a copy of the GNU General Public License
+ * along with this program; if not, write to the Free Software
+ * Foundation, Inc., 675 Mass Ave, Cambridge, MA 02139, USA.
  */
 
 /*
  * HillClimber.java
- * Copyright (C) 2004-2012 University of Waikato, Hamilton, New Zealand
+ * Copyright (C) 2004 University of Waikato, Hamilton, New Zealand
  * 
  */
- 
-package at.ac.tuwien.machine_learning;
 
 import java.io.Serializable;
 import java.util.Enumeration;
+import java.util.Random;
 import java.util.Vector;
 
 import weka.classifiers.bayes.BayesNet;
 import weka.classifiers.bayes.net.ParentSet;
-import weka.classifiers.bayes.net.search.local.LocalScoreSearchAlgorithm;
+import weka.classifiers.bayes.net.search.global.GlobalScoreSearchAlgorithm;
 import weka.core.Instances;
 import weka.core.Option;
 import weka.core.RevisionHandler;
@@ -59,37 +61,40 @@ import weka.core.Utils;
[…]

### 2013-01-05 exercise2/R/bayesNetTest.R
asdf

@@ -6,12 +6,11 @@ require("rJava")
 
 bayesNet <- make_Weka_classifier("weka/classifiers/bayes/BayesNet")
 
-res = bayesNet(Species~., data=iris, control = Weka_control(D="",  Q = "at.ac.tuwien.machine_learning.UltraSearcher", "--", P=1, S="BAYES", E="weka.classifiers.bayes.net.estimate.SimpleEstimator", "--", A=0.5))
-predict(res)
+model = bayesNet(Species~., data=iris, control = Weka_control(D="",  Q = "weka.classifiers.bayes.net.search.global.IteratedLocalSearch", "--", P=1, S="BAYES", E="weka.classifiers.bayes.net.estimate.SimpleEstimator", "--", A=0.5))
+predict(model)
 
 
 
-result = table(iris$Species, predict(res))
-result
+result = table(iris$Species, predict(model))
 
-
+evaluate_Weka_classifier(model)

### 2013-01-05 exercise2/BayesianNetwork/src/main/java/weka/classifiers/bayes/net/search/global/IteratedLocalSearch.java
added changes of benjamin

@@ -1,5 +1,3 @@
-package weka.classifiers.bayes.net.search.global;
-
 /*
  * This program is free software; you can redistribute it and/or modify
  * it under the terms of the GNU General Public License as published by
@@ -22,218 +20,264 @@ package weka.classifiers.bayes.net.search.global;
  * 
  */
 
+package weka.classifiers.bayes.net.search.global;
+
 import java.io.Serializable;
+import java.util.ArrayList;
 import java.util.Enumeration;
+import java.util.List;
 import java.util.Random;
 import java.util.Vector;
 
 import weka.classifiers.bayes.BayesNet;
 import weka.classifiers.bayes.net.ParentSet;
-import weka.classifiers.bayes.net.search.global.GlobalScoreSearchAlgorithm;
 import weka.core.Instances;
 import weka.core.Option;
 import weka.core.RevisionHandler;
 import weka.core.RevisionUtils;
 import weka.core.Utils;
 
-/** 
- <!-- globalinfo-start -->
- * This Bayes Network learning algorithm uses a hill climbing algorithm adding, deleting and reversing arcs. The search is not restricted by an order on the variables (unlike K2). The difference with B and B2 is that this hill climber also considers arrows part of the naive Bayes structure for deletion.
+/**
+ * <!-- globalinfo-start --> This Bayes Network learning algorithm uses a hill
+ * climbing algorithm adding, deleting and reversing arcs. The search is not
+ * restricted by an order on the variables (unlike K2). The difference with B
+ * and B2 is that this hill climber also considers arrows part of the naive
+ * Bayes structure for deletion.
  * <p/>
- <!-- globalinfo-end -->
- *
- <!-- options-start -->
- * Valid options are: <p/>
+ * <!-- globalinfo-end -->
  * 
- * <pre> -P &lt;nr of parents&gt;
- *  Maximum number of parents</pre>
+ * <!-- options-start --> Valid options are:
+ * <p/>
  * 
- * <pre> -R
+ * <pre>
+ * -P &lt;nr of parents&gt;
+ *  Maximum number of parents
+ * </pre>
+ * 
+ * <pre>
+ * -R
  *  Use arc reversal operation.
- *  (default false)</pre>
+ *  (default false)
+ * </pre>
  * 
- * <pre> -N
- *  Initial structure is empty (instead of Naive Bayes)</pre>
+ * <pre>
+ * -N
+ *  Initial structure is empty (instead of Naive Bayes)
+ * </pre>
  * 
- * <pre> -mbc
+ * <pre>
[…]

### 2013-01-05 exercise2/R/bayesNetTest.R
R-script now uses cmc and adult data set

@@ -6,9 +6,13 @@ require("rJava")
 
 bayesNet <- make_Weka_classifier("weka/classifiers/bayes/BayesNet")
 
+cmc = read.arff("../../exercise1/samples/cmc_preprocessed.arff")
+adult =  read.arff("../../exercise1/samples/census/adult_preprocessed.arff")
 
-hcModel = bayesNet(Species~., data=iris, control = Weka_control(D="",  Q = "weka.classifiers.bayes.net.search.global.HillClimber", "--", P=1, S="BAYES", E="weka.classifiers.bayes.net.estimate.SimpleEstimator", "--", A=0.5))
-extendedModel = bayesNet(Species~., data=iris, control = Weka_control(D="",  Q = "weka.classifiers.bayes.net.search.global.IteratedLocalSearch", "--", P=1, S="BAYES", E="weka.classifiers.bayes.net.estimate.SimpleEstimator", "--", A=0.5))
+names(adult)
+
+hcModel = bayesNet(`contraceptive-method`~., data=cmc, control = Weka_control(D="",  Q = "weka.classifiers.bayes.net.search.global.HillClimber", "--", P=1, S="BAYES", E="weka.classifiers.bayes.net.estimate.SimpleEstimator", "--", A=0.5))
+extendedModel = bayesNet(`contraceptive-method`~., data=cmc, control = Weka_control(D="",  Q = "weka.classifiers.bayes.net.search.global.IteratedLocalSearch", "--", P=1, S="BAYES", E="weka.classifiers.bayes.net.estimate.SimpleEstimator", "--", A=0.5))
 
 
 

### 2013-01-05 exercise2/BayesianNetwork/src/main/java/weka/classifiers/bayes/net/search/global/IteratedLocalSearch.java
added "times" parameter to pertubate(.. method

@@ -186,7 +186,7 @@ public class IteratedLocalSearch extends GlobalScoreSearchAlgorithm {
 		int numberOfIterationsWithNoBetterSolution = 0;
 
 		do {
-			perturbate(bayesNet, instances);
+			perturbate(bayesNet, instances, 3);
 			localSearch(bayesNet, instances);
 			scoreOfFoundSolution = calcScore(bayesNet);
 			System.out.println("Score of found solution: " + scoreOfFoundSolution);
@@ -216,14 +216,16 @@ public class IteratedLocalSearch extends GlobalScoreSearchAlgorithm {
 		System.out.println("End local search");
 	} // search
 
-	private void perturbate(BayesNet bayesNet, Instances instances)
+	private void perturbate(BayesNet bayesNet, Instances instances, Integer times)
 			throws Exception {
-		System.out.println("Start perturbation");
+		System.out.println("Start " + times + " perturbations");
+		for (@SuppressWarnings("unused") Integer i : new int[times]){
 		Operation randomOperation = getRandomOperation(bayesNet, instances);
 		performOperation(bayesNet, instances, randomOperation);
+		}
 		System.out.println("End perturbation");
 	}
-
+	
 	private ParentSet[] copyParentSets(ParentSet[] parentSets) {
 		ParentSet[] copy = new ParentSet[parentSets.length];
 		for (int i = 0; i < copy.length; i++) {

### 2013-01-05 exercise2/paper/paper.tex

@@
 
+%
+%
+%
 \section{TODO}
@@
 \item description of the setup/infrastructure (weka, custom searcher, R, runtime, comparison, data set, ...)
-\item R script that runs the custom searcher
-\item R script that runs a default searcher
-\item find and describe a custom search algorithm we want to uses
 \end{itemize}
 
+%
+%
+%
 \section{Preface}
@@
 \paragraph{}This assignment is about extending a Bayes Network Classifier by adding a custom search algorithm.
+We chose the Hill Climber search algorithm for an implementation of iterative local search algorithm.
 
+%
+%
+%
 \section{Setup}
 
-Used stuff: R.
-RWeka -- an interface in R for Weka.
-implement a searcher by extending a java class from the weka library/framework.
-running a R script that starts the weka BayesNet-classifier with our custom searcher.
+\paragraph{}To run the search algorithm an R script has be written.
+It loads Weka,builds the classifier, runs it and evaluates the model.
+It also shows the performance tables and the differences of the original Hill Climber implementation and our improved one.
 
-\subsection{RWeka}
+\paragraph{} Tools needed to run the code: R (RWeka, rJava), Java (JRE).
 
-RWeka is a interface in R that wraps the Weka framework and has a really good documentation.
+\paragraph{} RWeka is a interface in R that wraps the Weka framework and has a really good documentation.
 
-Some good stuff to know:
+%
+%
+\section{Current Implementation}
 
-\texttt{WOW("at.ac.tuwien.machine\_learning.UltraSearcher")} WOW is a function in R that reads and prints all possible options of a weka class.
-This can be done for classifiers, filters, searchers and all other types in weka.
-
-\subsection{Weka Javacode}
-
-\paragraph{initStructure()} of class BayesNet initializes the Bayesnetwork. It creates an adjacency matrix where each attribute has an array of parents. These arrays are empty after initialization.
+First we take a look at the BayesNet classifier implementation.
+When the classifier is startet, first thing is to generate the classification model.
+Therefore the BayesNet classifier initializes (\texttt{initStructure()})an adjacency matrix where each attribute has an array of parents.
+These arrays are empty after initialization.
 
@@
 
+%
+%
+%
+\section{Hill Climber}
+
+Current implementation of Weka's Hill Climber algorithm.
+
+\begin{lstlisting}
+double fScore = calcScore(bayesNet);
+Operation oOperation = getOptimalOperation(bayesNet, instances);
+while ((oOperation != null) && (oOperation.m_fScore > fScore)) {
+	performOperation(bayesNet, instances, oOperation);
+	fScore = oOperation.m_fScore;
+	oOperation = getOptimalOperation(bayesNet, instances);
+}  
+\end{lstlisting}
+
+\paragraph{}As shown in the listing, Hill Climber calculates an optimal solution based on the current network iteratively until no opration can be found or the new network is only slightly better than the previous one.
+To find the at least worse next operation all possible operations get evaluated and the best one is returned.
+
+\section{Extended Hill Climber}
+
+TODO: describe our extension.
+
+\section{Comparison}
+
+TODO:
+* used datasets
+* accuracy results
+* runtime/performance
+
 \bibliography{references}

### 2013-01-05 exercise2/BayesianNetwork/src/main/java/weka/classifiers/bayes/net/search/global/IteratedLocalSearch.java
added console option to iteratedLocalSearch algorithm

@@ -161,6 +161,17 @@ public class IteratedLocalSearch extends GlobalScoreSearchAlgorithm {
 
 	/** use the arc reversal operator **/
 	boolean m_bUseArcReversal = false;
+	
+	/** how many random operation per pertubation ? **/ 
+	int numberOfPertubations = 1;
+
+	public int getNumberOfPertubations() {
+		return numberOfPertubations;
+	}
+
+	public void setNumberOfPertubations(int numberOfPertubations) {
+		this.numberOfPertubations = numberOfPertubations;
+	}
 
 	/**
 	 * search determines the network structure/graph of the network with the
@@ -186,7 +197,7 @@ public class IteratedLocalSearch extends GlobalScoreSearchAlgorithm {
 		int numberOfIterationsWithNoBetterSolution = 0;
 
 		do {
-			perturbate(bayesNet, instances, 3);
+			perturbate(bayesNet, instances, getNumberOfPertubations());
 			localSearch(bayesNet, instances);
 			scoreOfFoundSolution = calcScore(bayesNet);
 			System.out.println("Score of found solution: " + scoreOfFoundSolution);
@@ -218,7 +229,7 @@ public class IteratedLocalSearch extends GlobalScoreSearchAlgorithm {
 
 	private void perturbate(BayesNet bayesNet, Instances instances, Integer times)
 			throws Exception {
-		System.out.println("Start " + times + " perturbations");
+		System.out.println("Start pertubation with " + times + " random operations");
 		for (@SuppressWarnings("unused") Integer i : new int[times]){
 		Operation randomOperation = getRandomOperation(bayesNet, instances);
 		performOperation(bayesNet, instances, randomOperation);
@@ -576,6 +587,11 @@ public class IteratedLocalSearch extends GlobalScoreSearchAlgorithm {
 		newVector.addElement(new Option(
 				"\tInitial structure is empty (instead of Naive Bayes)", "N",
 				0, "-N"));
+		
+		newVector.addElement(new Option(
+				"\tNumber of random operations during pertubation", "T",
+				1, "-T <nr of random operations>"));
+		
 
 		Enumeration enu = super.listOptions();
 		while (enu.hasMoreElements()) {
@@ -644,9 +660,14 @@ public class IteratedLocalSearch extends GlobalScoreSearchAlgorithm {
 		} else {
 			setMaxNrOfParents(100000);
 		}
+		
[…]

### 2013-01-05 exercise2/BayesianNetwork/src/main/java/weka/classifiers/bayes/net/search/global/IteratedLocalSearch.java
refactord iterated local search

@@ -136,8 +136,7 @@ public class IteratedLocalSearch extends GlobalScoreSearchAlgorithm {
 			if (other == null) {
 				return false;
 			}
-			return ((m_nOperation == other.m_nOperation)
-					&& (m_nHead == other.m_nHead) && (m_nTail == other.m_nTail));
+			return ((m_nOperation == other.m_nOperation) && (m_nHead == other.m_nHead) && (m_nTail == other.m_nTail));
 		} // equals
 
 		/** number of the tail node **/
@@ -161,8 +160,8 @@ public class IteratedLocalSearch extends GlobalScoreSearchAlgorithm {
 
 	/** use the arc reversal operator **/
 	boolean m_bUseArcReversal = false;
-	
-	/** how many random operation per pertubation ? **/ 
+
+	/** how many random operation per pertubation ? **/
 	int numberOfPertubations = 1;
 
 	public int getNumberOfPertubations() {
@@ -184,8 +183,7 @@ public class IteratedLocalSearch extends GlobalScoreSearchAlgorithm {
 	 * @throws Exception
 	 *             if something goes wrong
 	 */
-	protected void search(BayesNet bayesNet, Instances instances)
-			throws Exception {
+	protected void search(BayesNet bayesNet, Instances instances) throws Exception {
 		m_BayesNet = bayesNet;
 		bayesNet.setDebug(true);
 		System.out.println("Score before local search: " + calcScore(bayesNet));
@@ -213,8 +211,7 @@ public class IteratedLocalSearch extends GlobalScoreSearchAlgorithm {
 		} while (numberOfIterationsWithNoBetterSolution < 5);
 	} // search
 
-	protected void localSearch(BayesNet bayesNet, Instances instances)
-			throws Exception {
+	protected void localSearch(BayesNet bayesNet, Instances instances) throws Exception {
 		System.out.println("Start local search");
 		double fScore = calcScore(bayesNet);
 		// go do the search
@@ -227,16 +224,17 @@ public class IteratedLocalSearch extends GlobalScoreSearchAlgorithm {
 		System.out.println("End local search");
 	} // search
 
-	private void perturbate(BayesNet bayesNet, Instances instances, Integer times)
-			throws Exception {
+	private void perturbate(BayesNet bayesNet, Instances instances, Integer times) throws Exception {
[…]

### 2013-01-05 exercise2/R/bayesNetTest.R
added command line argument (number of random operations per pertubation) to the R script

@@ -9,10 +9,8 @@ bayesNet <- make_Weka_classifier("weka/classifiers/bayes/BayesNet")
 cmc = read.arff("../../exercise1/samples/cmc_preprocessed.arff")
 adult =  read.arff("../../exercise1/samples/census/adult_preprocessed.arff")
 
-names(adult)
-
-hcModel = bayesNet(`contraceptive-method`~., data=cmc, control = Weka_control(D="",  Q = "weka.classifiers.bayes.net.search.global.HillClimber", "--", P=1, S="BAYES", E="weka.classifiers.bayes.net.estimate.SimpleEstimator", "--", A=0.5))
-extendedModel = bayesNet(`contraceptive-method`~., data=cmc, control = Weka_control(D="",  Q = "weka.classifiers.bayes.net.search.global.IteratedLocalSearch", "--", P=1, S="BAYES", E="weka.classifiers.bayes.net.estimate.SimpleEstimator", "--", A=0.5))
+hcModel = bayesNet(`Species`~., data=iris, control = Weka_control(D="",  Q = "weka.classifiers.bayes.net.search.global.HillClimber", "--", P=1, S="BAYES", E="weka.classifiers.bayes.net.estimate.SimpleEstimator", "--", A=0.5))
+extendedModel = bayesNet(`Species`~., data=iris, control = Weka_control(D="",  Q = "weka.classifiers.bayes.net.search.global.IteratedLocalSearch", "--", T=5, P=1, S="BAYES", E="weka.classifiers.bayes.net.estimate.SimpleEstimator", "--", A=0.5))
 
 
 
@@ -23,10 +21,11 @@ hcEvaluation$details
 
 
 extendedEvaluation = evaluate_Weka_classifier(extendedModel)
-"New Iteradet Local Search Classifier"
+"New Iteratet tocal Search Classifier"
 extendedEvaluation$details
 
 print("Difference of details")
 extendedEvaluation$details - hcEvaluation$details 
 print("Difference of confusion matrices")
 extendedEvaluation$confusionMatrix- hcEvaluation$confusionMatrix
+

### 2013-01-06 exercise2/BayesianNetwork/src/main/java/weka/classifiers/bayes/net/search/global/IteratedLocalSearch.java
fix of silly bug

@@ -318,12 +318,12 @@ public class IteratedLocalSearch extends GlobalScoreSearchAlgorithm {
 
 		switch (operationType) {
 		case (Operation.OPERATION_ADD):
-			int attempt = 0, maxAttempts = bayesNet.getNrOfNodes(); // maxAttempts is rather arbitrarlily chosen
+			int attempt = 0, maxAttempts = 2*bayesNet.getNrOfNodes(); // maxAttempts is rather arbitrarlily chosen
 			do {
 				iAttributeHead = nodesWithMissingParents.get(random.nextInt(nodesWithMissingParents.size()));
 				iAttributeTail = random.nextInt(instances.numAttributes());
 				attempt++;
-				if (attempt < maxAttempts) return null; // seems we can't add valid arcs?
+				if (attempt >= maxAttempts) return null; // seems we can't add valid arcs?
 				
 			} while (!addArcMakesSense(bayesNet, instances, iAttributeHead, iAttributeTail));
 		// possible infinite loop?

### 2013-01-07 exercise2/paper/paper.tex
added stuff to paper

--- 2013-01-05 exercise2/paper/paper.tex
@@
 \lstset{
-    captionpos=b
+    captionpos=b,
+    basicstyle=\small
 }
@@
 
+\setlength{\parskip}{\baselineskip}%
 \setlength\parindent{0pt} % Removes all indentation from paragraphs - comment this line for an assignment with lots of text
-
 %----------------------------------------------------------------------------------------
@@
 \begin{itemize}
-\item Structure for paper
 \item Implementation of cutsom bn-searcher
@@
 \item comparison and interpretation of the results
-\item description of the used R and Weka features and commands
-\item description of the setup/infrastructure (weka, custom searcher, R, runtime, comparison, data set, ...)
-\item R script that runs the custom searcher
-\item R script that runs a default searcher
-\item find and describe a custom search algorithm we want to uses
 \end{itemize}
 
+%
+%
+%
 \section{Preface}
 
-\paragraph{}This assignment is about extending a Bayes Network Classifier by adding a custom search algorithm.
+This assignment is about extending a Bayes Network Classifier by adding a custom search algorithm.
+We chose the Hill Climber search algorithm for an implementation of an iterative local search algorithm.
 
+%
+%
+%
 \section{Setup}
 
-Used stuff: R.
-RWeka -- an interface in R for Weka.
-implement a searcher by extending a java class from the weka library/framework.
-running a R script that starts the weka BayesNet-classifier with our custom searcher.
+There is a R script to execute the searcher in the handed in package at \texttt{R/bayesNetTest.R}.
+It loads Weka, builds the classifier, runs it and evaluates the model.
+It also shows the performance tables and the differences of the original Hill Climber implementation and our improved one.
+To run the script the following Tools need to be installed: R with the packages RWeka and rJava, Java (JRE).
 
-\subsection{RWeka}
+The used package RWeka is an interface in R that wraps the Weka framework and has a really good documentation.
 
-RWeka is a interface in R that wraps the Weka framework and has a really good documentation.
+%
+%
+\section{BayesNet}
 
-Some good stuff to know:
+First we take a look at the BayesNet classifier implementation.
+When the classifier is startet, first thing is to generate the classification model.
+Therefore the BayesNet classifier initializes an adjacency matrix (\texttt{initStructure()}) where each attribute has an array of parents.
+These arrays are empty after initialization.
 
-\texttt{WOW("at.ac.tuwien.machine\_learning.UltraSearcher")} WOW is a function in R that reads and prints all possible options of a weka class.
-This can be done for classifiers, filters, searchers and all other types in weka.
-
-\subsection{Weka Javacode}
-
-\paragraph{initStructure()} of class BayesNet initializes the Bayesnetwork. It creates an adjacency matrix where each attribute has an array of parents. These arrays are empty after initialization.
-
-\paragraph{} The default SearchAlgorithm.class initializes the network like a Naive Bayes network. This shows how we can use the adjacency matrix of BayesNet:
-
-\begin{lstlisting}
+\begin{lstlisting}[caption={Initialization of BayesNet},label=initStructure]
 ...
@@
 
-\paragraph{} Where \texttt{BayesNet} is the classifier that holds the network datastructure and \texttt{getParentSet(n)} returns the \texttt{nth} row in the adjacency matrix.
+The default Weka's SearchAlgorithm initializes the network like a Naive Bayes network.
+Listing \ref{initStructure} shows how we can use the adjacency matrix of BayesNet.
+Where \texttt{BayesNet} is the classifier that holds the network datastructure and \texttt{getParentSet(n)} returns the \texttt{nth} row in the adjacency matrix.
+
+%
+%
+%
+\section{Hill Climber}
+
+Implementation of Weka's Hill Climber algorithm.
+
+%
+%
+\subsection{Description}
+
+As shown in listing \ref{hcSearch}, Hill Climber calculates an optimal solution based on the current network iteratively until no opration can be found to improve the network or the new network is only slightly better than the previous one.
+To find the best operation all possible operations get evaluated and the best one is returned.
+The algorithm stops, when it finds a local optima.
+
+\begin{lstlisting}[caption={HillClimbers search algorithm},label=hcSearch]
+double fScore = calcScore(bayesNet);
+Operation oOperation = getOptimalOperation(bayesNet, instances);
+while ((oOperation != null) && (oOperation.m_fScore > fScore)) {
+	performOperation(bayesNet, instances, oOperation);
+	fScore = oOperation.m_fScore;
+	oOperation = getOptimalOperation(bayesNet, instances);
+}  
+\end{lstlisting}
+
+%
+%
+\subsection{Test Results}
+
+Two data sets are used to test the search algorithm. These are x and y from the assignment 1.
+
+\todo{insert results of test}
+
+As shown in the result listing, \todo{describe result} ....
+
+%
+%
+%
+\section{Extended Hill Climber}
+
+To improve the Hill Climber algorithm we used it with an iterated local search algorithm.
+
+\subsection{Description}
+
+Like the Hill Climber, our improved search algorithm introduces an initial Naive Bayes like network where each attribute has the classification attribute as parent.
+
+Then the Hill Climber algorithm is executed to find a local optima.
+To find other local optima, our search algorithm iteratively selects a neighbor of the found network and uses it as initial network for the Hill Climber to hopefully find better local optima than the first one.
+
+ The intial network we use in each iteration is a random neighbor of the previously found one. It does not matter if it is better or worse.
+
+If the Hill Climber finds a better network it is stored as the current best found network and if the seacher cannot find a better network after a given number of iterations, the algorithm stops and yields the best solution found.
+
+%
+%
+%
+\section{Comparison}
+
+TODO:
+* used datasets
+* accuracy results
+* runtime/performance
 

### 2013-01-08 exercise2/BayesianNetwork/src/main/java/weka/classifiers/bayes/net/search/global/IteratedLocalSearch.java
changed some of the code in the perturbation and added a neighborhood-operator

@@ -107,6 +107,7 @@ public class IteratedLocalSearch extends GlobalScoreSearchAlgorithm {
 		final static int OPERATION_ADD = 0;
 		final static int OPERATION_DEL = 1;
 		final static int OPERATION_REVERSE = 2;
+		final static int OPERATION_HEADCHANGE = 3;
 
 		/** c'tor **/
 		public Operation() {
@@ -136,7 +137,8 @@ public class IteratedLocalSearch extends GlobalScoreSearchAlgorithm {
 			if (other == null) {
 				return false;
 			}
-			return ((m_nOperation == other.m_nOperation) && (m_nHead == other.m_nHead) && (m_nTail == other.m_nTail));
+			return ((m_nOperation == other.m_nOperation)
+					&& (m_nHead == other.m_nHead) && (m_nTail == other.m_nTail));
 		} // equals
 
 		/** number of the tail node **/
@@ -162,7 +164,7 @@ public class IteratedLocalSearch extends GlobalScoreSearchAlgorithm {
 	boolean m_bUseArcReversal = false;
 
 	/** how many random operation per pertubation ? **/
-	int numberOfPertubations = 1;
+	int numberOfPertubations = 3;
 
 	public int getNumberOfPertubations() {
 		return numberOfPertubations;
@@ -183,24 +185,29 @@ public class IteratedLocalSearch extends GlobalScoreSearchAlgorithm {
 	 * @throws Exception
 	 *             if something goes wrong
 	 */
-	protected void search(BayesNet bayesNet, Instances instances) throws Exception {
+	protected void search(BayesNet bayesNet, Instances instances)
+			throws Exception {
 		m_BayesNet = bayesNet;
 		bayesNet.setDebug(true);
 		System.out.println("Score before local search: " + calcScore(bayesNet));
 		localSearch(bayesNet, instances);
 		ParentSet[] bestSolution = copyParentSets(bayesNet.getParentSets());
 		double scoreOfBestSolution = calcScore(bayesNet);
-		System.out.println("Score of solution after local search: " + scoreOfBestSolution);
+		System.out.println("Score of solution after local search: "
+				+ scoreOfBestSolution);
 		double scoreOfFoundSolution;
 		int numberOfIterationsWithNoBetterSolution = 0;
 
 		do {
 			perturbate(bayesNet, instances, getNumberOfPertubations());
+			scoreOfFoundSolution = calcScore(bayesNet);
 			localSearch(bayesNet, instances);
[…]

### 2013-01-08 exercise2/paper/paper.tex
changed a lot

--- 2013-01-05 exercise2/paper/paper.tex
@@
 \item comparison and interpretation of the results
-\item description of the used R and Weka features and commands
-\item description of the setup/infrastructure (weka, custom searcher, R, runtime, comparison, data set, ...)
-\item R script that runs the custom searcher
-\item R script that runs a default searcher
-\item find and describe a custom search algorithm we want to uses
 \end{itemize}
 
+%
+%
+%
 \section{Preface}
 
-\paragraph{}This assignment is about extending a Bayes Network Classifier by adding a custom search algorithm.
+This assignment is about extending a Bayes Network Classifier by improving an existing search heuristic in Weka.
+We chose the Hill Climber search algorithm for an implementation of an iterative local search algorithm.
 
+%
+%
+%
 \section{Setup}
 
-Used stuff: R.
-RWeka -- an interface in R for Weka.
-implement a searcher by extending a java class from the weka library/framework.
-running a R script that starts the weka BayesNet-classifier with our custom searcher.
+There is a R script to execute the searcher in the handed in package at \texttt{R/bayesNetTest.R}.
+It loads Weka, builds the classifier, runs it and evaluates the model.
+It also shows the performance tables and the differences of the original Hill Climber implementation and our improved one.
+To run the script the following Tools need to be installed: R with the packages RWeka and rJava, Java (JRE).
 
-\subsection{RWeka}
+The used package RWeka is an interface in R that wraps the Weka framework and has a really good documentation.
 
-RWeka is a interface in R that wraps the Weka framework and has a really good documentation.
+%
+%
+%\section{Weka-Implementation of the Bayesian Network}
+%
+%
+%First we take a look at the BayesNet classifier implementation.
+When the classifier is startet, first thing is to generate the classification model.
+Therefore the BayesNet classifier initializes an adjacency matrix (\texttt{initStructure()}) where each attribute has an array of parents.
+These arrays are empty after initialization.
+%
+%%
+%%\begin{lstlisting}[caption={Initialization of BayesNet},label=initStructure]
+%%...
+%%// initialize parent sets to have arrow from classifier node to
+%%// each of the other nodes
+%%for (int iAttribute = 0; iAttribute < instances.numAttributes(); iAttribute++) {
+%%    if (iAttribute != iClass) {
+%%        bayesNet.getParentSet(iAttribute).addParent(iClass, instances);
+%%    }
+%%}
+%%...
+%%\end{lstlisting}
+%
+%The default Weka's SearchAlgorithm initializes the network like a Naive Bayes network.
+Listing \ref{initStructure} shows how we can use the adjacency matrix of BayesNet.
+Where \texttt{BayesNet} is the classifier that holds the network datastructure and \texttt{getParentSet(n)} returns the \texttt{nth} row in the adjacency matrix.
 
-Some good stuff to know:
+%
+%
+%
+\section{Weka-Implementation of Hill Climber}
 
-\texttt{WOW("at.ac.tuwien.machine\_learning.UltraSearcher")} WOW is a function in R that reads and prints all possible options of a weka class.
-This can be done for classifiers, filters, searchers and all other types in weka.
+We chose to implement an iterated local search which is based on the pre-implemented Hill Climber algorithm in Weka.
+There exist implementations as local score based algorithm and as global score based algorithm.
+We chose the global score based algorithm because this is better suited to iterated local search (for the perturbation and comparison with other solutions).
+The existing Hill Climber algorithm does the following:
 
-\subsection{Weka Javacode}
+\begin{enumerate}
+	\item Create a start solution: One can choose between an empty network which contains no arcs or a Naive Bayes network as start solution.
+In the Naive Bayes network the classification node is the parent of all other attribute nodes.
+	\item Local search: The algorithm chooses always the best solution from its neighborhood. The neighborhood is defined as all solutions which are accessible via the following operations:
+		\begin{itemize}
+			\item Add a new arc.
+			\item Delete an existing arc.
+			\item Reverse an existing arc.
+		\end{itemize}
+		As normal in local search the neighborhood is exploited until a local optimum is reached. The best solution is then returned as the Bayesian network.
+\end{enumerate}
 
-\paragraph{initStructure()} of class BayesNet initializes the Bayesnetwork. It creates an adjacency matrix where each attribute has an array of parents. These arrays are empty after initialization.
+\section{Classification results of the pre-implemented Hill Climber}
 
-\paragraph{} The default SearchAlgorithm.class initializes the network like a Naive Bayes network. This shows how we can use the adjacency matrix of BayesNet:
+\todo{give a table with the results for hill climber with empty start solution as well as with naive bayes.}
 
-\begin{lstlisting}
-...
-// initialize parent sets to have arrow from classifier node to
-// each of the other nodes
-for (int iAttribute = 0; iAttribute < instances.numAttributes(); iAttribute++) {
-    if (iAttribute != iClass) {
-        bayesNet.getParentSet(iAttribute).addParent(iClass, instances);
-    }
-}
-...
-\end{lstlisting}
+\section{Improvement of the Hill Climber: Iterated Local Search}
 
-\paragraph{} Where \texttt{BayesNet} is the classifier that holds the network datastructure and \texttt{getParentSet(n)} returns the \texttt{nth} row in the adjacency matrix.
+We improved the existing Hill Climber algorithm by making an iterated local search out of it. The basic functionality is as follows:
+
+\begin{itemize}
+	\item Create a start solution.
+	\item Apply local search to the start solution. The solution obtained is the new best solution.
+	\item Repeat the following procedure until there is no new best solution for five iterations:
+		\begin{itemize}
+			\item Perturbate the current best solution.
+			\item Apply local search to the perturbated solution.
+			\item If the solution obtained after local search is better than the current best solution then this solution is the new best solution.
+		\end{itemize}
+\end{itemize}
+
+It remains to describe what the Perturbation-procedure does: The perturbation-procedure modifies the given net by applying some number of random moves in the neighborhood.
+We implemented the search procedure in a way such that it is possible for the user to choose the number of random moves which are applied in every perturbation step.
+The standard value of random moves is three.
+Since it is not always possible to apply a move there are some tests to make sure that a random move is possible.
+This is necessary in the following cases:
+
+\begin{itemize}
+	\item The network doesn't contain an arc. (Deletion of an arc is not allowed/possible.)
+	\item Every node has a maximum number of parents. (Addition of an arc is not allowed/possible.)
+	\item Addition of an arc creates a cycle. (This test is pre-implemented in Weka.
+	\item etc.
+\end{itemize}
+
+\section{Classification results of the iterated local search}
+
+\todo{Insert tables containing the classification results of the iterated local search}.
+
+\section{Comparison of the Hill Climber algorithm and the iterated local search}
+
+One can see that the iterated local search does not or in some cases just slightly perform better than the pre-implemented Hill-Climber algorithm.
+This is mostly due to the fact that for our test data sets the Naive Bayes representation yields the best or at least nearly the best result.
+The Hill climber doesn't perform many neighborhood operations.
+For the contraception data set there are no neighborhood operations at all and for the adult data set its just a few operations (about five operations).
+The perturbation normally leads to worse solutions and local search then finds a net which is close to Naive Bayes or pretty similar to it so we don't get a better solution than the solution after the local search.
+
+%
+%
+%\subsection{Description}
+%
+%As shown in listing \ref{hcSearch}, Hill Climber calculates an optimal solution based on the current network iteratively until no opration can be found to improve the network or the new network is only slightly better than the previous one.
+To find the best operation all possible operations get evaluated and the best one is returned.
+The algorithm stops, when it finds a local optima.
+%
+%\begin{lstlisting}[caption={HillClimbers search algorithm},label=hcSearch]
+%double fScore = calcScore(bayesNet);
+%Operation oOperation = getOptimalOperation(bayesNet, instances);
+%while ((oOperation != null) && (oOperation.m_fScore > fScore)) {
+%	performOperation(bayesNet, instances, oOperation);
+%	fScore = oOperation.m_fScore;
+%	oOperation = getOptimalOperation(bayesNet, instances);
+%}  
+%\end{lstlisting}
+
+%
+%
+%\subsection{Test Results}
+%
+%Two data sets are used to test the search algorithm. These are x and y from the assignment 1.
+%
+%\todo{insert results of test}
+%
+%As shown in the result listing, \todo{describe result} ....
+
+%
+%
+
+%\subsection{Description}
+%
+%Like the Hill Climber, our improved search algorithm introduces an initial Naive Bayes like network where each attribute has the classification attribute as parent.
+%
+%Then the Hill Climber algorithm is executed to find a local optima.
+To find other local optima, our search algorithm iteratively selects a neighbor of the found network and uses it as initial network for the Hill Climber to hopefully find better local optima than the first one.
+%
+% The intial network we use in each iteration is a random neighbor of the previously found one. It does not matter if it is better or worse.
+%
+%If the Hill Climber finds a better network it is stored as the current best found network and if the seacher cannot find a better network after a given number of iterations, the algorithm stops and yields the best solution found.
+
+%
+%
+%%
+%\section{Comparison}
+%
+%TODO:
+%* used datasets
+%* accuracy results
+%* runtime/performance
 

### 2013-01-08 exercise2/paper/paper.tex
handed in

@@
 %
-\section{TODO}
-
-\begin{itemize}
-\item Implementation of cutsom bn-searcher
-\item documentation of cutsom bn-searcher
-\item comparison and interpretation of the results
-\end{itemize}
-
-%
-%
-%
 \section{Preface}
@@
 
+\newpage
+
 \section{Classification results of the pre-implemented Hill Climber}
 
-\todo{give a table with the results for hill climber with empty start solution as well as with naive bayes.}
+\begin{lstlisting}[caption={Classification result with default Hill Climber},label=listingDefault]
+"ADULT DATA SET"
+"Original Hill Climber classifier"
+              pctCorrect             pctIncorrect          pctUnclassified 
+              84.7025583               15.2974417                0.0000000 
+                   kappa        meanAbsoluteError     rootMeanSquaredError 
+               0.6078686                0.1722236                0.3319906 
+   relativeAbsoluteError rootRelativeSquaredError 
+              47.1008155               77.6450216 
+\end{lstlisting}
 
@@
 
-\todo{Insert tables containing the classification results of the iterated local search}.
+\begin{lstlisting}[caption={Classification result with iterated local search and Hill Climber},label=listingExtended]
+"ADULT DATA SET"
+"New iterated local search classifier"
+              pctCorrect             pctIncorrect          pctUnclassified 
+              84.8376893               15.1623107                0.0000000 
+                   kappa        meanAbsoluteError     rootMeanSquaredError 
+               0.6114594                0.1711662                0.3298876 
+   relativeAbsoluteError rootRelativeSquaredError 
+              46.8116196               77.1531638
+
+\end{lstlisting}
 
@@
 
-One can see that the iterated local search does not or in some cases just slightly perform better than the pre-implemented Hill-Climber algorithm.
-This is mostly due to the fact that for our test data sets the Naive Bayes representation yields the best or at least nearly the best result.
-The Hill climber doesn't perform many neighborhood operations.
-For the contraception data set there are no neighborhood operations at all and for the adult data set its just a few operations (about five operations).
+One can see that the iterated local search does improve the results slightly the pre-implemented Hill-Climber algorithm.
 The perturbation normally leads to worse solutions and local search then finds a net which is close to Naive Bayes or pretty similar to it so we don't get a better solution than the solution after the local search.
 
-%
-%
-%\subsection{Description}
-%
-%As shown in listing \ref{hcSearch}, Hill Climber calculates an optimal solution based on the current network iteratively until no opration can be found to improve the network or the new network is only slightly better than the previous one.
-To find the best operation all possible operations get evaluated and the best one is returned.
-The algorithm stops, when it finds a local optima.
-%
-%\begin{lstlisting}[caption={HillClimbers search algorithm},label=hcSearch]
-%double fScore = calcScore(bayesNet);
-%Operation oOperation = getOptimalOperation(bayesNet, instances);
-%while ((oOperation != null) && (oOperation.m_fScore > fScore)) {
-%	performOperation(bayesNet, instances, oOperation);
-%	fScore = oOperation.m_fScore;
-%	oOperation = getOptimalOperation(bayesNet, instances);
-%}  
-%\end{lstlisting}
+\begin{lstlisting}[caption={Comparison of classification results},label=listingComparison]
+"Difference of details"
+              pctCorrect             pctIncorrect          pctUnclassified 
+             0.135130985             -0.135130985              0.000000000 
+                   kappa        meanAbsoluteError     rootMeanSquaredError 
+             0.003590769             -0.001057442             -0.002103060 
+   relativeAbsoluteError rootRelativeSquaredError 
+            -0.289195892             -0.491857758 
+"Difference of confusion matrices"
+       predicted
+        >50K <=50K
+  >50K    26   -26
+  <=50K  -18    18
 
-%
-%
-%\subsection{Test Results}
-%
-%Two data sets are used to test the search algorithm. These are x and y from the assignment 1.
-%
-%\todo{insert results of test}
-%
-%As shown in the result listing, \todo{describe result} ....
-
-%
-%
-
-%\subsection{Description}
-%
-%Like the Hill Climber, our improved search algorithm introduces an initial Naive Bayes like network where each attribute has the classification attribute as parent.
-%
-%Then the Hill Climber algorithm is executed to find a local optima.
-To find other local optima, our search algorithm iteratively selects a neighbor of the found network and uses it as initial network for the Hill Climber to hopefully find better local optima than the first one.
-%
-% The intial network we use in each iteration is a random neighbor of the previously found one. It does not matter if it is better or worse.
-%
-%If the Hill Climber finds a better network it is stored as the current best found network and if the seacher cannot find a better network after a given number of iterations, the algorithm stops and yields the best solution found.
-
-%
-%
-%%
-%\section{Comparison}
-%
-%TODO:
-%* used datasets
-%* accuracy results
-%* runtime/performance
+\end{lstlisting}
 

### 2013-01-08 exercise2/assignment2-kiesl-steinwender-schaefer/R/bayesNetTest.R

@@ -6,16 +6,10 @@ require("rJava")
 
 bayesNet <- make_Weka_classifier("weka/classifiers/bayes/BayesNet")
 
-cmc = read.arff("../../exercise1/samples/cmc_preprocessed.arff")
-adult =  read.arff("../../exercise1/samples/census/adult_preprocessed.arff")
+adult =  read.arff("adult_preprocessed.arff")
 
-#hcModel = bayesNet(`Species`~., data=iris, control = Weka_control(D="",  Q = "weka.classifiers.bayes.net.search.global.HillClimber", "--", P=1, S="BAYES", E="weka.classifiers.bayes.net.estimate.SimpleEstimator", "--", A=0.5))
 hcAdultModel = bayesNet(`income`~., data=adult, control = Weka_control(D="",  Q = "weka.classifiers.bayes.net.search.global.HillClimber", "--", P=1, S="BAYES", E="weka.classifiers.bayes.net.estimate.SimpleEstimator", "--", A=0.5))
 extendedAdultModel = bayesNet(`income`~., data=adult, control = Weka_control(D="",  Q = "weka.classifiers.bayes.net.search.global.IteratedLocalSearch", "--", T=5, P=1, S="BAYES", E="weka.classifiers.bayes.net.estimate.SimpleEstimator", "--", A=0.5))
-
-#hcCmcModel = bayesNet(`income`~., data=adult, control = Weka_control(D="",  Q = "weka.classifiers.bayes.net.search.global.HillClimber", "--", P=1, S="BAYES", E="weka.classifiers.bayes.net.estimate.SimpleEstimator", "--", A=0.5))
-#extendedCmcModel = bayesNet(`Species`~., data=iris, control = Weka_control(D="",  Q = "weka.classifiers.bayes.net.search.global.IteratedLocalSearch", "--", T=5, P=1, S="BAYES", E="weka.classifiers.bayes.net.estimate.SimpleEstimator", "--", A=0.5))
-
 
 "ADULT DATA SET"
 hcAdultEvaluation = evaluate_Weka_classifier(hcAdultModel)

### 2013-01-24 exercise3/knnDemonstration.R
provided new knnDemonstration script, works quite well it     builds the model on the one half of our skin segmentation arff     and tests on the other half split     then plots each classified instance and colors it depending whether     the instance was classified correct or not

@@ -1,0 +1,43 @@
+#!/usr/bin/Rscript
+
+# Weka bridge
+library(RWeka)
+
+# plotting library
+library(rgl)
+
+# library to wait for user input
+library(tcltk)
+
+Nth.delete<-function(dataframe, n)dataframe[(seq(n,to=nrow(dataframe),by=n)),]
+
+# splitdf function will return a list of training and testing sets
+splitdf <- function(dataframe, seed=NULL) {
+    if (!is.null(seed)) set.seed(seed)
+    index <- 1:nrow(dataframe)
+    trainindex <- sample(index, trunc(length(index)/2))
+    trainset <- dataframe[trainindex, ]
+    testset <- dataframe[-trainindex, ]
+    list(trainset=trainset,testset=testset)
+}
+
+skin_arff <- read.arff("../exercise1/samples/skin\ color/Skin_NonSkin.arff")
+# We choose every 1000 row as an instance for our sample set
+sample <- Nth.delete(skin_arff,100)
+splits <- splitdf(sample, seed=12345)
+training <- splits$trainset
+testing <- splits$testset
+
+classifier <-IBk(Y~., training, control = Weka_control(K = 20))
+predictedData <- testing
+predictedData$P <- predict(classifier, testing)
+predictedData$Color <- (predictedData$Y == predictedData$P)
+str(predictedData)
+plot3d(predictedData$R, predictedData$G, predictedData$B, col=predictedData$Color, size=2, type='s')
+tk_messageBox(message = "Press a key to exit")
+
+#evaluation <- evaluate_Weka_classifier(classifier)
+#evaluation$confusionMatrix
+
+
+

### 2013-01-25 exercise3/knnDemonstration.R
refactored useless function

@@ -9,24 +9,18 @@ library(rgl)
 # library to wait for user input
 library(tcltk)
 
-Nth.delete<-function(dataframe, n)dataframe[(seq(n,to=nrow(dataframe),by=n)),]
+Nth.rows<-function(dataframe, n)dataframe[(seq(n,to=nrow(dataframe),by=n)),]
 
-# splitdf function will return a list of training and testing sets
-splitdf <- function(dataframe, seed=NULL) {
-    if (!is.null(seed)) set.seed(seed)
-    index <- 1:nrow(dataframe)
-    trainindex <- sample(index, trunc(length(index)/2))
-    trainset <- dataframe[trainindex, ]
-    testset <- dataframe[-trainindex, ]
-    list(trainset=trainset,testset=testset)
-}
 
 skin_arff <- read.arff("../exercise1/samples/skin\ color/Skin_NonSkin.arff")
 # We choose every 1000 row as an instance for our sample set
-sample <- Nth.delete(skin_arff,100)
-splits <- splitdf(sample, seed=12345)
+sample <-Nth.rows(skin_arff,100)
+splits <- split(sample, c("trainset", "testset"))
+
 training <- splits$trainset
 testing <- splits$testset
+nrow(training)
+nrow(testing)
 
 classifier <-IBk(Y~., training, control = Weka_control(K = 20))
 predictedData <- testing

### 2013-01-25 exercise3/knnDemonstration.R
knnDemonstration script has become a useful command line tool

@@ -1,5 +1,27 @@
 #!/usr/bin/Rscript
 
+# library for option parsing
+library("optparse")
+
+option_list <- list(
+                    make_option(c("-t", "--train"), help="load training data", metavar="ARFF_FILE"),
+                    make_option(c("-T", "--test"), help="load test data, if ommitted, training data is splitted into parts", metavar="ARFF_FILE"),
+                    make_option(c("-k", "--kneighbors"), type="integer", help="k parameter for nearest neighbor classifier", default=as.integer(5), metavar="K"),
+                    make_option(c("--extract"), type="integer", help="reduce input data set by extracting every nth instance", metavar="N")
+                    )
+parser <- OptionParser(usage="knnDemonstration [option] trainingfile", option_list=option_list)
+arguments <- parse_args(parser, positional_arguments = TRUE)
+opt <- arguments$options
+
+if(is.null(opt$train)){
+    cat("Please provide some training data.. \n")
+    stop()
+}
+if(file.access(opt$train) == -1){
+    stop(sprintf("Specified file ( $s ) does not exist", opt$input))
+}
+
+
 # Weka bridge
 library(RWeka)
 
@@ -9,24 +31,27 @@ library(rgl)
 # library to wait for user input
 library(tcltk)
 
+
 Nth.rows<-function(dataframe, n)dataframe[(seq(n,to=nrow(dataframe),by=n)),]
 
 
-skin_arff <- read.arff("../exercise1/samples/skin\ color/Skin_NonSkin.arff")
-# We choose every 1000 row as an instance for our sample set
-sample <-Nth.rows(skin_arff,100)
+skin_arff <- read.arff(opt$train)
+# If the user specified --extract, we choose every nth row as an instance for our sample set
+if (!(is.null(opt$extract))){
+    sample <-Nth.rows(skin_arff,opt$extract)
+} else {
+    sample <- skin_arff
+}
 splits <- split(sample, c("trainset", "testset"))
 
 training <- splits$trainset
+# TODO: provide logic for test data via command line
 testing <- splits$testset
-nrow(training)
-nrow(testing)
 
-classifier <-IBk(Y~., training, control = Weka_control(K = 20))
+classifier <-IBk(Y~., training, control = Weka_control(K = opt$k))
 predictedData <- testing
 predictedData$P <- predict(classifier, testing)
 predictedData$Color <- (predictedData$Y == predictedData$P)
-str(predictedData)
[…]

### 2013-01-25 exercise3/knnDemonstration.R
tiny commit, just to inform on missing implementation in help function

@@ -5,7 +5,7 @@ library("optparse")
 
 option_list <- list(
                     make_option(c("-t", "--train"), help="load training data", metavar="ARFF_FILE"),
-                    make_option(c("-T", "--test"), help="load test data, if ommitted, training data is splitted into parts", metavar="ARFF_FILE"),
+                    make_option(c("-T", "--test"), help="NOT IMPLEMENTED YET", metavar="ARFF_FILE"),
                     make_option(c("-k", "--kneighbors"), type="integer", help="k parameter for nearest neighbor classifier", default=as.integer(5), metavar="K"),
                     make_option(c("--extract"), type="integer", help="reduce input data set by extracting every nth instance", metavar="N")
                     )

### 2013-01-25 exercise3/knnDemonstration.R
new axis description providing test data should work

@@ -5,7 +5,7 @@ library("optparse")
 
 option_list <- list(
                     make_option(c("-t", "--train"), help="load training data", metavar="ARFF_FILE"),
-                    make_option(c("-T", "--test"), help="NOT IMPLEMENTED YET", metavar="ARFF_FILE"),
+                    make_option(c("-T", "--test"), help="load test data, if ommitted, training data is splitted into parts", metavar="ARFF_FILE"),
                     make_option(c("-k", "--kneighbors"), type="integer", help="k parameter for nearest neighbor classifier", default=as.integer(5), metavar="K"),
                     make_option(c("--extract"), type="integer", help="reduce input data set by extracting every nth instance", metavar="N")
                     )
@@ -13,14 +13,19 @@ parser <- OptionParser(usage="knnDemonstration [option] trainingfile", option_li
 arguments <- parse_args(parser, positional_arguments = TRUE)
 opt <- arguments$options
 
+# handle training data
 if(is.null(opt$train)){
     cat("Please provide some training data.. \n")
     stop()
 }
 if(file.access(opt$train) == -1){
-    stop(sprintf("Specified file ( $s ) does not exist", opt$input))
+    stop(sprintf("Specified file ( %s ) does not exist", opt$input))
 }
 
+# handle test data
+if( !is.null(opt$test) && (file.access(opt$test) == -1)){
+    stop(sprintf("Specified file ( %s ) does not exist", opt$test))
+}
 
 # Weka bridge
 library(RWeka)
@@ -31,29 +36,39 @@ library(rgl)
 # library to wait for user input
 library(tcltk)
 
-
+# helper function to extract a fraction amount of the training data
 Nth.rows<-function(dataframe, n)dataframe[(seq(n,to=nrow(dataframe),by=n)),]
 
 
-skin_arff <- read.arff(opt$train)
+inputArff <- read.arff(opt$train)
 # If the user specified --extract, we choose every nth row as an instance for our sample set
 if (!(is.null(opt$extract))){
-    sample <-Nth.rows(skin_arff,opt$extract)
+    sample <-Nth.rows(inputArff, opt$extract)
 } else {
-    sample <- skin_arff
+    sample <- inputArff
 }
-splits <- split(sample, c("trainset", "testset"))
 
-training <- splits$trainset
-# TODO: provide logic for test data via command line
-testing <- splits$testset
[…]

### 2013-01-25 exercise3/knnDemonstration.R
Rscript interactively asks for a user for a k value!

@@ -33,8 +33,8 @@ library(RWeka)
 # plotting library
 library(rgl)
 
-# library to wait for user input
-library(tcltk)
+# library for the text box
+library(rpanel)
 
 # helper function to extract a fraction amount of the training data
 Nth.rows<-function(dataframe, n)dataframe[(seq(n,to=nrow(dataframe),by=n)),]
@@ -59,19 +59,19 @@ if (is.null(opt$test)){
 }
 
 
-classifier <-IBk(Y~., training, control = Weka_control(K = opt$k))
-predictedData <- testing
-predictedData$P <- predict(classifier, testing)
-predictedData$Color <- (predictedData$Y == predictedData$P)
-plot3d(predictedData$R, predictedData$G, predictedData$B, xlab="B", ylab="G", zlab="R", col=predictedData$Color, size=2, type='s')
-tk_messageBox(message = "Press a key to exit")
 
-
-
-
-
-#evaluation <- evaluate_Weka_classifier(classifier)
-#evaluation$confusionMatrix
-
-
-
+interface.draw <- function(panel) {
+    classifier <-IBk(Y~., training, control = Weka_control(K = as.numeric(panel$k)))
+    predictedData <- testing
+    predictedData$P <- predict(classifier, testing)
+    predictedData$Color <- (predictedData$Y == predictedData$P)
+    # TODO: maybe output some evaluation data? e.g. percentage of misclassified instances
+    plot3d(predictedData$R, predictedData$G, predictedData$B, xlab="B", ylab="G", zlab="R", col=predictedData$Color, size=1, type='s')
+    panel
+}
+interface.panel <- rp.control("K nearest neighbor", k=opt$k)
+rp.textentry(panel=interface.panel, var=k, action=interface.draw, names="K nearest neighbor", labels="K=")
+# run it!
+rp.do(interface.panel, interface.draw)
+# user must close the panel to exit the program
+rp.block(interface.panel)

### 2013-01-25 exercise3/knnDemonstration.R
added column check for input data

@@ -58,6 +58,10 @@ if (is.null(opt$test)){
     testing <- read.arff(opt$test)
 }
 
+if( ncol(training) < 3 || 4 < ncol(training)){
+    stop(sprintf("The provided training has got %s columns, whith the last column considered as class attribute, but we can only plot 2dimensional or 3dimensional data!",  ncol(training)))
+}
+
 
 
 interface.draw <- function(panel) {

### 2013-01-25 exercise3/knnDemonstration.R
added column check for training and test data

@@ -62,6 +62,9 @@ if( ncol(training) < 3 || 4 < ncol(training)){
     stop(sprintf("The provided training has got %s columns, whith the last column considered as class attribute, but we can only plot 2dimensional or 3dimensional data!",  ncol(training)))
 }
 
+if( ncol(testing) != ncol(training)){
+    stop(sprintf("The number of columns for the provided data sets is not the same: %s columns for trainingdata and %s columns for testdata.",  ncol(training), ncol(testing)))
+}
 
 
 interface.draw <- function(panel) {

### 2013-01-25 exercise3/kNN/src/at/ac/tuwien/knn/data/DataSetParser.java
Added a java version of exercise 3

@@ -1,0 +1,43 @@
+package at.ac.tuwien.knn.data;
+import java.io.BufferedReader;
+import java.io.DataInputStream;
+import java.io.File;
+import java.io.FileInputStream;
+import java.io.FileNotFoundException;
+import java.io.IOException;
+import java.io.InputStreamReader;
+import java.util.Collection;
+import java.util.LinkedList;
+
+public class DataSetParser {
+	public static DataSet parseFile(File dataFile) {
+		Collection<DataPoint> datapoints = new LinkedList<DataPoint>();
+
+		BufferedReader fileReader;
+		try {
+			fileReader = new BufferedReader(new InputStreamReader(
+					new DataInputStream(new FileInputStream(dataFile))));
+			String line;
+			boolean dataLines = false;
+			int i=0;
+			while ((line = fileReader.readLine()) != null && i < 10) {
+				if(dataLines && !line.startsWith("//")){
+					String[] columns = line.split(",");
+					double x = Double.valueOf(columns[0]);
+					double y = Double.valueOf(columns[1]);
+					int classification = Integer.valueOf(columns[2]);
+					datapoints.add(new DataPoint(x,y,classification));
+				}
+				if(line.startsWith("@data")){
+					dataLines = true;
+				}
+			}
+			fileReader.close();
+		} catch (FileNotFoundException e) {
+			e.printStackTrace();
+		} catch (IOException e) {
+			e.printStackTrace();
+		}
+		return new DataSet(datapoints);
+	}
+}

### 2013-01-25 exercise3/kNN/src/at/ac/tuwien/knn/gui/DrawingArea.java

@@ -1,0 +1,173 @@
+package at.ac.tuwien.knn.gui;
+
+import java.awt.Color;
+import java.awt.Graphics;
+import java.awt.Graphics2D;
+import java.awt.Point;
+import java.awt.RenderingHints;
+import java.util.Collection;
+import java.util.HashMap;
+import java.util.LinkedList;
+import java.util.Random;
+
+import javax.swing.JPanel;
+
+import at.ac.tuwien.knn.data.DataPoint;
+import at.ac.tuwien.knn.data.DataSet;
+
+/**
+ * The panel in which all the data points are drawn
+ * @author BK
+ *
+ */
+public class DrawingArea extends JPanel {
+	private DataSet dataSet;
+	private double minX, maxX, minY, maxY;
+	private boolean showTrainingData = true;
+	private boolean showTestData = false;
+	
+	private static Color[] colors = new Color[]{new Color(230,0,0),//red
+												new Color(0,0,230),//blue
+												new Color(0,220,0),//green
+												new Color(102,0,102),//violet
+												new Color(255,153,0),//orange
+												new Color(255,255,100)//beige
+												};
+	private HashMap<Integer,Color> classColorMap = new HashMap<Integer, Color>();
+	private int pointRadius = 3;
+
+	@Override
+	protected void paintComponent(Graphics graphics) {
+		super.paintComponent(graphics);
+		if(this.dataSet == null || this.dataSet.getAllDataPoints() == null) return;
+		this.initColorMap(this.dataSet.getTrainingSet());
+		this.setMinMaxCoordinates(this.dataSet.getAllDataPoints());
+		if(showTrainingData){
+			this.paintDataPoints((Graphics2D)graphics, this.dataSet.getTrainingSet());
+		}
+		if(showTestData){
+			this.paintTestSet((Graphics2D)graphics, this.dataSet.getTestSet());
+		}
+	}
+	
+	/**
+	 * Paints the test data points into the drawing area. For every test-data point we draw a grey circle around it. At this 
+	 * time the circle has just an arbitrary hardcoded radius but in the future the circle should be exactly as big such that
+	 * it contains for every test data point the k nearest neighbours of this data point.
+	 * @param g The graphics element in which we draw.
+	 * @param dataPoints The datapoints which should be drawn.
+	 */
[…]

### 2013-01-25 exercise3/kNN/src/at/ac/tuwien/knn/gui/MainWindow.java

@@ -1,0 +1,144 @@
+package at.ac.tuwien.knn.gui;
+
+import java.awt.Color;
+import java.awt.EventQueue;
+import java.awt.SystemColor;
+import java.awt.event.ItemEvent;
+import java.awt.event.ItemListener;
+import java.awt.event.MouseAdapter;
+import java.awt.event.MouseEvent;
+
+import javax.swing.JButton;
+import javax.swing.JCheckBox;
+import javax.swing.JFileChooser;
+import javax.swing.JFrame;
+import javax.swing.JLabel;
+import javax.swing.JSpinner;
+import javax.swing.border.LineBorder;
+import javax.swing.event.ChangeEvent;
+import javax.swing.event.ChangeListener;
+
+import at.ac.tuwien.knn.data.DataSet;
+import at.ac.tuwien.knn.data.DataSetParser;
+
+
+public class MainWindow {
+
+	private JFrame frame;
+	private DrawingArea panel;
+	private JLabel lbFilename;
+
+	/**
+	 * Launch the application.
+	 */
+	public static void main(String[] args) {
+		EventQueue.invokeLater(new Runnable() {
+			public void run() {
+				try {
+					MainWindow window = new MainWindow();
+					window.frame.setVisible(true);
+				} catch (Exception e) {
+					e.printStackTrace();
+				}
+			}
+		});
+		
+		//DrawingArea drawingArea = new DrawingArea();
+	}
+
+	/**
+	 * Create the application.
+	 */
+	public MainWindow() {
+		initialize();
+		
+	}
+
+	/**
+	 * Initialize the contents of the frame.
+	 */
+	private void initialize() {
+		frame = new JFrame();
+		frame.setBounds(100, 100, 1000, 800);
+		frame.setDefaultCloseOperation(JFrame.EXIT_ON_CLOSE);
+		frame.getContentPane().setLayout(null);
+		
+		JSpinner spinnerK = new JSpinner();
+		spinnerK.setBounds(827, 117, 45, 20);
+		frame.getContentPane().add(spinnerK);
+		
+		JLabel lblK = new JLabel("k:");
+		lblK.setBounds(810, 120, 16, 14);
+		frame.getContentPane().add(lblK);
+		
+		panel = new DrawingArea();
+		panel.setBackground(Color.WHITE);
+		panel.setBorder(new LineBorder(SystemColor.controlDkShadow));
+		panel.setBounds(33, 33, 737, 699);
[…]

### 2013-01-26 exercise3/knnDemonstration.R
the script should be independent of the naming of columns in the input files

@@ -67,13 +67,19 @@ if( ncol(testing) != ncol(training)){
 }
 
 
+# for convenience, rename class column to "Class"
+names(training)[4] <- "Class"
+names(testing)[4] <- "Class"
+
 interface.draw <- function(panel) {
-    classifier <-IBk(Y~., training, control = Weka_control(K = as.numeric(panel$k)))
-    predictedData <- testing
-    predictedData$P <- predict(classifier, testing)
-    predictedData$Color <- (predictedData$Y == predictedData$P)
+    classifier <-IBk(Class ~ ., training, control = Weka_control(K = as.numeric(panel$k)))
+    testing <- testing
+    testing$P <- predict(classifier, testing)
+    testing$Color <- (testing$Class == testing$P)
+
     # TODO: maybe output some evaluation data? e.g. percentage of misclassified instances
-    plot3d(predictedData$R, predictedData$G, predictedData$B, xlab="B", ylab="G", zlab="R", col=predictedData$Color, size=1, type='s')
+
+    plot3d(testing[[1]], testing[[2]], testing[[3]], xlab=names(testing)[1], ylab=names(testing)[2], zlab=names(testing)[3], col=testing$Color, size=1, type='s')
     panel
 }
 interface.panel <- rp.control("K nearest neighbor", k=opt$k)

### 2013-01-26 exercise3/knnDemonstration.R
silly 2d vs 3d dimensional switch, handling 2d data currently raises a NotImplementedYetError

@@ -58,7 +58,11 @@ if (is.null(opt$test)){
     testing <- read.arff(opt$test)
 }
 
-if( ncol(training) < 3 || 4 < ncol(training)){
+if ( ncol(training) == 3 ){
+    modus = "2D"
+} else if ( ncol(training) == 4 ){
+    modus = "3D"
+} else {
     stop(sprintf("The provided training has got %s columns, whith the last column considered as class attribute, but we can only plot 2dimensional or 3dimensional data!",  ncol(training)))
 }
 
@@ -66,25 +70,28 @@ if( ncol(testing) != ncol(training)){
     stop(sprintf("The number of columns for the provided data sets is not the same: %s columns for trainingdata and %s columns for testdata.",  ncol(training), ncol(testing)))
 }
 
+if (modus == "2D"){
+    stop("Handling of 2 dimennsional plots is not implemented yet!")
+} else if (modus == "3D"){
+    # for convenience, rename class column to "Class"
+    names(training)[4] <- "Class"
+    names(testing)[4] <- "Class"
 
-# for convenience, rename class column to "Class"
-names(training)[4] <- "Class"
-names(testing)[4] <- "Class"
+    interface.draw <- function(panel) {
+        classifier <-IBk(Class ~ ., training, control = Weka_control(K = as.numeric(panel$k)))
+        testing <- testing
+        testing$P <- predict(classifier, testing)
+        testing$Color <- (testing$Class == testing$P)
 
-interface.draw <- function(panel) {
-    classifier <-IBk(Class ~ ., training, control = Weka_control(K = as.numeric(panel$k)))
-    testing <- testing
-    testing$P <- predict(classifier, testing)
-    testing$Color <- (testing$Class == testing$P)
+        # TODO: maybe output some evaluation data? e.g. percentage of misclassified instances
 
-    # TODO: maybe output some evaluation data? e.g. percentage of misclassified instances
-
-    plot3d(testing[[1]], testing[[2]], testing[[3]], xlab=names(testing)[1], ylab=names(testing)[2], zlab=names(testing)[3], col=testing$Color, size=1, type='s')
-    panel
+        plot3d(testing[[1]], testing[[2]], testing[[3]], xlab=names(testing)[1], ylab=names(testing)[2], zlab=names(testing)[3], col=testing$Color, size=1, type='s')
+        panel
+    }
+    interface.panel <- rp.control("K nearest neighbor", k=opt$k)
[…]

### 2013-01-26 exercise3/kNN/src/at/ac/tuwien/knn/data/DataSet.java
did some stuff

@@ -1,80 +1,0 @@
-package at.ac.tuwien.knn.data;
-import java.util.Collection;
-import java.util.LinkedList;
-
-/**
- * The dataset stores the data points. It contains two separate collections which represent the training and the test data.
- * @author BK
- *
- */
-public class DataSet {
-	Collection<DataPoint> dataPoints;
-	Collection<DataPoint> trainingSet = new LinkedList<DataPoint>();
-	Collection<DataPoint> testSet = new LinkedList<DataPoint>();
-	
-	int trainingSetPercentage = 80;
-	
-	public DataSet(){}
-	
-	public DataSet(Collection<DataPoint> dataPoints){
-		this.dataPoints = dataPoints;
-	}
-	
-	public Collection<DataPoint> getTrainingSet() {
-		return trainingSet;
-	}
-
-	public void setTrainingSet(Collection<DataPoint> trainingSet) {
-		this.trainingSet = trainingSet;
-	}
-
-	public Collection<DataPoint> getTestSet() {
-		return testSet;
-	}
-
-	public void setTestSet(Collection<DataPoint> testSet) {
-		this.testSet = testSet;
-	}
-	
-	public Collection<DataPoint> getAllDataPoints() {
-		return this.dataPoints;
-	}
-
-	public void setAllDataPoints(Collection<DataPoint> datapoints) {
-		this.dataPoints = datapoints;
-	}
-
-	/**
-	 * Sets the percentage of data which should be used as training set.
-	 * @param percentage
-	 */
-	public void setTrainingSetPercentage(int percentage){
-		if(percentage > 100){
-			percentage = 100;
-		}else if(percentage < 0){
-			percentage = 0;
-		}
-		this.trainingSetPercentage = percentage;
-		this.split(percentage);
-	}
-	
-	/**
-	 * Splits the data into training and test data.
-	 * @param percentage
-	 */
-	public void split(int percentage){
-		double step = 100/(double)percentage;
-		double i=1;
-		for(DataPoint dataPoint : this.dataPoints){
-			if(i >= step){
-				i -= step;
-				this.trainingSet.add(dataPoint);
-			}else{
-				this.testSet.add(dataPoint);
-			}
-			i++;
-		}
-		int realPercentage = (int)(100 * this.trainingSet.size() / (double)this.dataPoints.size());
-		System.out.println("Percentage: " + realPercentage);
-	}
-}

### 2013-01-26 exercise3/kNN/src/at/ac/tuwien/knn/data/DataSetParser.java

@@ -1,43 +1,0 @@
-package at.ac.tuwien.knn.data;
-import java.io.BufferedReader;
-import java.io.DataInputStream;
-import java.io.File;
-import java.io.FileInputStream;
-import java.io.FileNotFoundException;
-import java.io.IOException;
-import java.io.InputStreamReader;
-import java.util.Collection;
-import java.util.LinkedList;
-
-public class DataSetParser {
-	public static DataSet parseFile(File dataFile) {
-		Collection<DataPoint> datapoints = new LinkedList<DataPoint>();
-
-		BufferedReader fileReader;
-		try {
-			fileReader = new BufferedReader(new InputStreamReader(
-					new DataInputStream(new FileInputStream(dataFile))));
-			String line;
-			boolean dataLines = false;
-			int i=0;
-			while ((line = fileReader.readLine()) != null && i < 10) {
-				if(dataLines && !line.startsWith("//")){
-					String[] columns = line.split(",");
-					double x = Double.valueOf(columns[0]);
-					double y = Double.valueOf(columns[1]);
-					int classification = Integer.valueOf(columns[2]);
-					datapoints.add(new DataPoint(x,y,classification));
-				}
-				if(line.startsWith("@data")){
-					dataLines = true;
-				}
-			}
-			fileReader.close();
-		} catch (FileNotFoundException e) {
-			e.printStackTrace();
-		} catch (IOException e) {
-			e.printStackTrace();
-		}
-		return new DataSet(datapoints);
-	}
-}

### 2013-01-26 exercise3/kNN/src/at/ac/tuwien/knn/gui/DrawingArea.java

@@ -1,173 +1,246 @@
 package at.ac.tuwien.knn.gui;
 
-import java.awt.Color;
-import java.awt.Graphics;
-import java.awt.Graphics2D;
-import java.awt.Point;
-import java.awt.RenderingHints;
-import java.util.Collection;
+import at.ac.tuwien.knn.data.DataSets;
+import at.ac.tuwien.knn.weka.WekaApi;
+import weka.core.Instance;
+import weka.core.Instances;
+
+import javax.swing.*;
+import java.awt.*;
+import java.io.File;
 import java.util.HashMap;
-import java.util.LinkedList;
+import java.util.Map;
 import java.util.Random;
-
-import javax.swing.JPanel;
-
-import at.ac.tuwien.knn.data.DataPoint;
-import at.ac.tuwien.knn.data.DataSet;
 
 /**
  * The panel in which all the data points are drawn
- * @author BK
  *
+ * @author BK
  */
 public class DrawingArea extends JPanel {
-	private DataSet dataSet;
-	private double minX, maxX, minY, maxY;
-	private boolean showTrainingData = true;
-	private boolean showTestData = false;
-	
-	private static Color[] colors = new Color[]{new Color(230,0,0),//red
-												new Color(0,0,230),//blue
-												new Color(0,220,0),//green
-												new Color(102,0,102),//violet
-												new Color(255,153,0),//orange
-												new Color(255,255,100)//beige
-												};
-	private HashMap<Integer,Color> classColorMap = new HashMap<Integer, Color>();
-	private int pointRadius = 3;
 
-	@Override
-	protected void paintComponent(Graphics graphics) {
-		super.paintComponent(graphics);
-		if(this.dataSet == null || this.dataSet.getAllDataPoints() == null) return;
-		this.initColorMap(this.dataSet.getTrainingSet());
-		this.setMinMaxCoordinates(this.dataSet.getAllDataPoints());
-		if(showTrainingData){
-			this.paintDataPoints((Graphics2D)graphics, this.dataSet.getTrainingSet());
-		}
-		if(showTestData){
-			this.paintTestSet((Graphics2D)graphics, this.dataSet.getTestSet());
-		}
-	}
-	
-	/**
-	 * Paints the test data points into the drawing area. For every test-data point we draw a grey circle around it. At this 
-	 * time the circle has just an arbitrary hardcoded radius but in the future the circle should be exactly as big such that
-	 * it contains for every test data point the k nearest neighbours of this data point.
-	 * @param g The graphics element in which we draw.
[…]

### 2013-01-26 exercise3/kNN/src/at/ac/tuwien/knn/gui/MainWindow.java

@@ -8,18 +8,13 @@ import java.awt.event.ItemListener;
 import java.awt.event.MouseAdapter;
 import java.awt.event.MouseEvent;
 
-import javax.swing.JButton;
-import javax.swing.JCheckBox;
-import javax.swing.JFileChooser;
-import javax.swing.JFrame;
-import javax.swing.JLabel;
-import javax.swing.JSpinner;
+import javax.swing.*;
 import javax.swing.border.LineBorder;
 import javax.swing.event.ChangeEvent;
 import javax.swing.event.ChangeListener;
 
-import at.ac.tuwien.knn.data.DataSet;
-import at.ac.tuwien.knn.data.DataSetParser;
+import at.ac.tuwien.knn.weka.WekaApi;
+import weka.core.Instances;
 
 
 public class MainWindow {
@@ -27,6 +22,7 @@ public class MainWindow {
 	private JFrame frame;
 	private DrawingArea panel;
 	private JLabel lbFilename;
+    private int defaultK = 3;
 
 	/**
 	 * Launch the application.
@@ -58,25 +54,39 @@ public class MainWindow {
 	 * Initialize the contents of the frame.
 	 */
 	private void initialize() {
+
+        // MAIN
+        //
 		frame = new JFrame();
 		frame.setBounds(100, 100, 1000, 800);
 		frame.setDefaultCloseOperation(JFrame.EXIT_ON_CLOSE);
 		frame.getContentPane().setLayout(null);
-		
-		JSpinner spinnerK = new JSpinner();
+
+        // SPINNER K
+        //
+		final JSpinner spinnerK = new JSpinner(new SpinnerNumberModel(defaultK, 1, 100, 1));
 		spinnerK.setBounds(827, 117, 45, 20);
+        spinnerK.addChangeListener(new ChangeListener() {
+            @Override
+            public void stateChanged(ChangeEvent e) {
+                panel.updateK((Integer)spinnerK.getModel().getValue());
+            }
+        });
 		frame.getContentPane().add(spinnerK);
-		
 		JLabel lblK = new JLabel("k:");
 		lblK.setBounds(810, 120, 16, 14);
 		frame.getContentPane().add(lblK);
-		
+
+        // DRAWING PANEL
+        //
 		panel = new DrawingArea();
 		panel.setBackground(Color.WHITE);
 		panel.setBorder(new LineBorder(SystemColor.controlDkShadow));
[…]

### 2013-01-26 exercise3/kNN/src/at/ac/tuwien/knn/gui/DrawingArea.java
finished first version. some stuff is missing

@@ -2,6 +2,7 @@ package at.ac.tuwien.knn.gui;
 
 import at.ac.tuwien.knn.data.DataSets;
 import at.ac.tuwien.knn.weka.WekaApi;
+import weka.classifiers.lazy.IBk;
 import weka.core.Instance;
 import weka.core.Instances;
 
@@ -23,6 +24,9 @@ public class DrawingArea extends JPanel {
     private DataSets dataSets;
     private Instances originalInstances;
 
+    // classifier
+    private IBk classifier;
+
     // drawing flags
     private boolean showTrainingData = true;
     private boolean showTestData = false;
@@ -30,6 +34,9 @@ public class DrawingArea extends JPanel {
     // coordinate and drawing stuff
     private double minX, maxX, minY, maxY;
     private int pointRadius = 5;
+    private int highlightRadius = 20;
+    private Color unknownColor = new Color(26, 26, 26);
+    private Color neighborConnectorColor = new Color(255, 163, 0);
     private static Color[] colors = new Color[]{new Color(230, 0, 0),//red
             new Color(0, 0, 230),//blue
             new Color(0, 220, 0),//green
@@ -44,14 +51,20 @@ public class DrawingArea extends JPanel {
     // CONTROL METHODS FOR USERS OF THIS COMPONENT (MAIN WINDOW)
     //
 
+    private void resetClassifier(int k) throws Exception {
+        classifier = WekaApi.getInstance().buildClassifierFromTrainingInstances(dataSets.getTrainingInstances(), k);
+    }
+
     /**
      * Update dataset file. Reload data, get a new classifier, re-everything.
+     *
      * @param dataFile
      * @throws Exception
      */
-    public void updateDataFile(File dataFile) throws Exception {
+    public void updateDataFile(File dataFile, Integer k) throws Exception {
         originalInstances = WekaApi.getInstance().loadData(dataFile);
         this.dataSets = WekaApi.getInstance().splitDataSet(originalInstances, 70);
+        resetClassifier(k);
         initColorMap(dataSets);
         calibrateDataSetsCoordinates(dataSets);
         repaint();
@@ -60,12 +73,12 @@ public class DrawingArea extends JPanel {
     /**
      * Update the value for k of kNN.
      * This builds a new classifier and restarts the classification.
-     * @param newK
+     *
+     * @param k
      */
-    public void updateK(Integer newK) {
+    public void updateK(Integer k) throws Exception {
         // create new classifier.
[…]

### 2013-01-26 exercise3/kNN/src/at/ac/tuwien/knn/gui/MainWindow.java

@@ -1,115 +1,114 @@
 package at.ac.tuwien.knn.gui;
 
-import java.awt.Color;
-import java.awt.EventQueue;
-import java.awt.SystemColor;
+import javax.swing.*;
+import javax.swing.border.LineBorder;
+import javax.swing.event.ChangeEvent;
+import javax.swing.event.ChangeListener;
+import java.awt.*;
 import java.awt.event.ItemEvent;
 import java.awt.event.ItemListener;
 import java.awt.event.MouseAdapter;
 import java.awt.event.MouseEvent;
 
-import javax.swing.*;
-import javax.swing.border.LineBorder;
-import javax.swing.event.ChangeEvent;
-import javax.swing.event.ChangeListener;
-
-import at.ac.tuwien.knn.weka.WekaApi;
-import weka.core.Instances;
-
 
 public class MainWindow {
 
-	private JFrame frame;
-	private DrawingArea panel;
-	private JLabel lbFilename;
+    private JFrame frame;
+    private DrawingArea panel;
+    private JLabel lbFilename;
     private int defaultK = 3;
+    private int defaultPercentage = 70;
 
-	/**
-	 * Launch the application.
-	 */
-	public static void main(String[] args) {
-		EventQueue.invokeLater(new Runnable() {
-			public void run() {
-				try {
-					MainWindow window = new MainWindow();
-					window.frame.setVisible(true);
-				} catch (Exception e) {
-					e.printStackTrace();
-				}
-			}
-		});
-		
-		//DrawingArea drawingArea = new DrawingArea();
-	}
+    /**
+     * Launch the application.
+     */
+    public static void main(String[] args) {
+        EventQueue.invokeLater(new Runnable() {
+            public void run() {
+                try {
+                    MainWindow window = new MainWindow();
+                    window.frame.setVisible(true);
+                } catch (Exception e) {
+                    e.printStackTrace();
+                }
+            }
+        });
 
-	/**
-	 * Create the application.
-	 */
-	public MainWindow() {
-		initialize();
-		
-	}
+        //DrawingArea drawingArea = new DrawingArea();
+    }
 
-	/**
-	 * Initialize the contents of the frame.
-	 */
-	private void initialize() {
+    /**
+     * Create the application.
+     */
+    public MainWindow() {
+        initialize();
+
+    }
+
+    /**
+     * Initialize the contents of the frame.
+     */
[…]

### 2013-01-26 exercise3/paper.tex
added initial paper.tex

--- 2013-01-07 exercise2/paper/paper.tex
@@
-%x%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Short Sectioned Assignment
@@
 
-\usepackage{color}
-\usepackage{float}
-\usepackage{placeins}
-\usepackage{natbib}
-% philipp: added for code listings
-\usepackage{listings}
-\lstset{
-    captionpos=b,
-    basicstyle=\small
-}
-\newcommand{\Hilight}{\makebox[0pt][l]{\color{red}\rule[-4pt]{0.65\linewidth}{14pt}}}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage{morefloats}
-\bibliographystyle{plain}
-
 \usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
@@
 \usepackage{lipsum} % Used for inserting dummy 'Lorem ipsum' text into the template
-\usepackage{todonotes}
 
 \usepackage{sectsty} % Allows customizing section commands
+
+\usepackage{todonotes}
 \allsectionsfont{\centering \normalfont\scshape} % Make all sections centered, the default font and small caps
@@
 \fancyfoot[R]{\thepage} % Page numbering for right footer
-% \renewcommand{\thesection}{Data Set \arabic{section}}
 \renewcommand{\headrulewidth}{0pt} % Remove header underlines
@@
 
-\setlength{\parskip}{\baselineskip}%
 \setlength\parindent{0pt} % Removes all indentation from paragraphs - comment this line for an assignment with lots of text
+
 %----------------------------------------------------------------------------------------
@@
 \normalfont \normalsize 
-\textsc{Vienna University of Technology} \\ [25pt] % Your university, school and/or department name(s)
+\textsc{university, school or department name} \\ [25pt] % Your university, school and/or department name(s)
 \horrule{0.5pt} \\[0.4cm] % Thin top horizontal rule
-\huge Experiments in Machine Learning 2 \\ % The assignment title
+\huge Assignment Title \\ % The assignment title
 \horrule{2pt} \\[0.5cm] % Thick bottom horizontal rule
@@
 
-\author{Benjamin Kiesl \and Philipp Steinwender \and Robert Sch\"{a}fer} % Your name
+\author{John Smith} % Your name
 
@@
 
-\tableofcontents
+\section{Our Task}
 
-%
-%
-%
-\section{TODO}
+You should pick a number of 2D (or optionally 3D) datasets (where the variables correspond to the coordinates in the 2D or 3D space), and demonstrate how different settings of the parameter k in k-NN modify the decision boundary.
+Your solution should work with arbitrary number of classes.
+Provide an interface that visualises the data set and its classes.
+In that interface, provide controls where the user can select a k value, and colour the points in the space according to which class they would get classified into.
+A simple approach is to iterate over all your points in the visualisation grid and ask the classifier for the classification of each point.
 
-\begin{itemize}
-\item Implementation of cutsom bn-searcher
-\item documentation of cutsom bn-searcher
-\item comparison and interpretation of the results
-\end{itemize}
+Finally, provide an option to the user to select a training/test set split, and indicate the classification of the test set in your plot.
+The data sets you provide should contain examples where low numbers of k are superior, as well as examples where higher number of k are more accurate.
 
-%
-%
-%
-\section{Preface}
+\section{Implementation}
 
-This assignment is about extending a Bayes Network Classifier by adding a custom search algorithm.
-We chose the Hill Climber search algorithm for an implementation of an iterative local search algorithm.
-
-%
-%
-%
-\section{Setup}
-
-There is a R script to execute the searcher in the handed in package at \texttt{R/bayesNetTest.R}.
-It loads Weka, builds the classifier, runs it and evaluates the model.
-It also shows the performance tables and the differences of the original Hill Climber implementation and our improved one.
-To run the script the following Tools need to be installed: R with the packages RWeka and rJava, Java (JRE).
-
-The used package RWeka is an interface in R that wraps the Weka framework and has a really good documentation.
-
-%
-%
-\section{BayesNet}
-
-First we take a look at the BayesNet classifier implementation.
-When the classifier is startet, first thing is to generate the classification model.
-Therefore the BayesNet classifier initializes an adjacency matrix (\texttt{initStructure()}) where each attribute has an array of parents.
-These arrays are empty after initialization.
-
-\begin{lstlisting}[caption={Initialization of BayesNet},label=initStructure]
-...
-// initialize parent sets to have arrow from classifier node to
-// each of the other nodes
-for (int iAttribute = 0; iAttribute < instances.numAttributes(); iAttribute++) {
-    if (iAttribute != iClass) {
-        bayesNet.getParentSet(iAttribute).addParent(iClass, instances);
-    }
-}
-...
-\end{lstlisting}
-
-The default Weka's SearchAlgorithm initializes the network like a Naive Bayes network.
-Listing \ref{initStructure} shows how we can use the adjacency matrix of BayesNet.
-Where \texttt{BayesNet} is the classifier that holds the network datastructure and \texttt{getParentSet(n)} returns the \texttt{nth} row in the adjacency matrix.
-
-%
-%
-%
-\section{Hill Climber}
-
-Implementation of Weka's Hill Climber algorithm.
-
-%
-%
-\subsection{Description}
-
-As shown in listing \ref{hcSearch}, Hill Climber calculates an optimal solution based on the current network iteratively until no opration can be found to improve the network or the new network is only slightly better than the previous one.
-To find the best operation all possible operations get evaluated and the best one is returned.
-The algorithm stops, when it finds a local optima.
-
-\begin{lstlisting}[caption={HillClimbers search algorithm},label=hcSearch]
-double fScore = calcScore(bayesNet);
-Operation oOperation = getOptimalOperation(bayesNet, instances);
-while ((oOperation != null) && (oOperation.m_fScore > fScore)) {
-	performOperation(bayesNet, instances, oOperation);
-	fScore = oOperation.m_fScore;
-	oOperation = getOptimalOperation(bayesNet, instances);
-}  
-\end{lstlisting}
-
-%
-%
-\subsection{Test Results}
-
-Two data sets are used to test the search algorithm. These are x and y from the assignment 1.
-
-\todo{insert results of test}
-
-As shown in the result listing, \todo{describe result} ....
-
-%
-%
-%
-\section{Extended Hill Climber}
-
-To improve the Hill Climber algorithm we used it with an iterated local search algorithm.
-
-\subsection{Description}
-
-Like the Hill Climber, our improved search algorithm introduces an initial Naive Bayes like network where each attribute has the classification attribute as parent.
-
-Then the Hill Climber algorithm is executed to find a local optima.
-To find other local optima, our search algorithm iteratively selects a neighbor of the found network and uses it as initial network for the Hill Climber to hopefully find better local optima than the first one.
-
- The intial network we use in each iteration is a random neighbor of the previously found one. It does not matter if it is better or worse.
-
-If the Hill Climber finds a better network it is stored as the current best found network and if the seacher cannot find a better network after a given number of iterations, the algorithm stops and yields the best solution found.
-
-%
-%
-%
-\section{Comparison}
-
-TODO:
-* used datasets
-* accuracy results
-* runtime/performance
-
-\bibliography{references}
-
+\section{Evaluation}
+\subsection{Data Sets}
+\todo[inline]{Wich data sets did we used?}
 \end{document}
-%------------------------------------------------

### 2013-01-26 exercise3/kNN/src/at/ac/tuwien/knn/gui/DrawingArea.java
ferdisch

@@ -8,6 +8,7 @@ import weka.core.Instances;
 
 import javax.swing.*;
 import java.awt.*;
+import java.awt.geom.Arc2D;
 import java.io.File;
 import java.util.HashMap;
 import java.util.Map;
@@ -33,10 +34,13 @@ public class DrawingArea extends JPanel {
 
     // coordinate and drawing stuff
     private double minX, maxX, minY, maxY;
+
     private int pointRadius = 5;
-    private int highlightRadius = 20;
-    private Color unknownColor = new Color(26, 26, 26);
+    private int highlightRadius = 15;
+
+    private Color highlightColor = new Color(230, 230, 230);
     private Color neighborConnectorColor = new Color(255, 163, 0);
+
     private static Color[] colors = new Color[]{new Color(230, 0, 0),//red
             new Color(0, 0, 230),//blue
             new Color(0, 220, 0),//green
@@ -63,7 +67,7 @@ public class DrawingArea extends JPanel {
      */
     public void updateDataFile(File dataFile, Integer k) throws Exception {
         originalInstances = WekaApi.getInstance().loadData(dataFile);
-        this.dataSets = WekaApi.getInstance().splitDataSet(originalInstances, 70);
+        dataSets = WekaApi.getInstance().splitDataSet(originalInstances, 70);
         resetClassifier(k);
         initColorMap(dataSets);
         calibrateDataSetsCoordinates(dataSets);
@@ -82,9 +86,15 @@ public class DrawingArea extends JPanel {
         repaint();
     }
 
+    /**
+     * Update the percentage value where to split the data set.
+     *
+     * @param percentage
+     * @throws Exception
+     */
     public void updatePercentage(Integer percentage) throws Exception {
         // make new split
-        this.dataSets = WekaApi.getInstance().splitDataSet(originalInstances, percentage);
+        dataSets = WekaApi.getInstance().splitDataSet(originalInstances, percentage);
         repaint();
     }
 
@@ -119,22 +129,28 @@ public class DrawingArea extends JPanel {
         super.paintComponent(graphics);
         try {
 
-            System.out.println("repaint");
-
             if (dataSets == null || dataSets.isEmpty()) {
-                System.out.println("no data set, return");
                 return;
             }
 
-            // paint all dots of the test set + highlighting ovals + connectors to the nearest neigbors
+            // order of how to paint all elements:
+            // test:
+            // - highlights of test data
+            // - knn connectors
[…]

### 2013-01-26 exercise3/kNN/src/at/ac/tuwien/knn/gui/MainWindow.java

@@ -9,6 +9,7 @@ import java.awt.event.ItemEvent;
 import java.awt.event.ItemListener;
 import java.awt.event.MouseAdapter;
 import java.awt.event.MouseEvent;
+import java.io.File;
 
 
 public class MainWindow {
@@ -53,9 +54,15 @@ public class MainWindow {
         // MAIN
         //
         frame = new JFrame();
-        frame.setBounds(100, 100, 1000, 800);
+        frame.setBounds(100, 100, 1000, 810);
         frame.setDefaultCloseOperation(JFrame.EXIT_ON_CLOSE);
         frame.getContentPane().setLayout(null);
+
+        JLabel jlDesc = new JLabel("Test data points are highlighted by a grey circle. The colors show different classes: (acutal : predicted)");
+        jlDesc.setBounds(35, 740, 800, 20);
+        frame.getContentPane().add(jlDesc);
+
+
 
         // SPINNER K
         //
@@ -166,5 +173,11 @@ public class MainWindow {
         chckbxShowTrainingData.setSelected(true);
         chckbxShowTrainingData.setBounds(810, 235, 152, 23);
         frame.getContentPane().add(chckbxShowTrainingData);
+
+        try {
+            panel.updateDataFile(new File("data/knn-data.arff"), 3);
+        } catch (Exception e) {
+            e.printStackTrace();
+        }
     }
 }

### 2013-01-26 exercise3/kNN/src/at/ac/tuwien/knn/gui/MainWindow.java
Merge branch 'master' of https://github.com/serprime/Machine-Learning-WS-2012
Conflicts:
	.gitignore

Signed-off-by: schubiduu <b.kiesl@atn.nu>

@@ -17,10 +17,22 @@ public class MainWindow {
     private JFrame frame;
     private DrawingArea panel;
     private JLabel lbFilename;
-    private int defaultK = 3;
+    private JLabel lblClasses;
+    
+	private int defaultK = 3;
     private int defaultPercentage = 70;
 
-    /**
+    
+    public JLabel getLblClasses() {
+		return lblClasses;
+	}
+    
+    public JFrame getFrame() {
+		return frame;
+	}
+
+
+	/**
      * Launch the application.
      */
     public static void main(String[] args) {
@@ -34,8 +46,6 @@ public class MainWindow {
                 }
             }
         });
-
-        //DrawingArea drawingArea = new DrawingArea();
     }
 
     /**
@@ -54,7 +64,7 @@ public class MainWindow {
         // MAIN
         //
         frame = new JFrame();
-        frame.setBounds(100, 100, 1000, 810);
+        frame.setBounds(100, 100, 1068, 810);
         frame.setDefaultCloseOperation(JFrame.EXIT_ON_CLOSE);
         frame.getContentPane().setLayout(null);
 
@@ -85,7 +95,7 @@ public class MainWindow {
 
         // DRAWING PANEL
         //
-        panel = new DrawingArea();
+        panel = new DrawingArea(this);
         panel.setBackground(Color.WHITE);
         panel.setBorder(new LineBorder(SystemColor.controlDkShadow));
         panel.setBounds(33, 33, 737, 699);
@@ -125,7 +135,7 @@ public class MainWindow {
         // FILE HANDLING
         //
         lbFilename = new JLabel("No dataset chosen");
-        lbFilename.setBounds(810, 73, 164, 20);
+        lbFilename.setBounds(810, 73, 232, 20);
         frame.getContentPane().add(lbFilename);
         JButton btnOpenFile = new JButton("Open file...");
         btnOpenFile.addMouseListener(new MouseAdapter() {
@@ -135,7 +145,7 @@ public class MainWindow {
                 int returnVal = fileChooser.showOpenDialog(frame);
                 if (returnVal == JFileChooser.APPROVE_OPTION) {
                     try {
-                        panel.updateDataFile(fileChooser.getSelectedFile(), (Integer) spinnerK.getModel().getValue());
+                        updateDataFile(fileChooser.getSelectedFile(), (Integer) spinnerK.getModel().getValue());
                         lbFilename.setText("Dataset: " + fileChooser.getSelectedFile().getName());
                     } catch (Exception e) {
[…]

### 2013-01-26 exercise3/kNN/src/at/ac/tuwien/knn/weka/WekaApi.java

@@ -1,6 +1,12 @@
 package at.ac.tuwien.knn.weka;
 
-import at.ac.tuwien.knn.data.DataSets;
+import java.awt.Color;
+import java.io.File;
+import java.util.Collection;
+import java.util.Enumeration;
+import java.util.HashSet;
+import java.util.Random;
+
 import weka.classifiers.lazy.IBk;
 import weka.core.Instance;
 import weka.core.Instances;
@@ -8,8 +14,7 @@ import weka.core.converters.ConverterUtils.DataSource;
 import weka.core.neighboursearch.LinearNNSearch;
 import weka.filters.Filter;
 import weka.filters.unsupervised.instance.RemovePercentage;
-
-import java.io.File;
+import at.ac.tuwien.knn.data.DataSets;
 
 
 /**
@@ -121,5 +126,18 @@ public class WekaApi {
         // get other part
         percentageFilter.setInvertSelection(inverse);
         return Filter.useFilter(instances, percentageFilter);
+    }
+    
+    public static Collection<Object> getAttributeValues(DataSets dataSets){
+    	Collection<Object> attributeValues = new HashSet<Object>();
+    	Enumeration<Object> testSetAttributes = dataSets.getTestInstances().classAttribute().enumerateValues();
+    	Enumeration<Object> trainingSetAttributes = dataSets.getTrainingInstances().classAttribute().enumerateValues();
+    	while(testSetAttributes.hasMoreElements()){
+        	attributeValues.add(testSetAttributes.nextElement());
+        }
+        while(trainingSetAttributes.hasMoreElements()){
+        	attributeValues.add(trainingSetAttributes.nextElement());
+        }
+    	return attributeValues;
     }
 }

### 2013-01-26 exercise3/kNN/src/at/ac/tuwien/knn/DistributionGenerator.java
added gauss data set generator

@@ -1,0 +1,52 @@
+package at.ac.tuwien.knn;
+
+import org.uncommons.maths.random.GaussianGenerator;
+
+import java.io.File;
+import java.io.FileNotFoundException;
+import java.io.FileOutputStream;
+import java.io.PrintStream;
+import java.util.Random;
+
+public class DistributionGenerator {
+
+    public static void main(String[] args) throws FileNotFoundException {
+        run();
+    }
+
+    public static void run() throws FileNotFoundException {
+        PrintStream out = null;
+        try {
+            File file = new File("data/knn-3-gauss.arff");
+            System.out.println(file.getAbsolutePath());
+            out = new PrintStream(new FileOutputStream(file));
+            System.setOut(out);
+
+            out.println("@RELATION knn-data\n" +
+                    "\n" +
+                    "@ATTRIBUTE x     NUMERIC\n" +
+                    "@ATTRIBUTE y     NUMERIC\n" +
+                    "@ATTRIBUTE class {1,2,3}\n" +
+                    "\n" +
+                    "@DATA" +
+                    "\n");
+
+            generateGaussFor(10, 0, 10, 0, "1", 30);
+            generateGaussFor(10, 0, 10, 30, "2", 30);
+            generateGaussFor(10, -30, 10, 0, "3", 30);
+        } finally {
+            out.close();
+        }
+
+
+    }
+
+    private static void generateGaussFor(int xWidth, int xCenter, int yWidth, int yCenter, String classLabel, int count) {
+        GaussianGenerator gg = new GaussianGenerator(0, 1, new Random());
+        for (int i = 0; i < count; i++) {
+            int x = (int) (gg.nextValue() * xWidth + xCenter);
+            int y = (int) (gg.nextValue() * yWidth + yCenter);
+            System.out.println(x + "," + y + "," + classLabel);
+        }
+    }
+}

### 2013-01-26 exercise3/kNN/src/at/ac/tuwien/knn/gui/MainWindow.java

@@ -1,5 +1,7 @@
 package at.ac.tuwien.knn.gui;
 
+import at.ac.tuwien.knn.DistributionGenerator;
+
 import javax.swing.*;
 import javax.swing.border.LineBorder;
 import javax.swing.event.ChangeEvent;
@@ -175,7 +177,8 @@ public class MainWindow {
         frame.getContentPane().add(chckbxShowTrainingData);
 
         try {
-            panel.updateDataFile(new File("data/knn-data.arff"), 3);
+            DistributionGenerator.run();
+            panel.updateDataFile(new File("data/knn-3-gauss.arff"), 3);
         } catch (Exception e) {
             e.printStackTrace();
         }

### 2013-01-26 exercise3/kNN/src/at/ac/tuwien/knn/weka/WekaApi.java

@@ -10,6 +10,7 @@ import weka.filters.Filter;
 import weka.filters.unsupervised.instance.RemovePercentage;
 
 import java.io.File;
+import java.util.Random;
 
 
 /**
@@ -76,6 +77,7 @@ public class WekaApi {
     public Instances loadData(File file) throws Exception {
         DataSource dataSource = new DataSource(file.getAbsolutePath());
         Instances instances = dataSource.getDataSet();
+        instances.randomize(new Random());
         instances.setClassIndex(instances.numAttributes() - 1);
         System.out.println("loaded data. class attribute: " + instances.classAttribute());
         return instances;

### 2013-01-26 exercise3/kNN/src/at/ac/tuwien/knn/gui/MainWindow.java
stuff

@@ -203,8 +203,8 @@ public class MainWindow {
         try {
             DistributionGenerator.run();
             this.updateDataFile(new File("data/knn-3-gauss.arff"), 3);
-            //this.updateDataFile(new File("data/knn-data.arff"), 3);
-            //this.updateDataFile(new File("data/knn_benjamin_data.arff"), 3);
+            // this.updateDataFile(new File("data/knn-data.arff"), 3);
+            // this.updateDataFile(new File("data/knn_benjamin_data.arff"), 3);
         } catch (Exception e) {
             e.printStackTrace();
         }

### 2013-01-26 exercise3/kNN/src/at/ac/tuwien/knn/DistributionGenerator.java
whatev

@@ -26,14 +26,14 @@ public class DistributionGenerator {
                     "\n" +
                     "@ATTRIBUTE x     NUMERIC\n" +
                     "@ATTRIBUTE y     NUMERIC\n" +
-                    "@ATTRIBUTE class {1,2,3}\n" +
+                    "@ATTRIBUTE class {welldone,medium,rare}\n" +
                     "\n" +
                     "@DATA" +
                     "\n");
 
-            generateGaussFor(10, 0, 10, 0, "1", 30);
-            generateGaussFor(10, 0, 10, 30, "2", 30);
-            generateGaussFor(10, -30, 10, 0, "3", 30);
+            generateGaussFor(10, 0, 10, 0, "welldone", 30);
+            generateGaussFor(10, 0, 10, 30, "medium", 30);
+            generateGaussFor(10, -30, 10, 0, "rare", 30);
         } finally {
             out.close();
         }

### 2013-01-26 exercise3/knnDemonstration.R
script checks whether both training and test date have the same schema

@@ -70,8 +70,14 @@ if( ncol(testing) != ncol(training)){
     stop(sprintf("The number of columns for the provided data sets is not the same: %s columns for trainingdata and %s columns for testdata.",  ncol(training), ncol(testing)))
 }
 
+for (i in 1:ncol(training)){
+    if (names(testing)[i] != names(training)[i]){
+        stop(sprintf("Please ensure that both training set and test data set have the same schema"))
+        }
+}
+
 if (modus == "2D"){
-    stop("Handling of 2 dimennsional plots is not implemented yet!")
+    stop("Handling of 2 dimensional plots is not implemented yet!")
 } else if (modus == "3D"){
     # for convenience, rename class column to "Class"
     names(training)[4] <- "Class"

### 2013-01-26 exercise3/kNN/src/at/ac/tuwien/knn/gui/MainWindow.java
einiges geändert

@@ -142,6 +142,7 @@ public class MainWindow {
             @Override
             public void mouseClicked(MouseEvent arg0) {
                 JFileChooser fileChooser = new JFileChooser();
+                fileChooser.setCurrentDirectory(new File("./data"));
                 int returnVal = fileChooser.showOpenDialog(frame);
                 if (returnVal == JFileChooser.APPROVE_OPTION) {
                     try {
@@ -206,7 +207,7 @@ public class MainWindow {
     }
     
     private void updateDataFile(File file, int k) throws Exception{
-    	panel.updateDataFile(new File("data/knn_benjamin_data.arff"), k);
+    	panel.updateDataFile(file, k);
     	this.lbFilename.setText("Open file: " + file.getName());
     }
 }

### 2013-01-26 exercise3/kNN/src/at/ac/tuwien/knn/gui/DrawingArea.java
some changes

@@ -1,6 +1,7 @@
 package at.ac.tuwien.knn.gui;
 
 import java.awt.Color;
+import java.awt.Component;
 import java.awt.Graphics;
 import java.awt.Graphics2D;
 import java.awt.Point;
@@ -8,8 +9,8 @@ import java.awt.RenderingHints;
 import java.awt.geom.Arc2D;
 import java.io.File;
 import java.util.Collection;
-import java.util.Enumeration;
 import java.util.HashMap;
+import java.util.LinkedList;
 import java.util.Map;
 import java.util.Random;
 
@@ -35,6 +36,7 @@ public class DrawingArea extends JPanel {
     // data, holds training and test data sets
     private DataSets dataSets;
     private Instances originalInstances;
+    private Collection<Component> legendElements = new LinkedList<Component>();
 
     // classifier
     private IBk classifier;
@@ -97,17 +99,22 @@ public class DrawingArea extends JPanel {
     	int i=0;
     	int vSpace = 25;
     	int hSpace = 10;
+    	for(Component c : this.legendElements){
+    		window.getFrame().remove(c);
+    	}
+    	this.legendElements.clear();
     	for(Object o : attributeValues){
     		JLabel lblClass = new JLabel(o.toString());
     		lblClass.setBounds(x + this.pointRadius*2 + hSpace, y + i*vSpace-2, 140, 14);
     		window.getFrame().getContentPane().add(lblClass);
+    		this.legendElements.add(lblClass);
     		ColorPoint colorPoint = new ColorPoint(this.classColorMap.get(o));
     		colorPoint.setLocation(x, y + i*vSpace);
     		colorPoint.setRadius(this.pointRadius);
     		window.getFrame().getContentPane().add(colorPoint);
+    		this.legendElements.add(colorPoint);
     		i++;
     	}
-
     }
 
     /**

### 2013-01-26 exercise3/kNN/src/at/ac/tuwien/knn/weka/WekaApi.java

@@ -1,12 +1,6 @@
 package at.ac.tuwien.knn.weka;
 
-import java.awt.Color;
-import java.io.File;
-import java.util.Collection;
-import java.util.Enumeration;
-import java.util.HashSet;
-import java.util.Random;
-
+import at.ac.tuwien.knn.data.DataSets;
 import weka.classifiers.lazy.IBk;
 import weka.core.Instance;
 import weka.core.Instances;
@@ -14,7 +8,12 @@ import weka.core.converters.ConverterUtils.DataSource;
 import weka.core.neighboursearch.LinearNNSearch;
 import weka.filters.Filter;
 import weka.filters.unsupervised.instance.RemovePercentage;
-import at.ac.tuwien.knn.data.DataSets;
+
+import java.io.File;
+import java.util.Collection;
+import java.util.Enumeration;
+import java.util.HashSet;
+import java.util.Random;
 
 
 /**
@@ -81,6 +80,7 @@ public class WekaApi {
     public Instances loadData(File file) throws Exception {
         DataSource dataSource = new DataSource(file.getAbsolutePath());
         Instances instances = dataSource.getDataSet();
+        instances.randomize(new Random());
         instances.setClassIndex(instances.numAttributes() - 1);
         System.out.println("loaded data. class attribute: " + instances.classAttribute());
         return instances;
@@ -127,17 +127,17 @@ public class WekaApi {
         percentageFilter.setInvertSelection(inverse);
         return Filter.useFilter(instances, percentageFilter);
     }
-    
-    public static Collection<Object> getAttributeValues(DataSets dataSets){
-    	Collection<Object> attributeValues = new HashSet<Object>();
-    	Enumeration<Object> testSetAttributes = dataSets.getTestInstances().classAttribute().enumerateValues();
-    	Enumeration<Object> trainingSetAttributes = dataSets.getTrainingInstances().classAttribute().enumerateValues();
-    	while(testSetAttributes.hasMoreElements()){
-        	attributeValues.add(testSetAttributes.nextElement());
+
+    public static Collection<Object> getAttributeValues(DataSets dataSets) {
+        Collection<Object> attributeValues = new HashSet<Object>();
+        Enumeration<Object> testSetAttributes = dataSets.getTestInstances().classAttribute().enumerateValues();
+        Enumeration<Object> trainingSetAttributes = dataSets.getTrainingInstances().classAttribute().enumerateValues();
+        while (testSetAttributes.hasMoreElements()) {
[…]

### 2013-01-26 exercise3/knnDemonstration.R
@wip: 2d plotting produces pdf output

@@ -77,11 +77,14 @@ for (i in 1:ncol(training)){
 }
 
 if (modus == "2D"){
-    stop("Handling of 2 dimensional plots is not implemented yet!")
+    # for convenience, rename class column to "Class"
+    names(training)[3] <- "Class"
+    names(testing)[3] <- "Class"
 } else if (modus == "3D"){
     # for convenience, rename class column to "Class"
     names(training)[4] <- "Class"
     names(testing)[4] <- "Class"
+}
 
     interface.draw <- function(panel) {
         classifier <-IBk(Class ~ ., training, control = Weka_control(K = as.numeric(panel$k)))
@@ -90,8 +93,11 @@ if (modus == "2D"){
         testing$Color <- (testing$Class == testing$P)
 
         # TODO: maybe output some evaluation data? e.g. percentage of misclassified instances
-
-        plot3d(testing[[1]], testing[[2]], testing[[3]], xlab=names(testing)[1], ylab=names(testing)[2], zlab=names(testing)[3], col=testing$Color, size=1, type='s')
+        if (modus == "3D"){
+            plot3d(testing[[1]], testing[[2]], testing[[3]], xlab=names(testing)[1], ylab=names(testing)[2], zlab=names(testing)[3], col=testing$Color, size=1, type='s')
+        } else if (modus == "2D"){
+            plot(testing[[1]], testing[[2]],xlab=names(testing)[1], ylab=names(testing)[2], col=testing$Color)
+        }
         panel
     }
     interface.panel <- rp.control("K nearest neighbor", k=opt$k)
@@ -100,4 +106,3 @@ if (modus == "2D"){
     rp.do(interface.panel, interface.draw)
     # user must close the panel to exit the program
     rp.block(interface.panel)
-}

### 2013-01-26 exercise3/knnDemonstration.R
Rscript is now able to color the points based on predicted class rather than the correctness of prediction

@@ -7,6 +7,7 @@ option_list <- list(
                     make_option(c("-t", "--train"), help="load training data", metavar="ARFF_FILE"),
                     make_option(c("-T", "--test"), help="load test data, if ommitted, training data is splitted into parts", metavar="ARFF_FILE"),
                     make_option(c("-k", "--kneighbors"), type="integer", help="k parameter for nearest neighbor classifier", default=as.integer(5), metavar="K"),
+                    make_option(c("-c", "--color"), type="character", default="Prediction", help="color the points based on the predicted value or based on the prediction correctness", metavar="Prediction|Correctness"),
                     make_option(c("--extract"), type="integer", help="reduce input data set by extracting every nth instance", metavar="N")
                     )
 parser <- OptionParser(usage="knnDemonstration [option] trainingfile", option_list=option_list)
@@ -26,6 +27,12 @@ if(file.access(opt$train) == -1){
 if( !is.null(opt$test) && (file.access(opt$test) == -1)){
     stop(sprintf("Specified file ( %s ) does not exist", opt$test))
 }
+
+# only values 'Prediction' or 'Correctness' as color values are allowed
+if (!opt$color %in% c("Prediction","Correctness")){
+    stop(sprintf("Expected 'Prediction' or 'Correctness' as 'color' parameter, but given %s", opt$color ))
+}
+
 
 # Weka bridge
 library(RWeka)
@@ -90,8 +97,11 @@ if (modus == "2D"){
         classifier <-IBk(Class ~ ., training, control = Weka_control(K = as.numeric(panel$k)))
         testing <- testing
         testing$P <- predict(classifier, testing)
-        testing$Color <- (testing$Class == testing$P)
-
+        if (opt$color == "Prediction"){
+            testing$Color <- testing$P
+        } else if (opt$color == "Correctness"){
+            testing$Color <- (testing$Class == testing$P)
+        }
         # TODO: maybe output some evaluation data? e.g. percentage of misclassified instances
         if (modus == "3D"){
             plot3d(testing[[1]], testing[[2]], testing[[3]], xlab=names(testing)[1], ylab=names(testing)[2], zlab=names(testing)[3], col=testing$Color, size=1, type='s')

### 2013-01-27 exercise3/knnDemonstration.R
refactoring of test/train data schema test added short version of parameter 'e'

@@ -8,7 +8,7 @@ option_list <- list(
                     make_option(c("-T", "--test"), help="load test data, if ommitted, training data is splitted into parts", metavar="ARFF_FILE"),
                     make_option(c("-k", "--kneighbors"), type="integer", help="k parameter for nearest neighbor classifier", default=as.integer(5), metavar="K"),
                     make_option(c("-c", "--color"), type="character", default="Prediction", help="color the points based on the predicted value or based on the prediction correctness", metavar="Prediction|Correctness"),
-                    make_option(c("--extract"), type="integer", help="reduce input data set by extracting every nth instance", metavar="N")
+                    make_option(c("-e","--extract"), type="integer", help="reduce input data set by extracting every nth instance", metavar="N")
                     )
 parser <- OptionParser(usage="knnDemonstration [option] trainingfile", option_list=option_list)
 arguments <- parse_args(parser, positional_arguments = TRUE)
@@ -77,10 +77,8 @@ if( ncol(testing) != ncol(training)){
     stop(sprintf("The number of columns for the provided data sets is not the same: %s columns for trainingdata and %s columns for testdata.",  ncol(training), ncol(testing)))
 }
 
-for (i in 1:ncol(training)){
-    if (names(testing)[i] != names(training)[i]){
-        stop(sprintf("Please ensure that both training set and test data set have the same schema"))
-        }
+if (!all(names(testing) == names(training))){
+    stop(sprintf("Please ensure that both training set and test data set have the same schema.\nThe training set looks like: %s \nAnd the test data looks like %s", paste(names(training),"",collapse=""),paste(names(testing),"",collapse="")))
 }
 
 if (modus == "2D"){
@@ -101,7 +99,7 @@ if (modus == "2D"){
             testing$Color <- testing$P
         } else if (opt$color == "Correctness"){
             testing$Color <- (testing$Class == testing$P)
-        }
+        } # TODO: does another switch makes sense? Maybe for 'Class', the true Class Attribute without prediction?
         # TODO: maybe output some evaluation data? e.g. percentage of misclassified instances
         if (modus == "3D"){
             plot3d(testing[[1]], testing[[2]], testing[[3]], xlab=names(testing)[1], ylab=names(testing)[2], zlab=names(testing)[3], col=testing$Color, size=1, type='s')

### 2013-01-27 exercise3/knnDemonstration.R
ensured that 2d plot will pop up a window

@@ -16,8 +16,7 @@ opt <- arguments$options
 
 # handle training data
 if(is.null(opt$train)){
-    cat("Please provide some training data.. \n")
-    stop()
+    stop("Please provide some training data.. ")
 }
 if(file.access(opt$train) == -1){
     stop(sprintf("Specified file ( %s ) does not exist", opt$input))
@@ -85,6 +84,14 @@ if (modus == "2D"){
     # for convenience, rename class column to "Class"
     names(training)[3] <- "Class"
     names(testing)[3] <- "Class"
+    # change default behaviour, plot to pdf, in plot to window
+    willFail <- tryCatch({
+        x11() # fails on windows
+        windows() # fails on linux
+        } , error = function(err){
+        err # Do nothing, just return the error
+        }
+    )
 } else if (modus == "3D"){
     # for convenience, rename class column to "Class"
     names(training)[4] <- "Class"

### 2013-01-27 exercise3/Paper/paper.tex
moved paper files into Paper/

--- 2013-01-26 exercise3/paper.tex
@@
 
+We used R as a wrapper for Weka and plotted the result with the native functionality in R.
+
 \section{Evaluation}
+Our Data Set were 
+\begin{enumerate}
+\item
+The 'Skin-NonSkin' Data Set, which is a collection of RGB-values and the 4th attribute states whether this is a skin color or not
+\item
+A truncation of the iris data set, with a removed column to fit the data set to 3 dimensions
+\end{enumerate}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{Skin_predicted_k5}
+\caption{The 'Skin-NonSkin' Data Set, colored by the predicted class attribute}
+\label{fig:skin:predicted:k5}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{Skin_correct_k5}
+\caption{The 'Skin-NonSkin' Data Set, colored by correct or misclassified instance}
+\label{fig:skin:correct:k5}
+\end{figure}
 \subsection{Data Sets}

### 2013-01-27 exercise3/Paper/paper.tex
amended commit, moved paper2 into paper

@@
 
+\usepackage{graphicx}
 \usepackage{todonotes}
+
 \allsectionsfont{\centering \normalfont\scshape} % Make all sections centered, the default font and small caps
@@
 \item
-A truncation of the iris data set, with a removed column to fit the data set to 3 dimensions
+A truncation of the iris data set, with a removed column to fit the data to 3 dimensions
+\item
+\todo[inline]{... a third data set}
 \end{enumerate}
-\begin{figure}[h]
-\centering
+
+The knn nearest neighbor performs especially well if we have a balanced data set with separated spaces where one of the class values is dominating.
+Figure \ref{fig:skin:predicted:k5} shows the prediction of K-Nearest-Neighbor for the 'Skin-Nonskin' data set.
+Obviously, there are a certain space where all the actual skin colors are located.
+This is an important fact, which facilitates the prediction of instances.
+In figure \ref{fig:skin:correct:k5} one can see the same data set visualized based on the correctness of the prediction.
+Black marks belong to correct instances and white marks belong to misclassified instances.
+In this example, k-nn performs very well, only few instances lying on the border to the skin color space are misclassified.
+
+\begin{figure}[\textwidth]
+    \begin{center}
         \includegraphics[width=\textwidth]{Skin_predicted_k5}
-\caption{The 'Skin-NonSkin' Data Set, colored by the predicted class attribute}
+    \end{center}
+    \caption['Skin-NonSkin' prediction with k=5]{The 'Skin-NonSkin' Data Set, colored by the predicted class attribute, with 5 neighbors as K parameter.}
     \label{fig:skin:predicted:k5}
@@
 
-\begin{figure}[h]
-\centering
+\begin{figure}[\textwidth]
+    \begin{center}
 \includegraphics[width=\textwidth]{Skin_correct_k5}
-\caption{The 'Skin-NonSkin' Data Set, colored by correct or misclassified instance}
+    \end{center}
+\caption['Skin-NonSkin' correctness with k=5]{The 'Skin-NonSkin' Data Set, colored by correct or misclassified instance, with 5 neighbors as K parameter.}
 \label{fig:skin:correct:k5}
 \end{figure}
+
+\begin{figure}[\textwidth]
+    \begin{center}
+\includegraphics[width=\textwidth]{Skin_predicted_k100}
+    \end{center}
+\caption['Skin-NonSkin' prediction with k=100]{The 'Skin-NonSkin' Data Set, colored by the predicted class attribute, with 100 neighbors as K parameter.}
+\label{fig:skin:predicted:k100}
+\end{figure}
+
+If we raise the K-paramter up to 100, we get worse results.
+As seen in figure \ref{fig:skin:predicted:k100} the K-NN algorithm misclassfies more instances that are located in one corner of the RGB space.
+Hence, we can see more white marks in figure \ref{fig:skin:correct:k100}.
+
+\begin{figure}[\textwidth]
+    \begin{center}
+\includegraphics[width=\textwidth]{Skin_correct_k100}
+    \end{center}
+\caption['Skin-NonSkin' correctness with k=100]{The 'Skin-NonSkin' Data Set, colored by correct or misclassified instance, with 100 neighbors as K parameter.}
+\label{fig:skin:correct:k100}
+\end{figure}
+
 \subsection{Data Sets}

### 2013-01-27 exercise3/Paper/paper.tex
minor changes

@@
 \normalfont \normalsize 
-\textsc{university, school or department name} \\ [25pt] % Your university, school and/or department name(s)
+\textsc{Vienna University of Technology} \\ [25pt] % Your university, school and/or department name(s)
 \horrule{0.5pt} \\[0.4cm] % Thin top horizontal rule
-\huge Assignment Title \\ % The assignment title
+\huge Demonstration of K-Nearest-Neighbor \\ % The assignment title
 \horrule{2pt} \\[0.5cm] % Thick bottom horizontal rule
@@
 
-\author{John Smith} % Your name
+\author{Benjamin Kiesl \and Philipp Steinwender \and Robert Schäfer} % Your name
 
@@
 
-If we raise the K-paramter up to 100, we get worse results.
+If we raise the K-parameter up to 100, we get worse results.
 As seen in figure \ref{fig:skin:predicted:k100} the K-NN algorithm misclassfies more instances that are located in one corner of the RGB space.
@@
 
-\subsection{Data Sets}
-\todo[inline]{Wich data sets did we used?}
 \end{document}

### 2013-01-27 exercise3/knnDemonstration.R
minor bugfix

@@ -19,7 +19,7 @@ if(is.null(opt$train)){
     stop("Please provide some training data.. ")
 }
 if(file.access(opt$train) == -1){
-    stop(sprintf("Specified file ( %s ) does not exist", opt$input))
+    stop(sprintf("Specified file ( %s ) does not exist", opt$train))
 }
 
 # handle test data

### 2013-01-27 exercise3/knnDemonstration.R
refactored command line parsing

@@ -4,15 +4,14 @@
 library("optparse")
 
 option_list <- list(
-                    make_option(c("-t", "--train"), help="load training data", metavar="ARFF_FILE"),
-                    make_option(c("-T", "--test"), help="load test data, if ommitted, training data is splitted into parts", metavar="ARFF_FILE"),
-                    make_option(c("-k", "--kneighbors"), type="integer", help="k parameter for nearest neighbor classifier", default=as.integer(5), metavar="K"),
-                    make_option(c("-c", "--color"), type="character", default="Prediction", help="color the points based on the predicted value or based on the prediction correctness", metavar="Prediction|Correctness"),
-                    make_option(c("-e","--extract"), type="integer", help="reduce input data set by extracting every nth instance", metavar="N")
+                    make_option(c("-t", "--train"), action="store", help="load training data", metavar="ARFF_FILE"),
+                    make_option(c("-T", "--test"), action="store", help="load test data, if ommitted, training data is splitted into parts", metavar="ARFF_FILE"),
+                    make_option(c("-k", "--kneighbors"), action="store", type="integer", help="k parameter for nearest neighbor classifier", default=as.integer(5), metavar="K"),
+                    make_option(c("-c", "--color"), action="store_true", default=FALSE, help="color the points based on the prediction correctness, otherwise on the predicted value"),
+                    make_option(c("-e","--extract"), action="store", type="integer", help="reduce input data set by extracting every nth instance", metavar="N")
                     )
 parser <- OptionParser(usage="knnDemonstration [option] trainingfile", option_list=option_list)
-arguments <- parse_args(parser, positional_arguments = TRUE)
-opt <- arguments$options
+opt <- parse_args(parser)
 
 # handle training data
 if(is.null(opt$train)){
@@ -26,12 +25,6 @@ if(file.access(opt$train) == -1){
 if( !is.null(opt$test) && (file.access(opt$test) == -1)){
     stop(sprintf("Specified file ( %s ) does not exist", opt$test))
 }
-
-# only values 'Prediction' or 'Correctness' as color values are allowed
-if (!opt$color %in% c("Prediction","Correctness")){
-    stop(sprintf("Expected 'Prediction' or 'Correctness' as 'color' parameter, but given %s", opt$color ))
-}
-
 
 # Weka bridge
 library(RWeka)
@@ -102,9 +95,9 @@ if (modus == "2D"){
         classifier <-IBk(Class ~ ., training, control = Weka_control(K = as.numeric(panel$k)))
[…]

### 2013-01-27 exercise3/knnDemonstration.R
added evaluation command line argument, to output the counts of the color values as a table

@@ -8,6 +8,7 @@ option_list <- list(
                     make_option(c("-T", "--test"), action="store", help="load test data, if ommitted, training data is splitted into parts", metavar="ARFF_FILE"),
                     make_option(c("-k", "--kneighbors"), action="store", type="integer", help="k parameter for nearest neighbor classifier", default=as.integer(5), metavar="K"),
                     make_option(c("-c", "--color"), action="store_true", default=FALSE, help="color the points based on the prediction correctness, otherwise on the predicted value"),
+                    make_option(c("-E", "--eval"), action="store_true", default=FALSE, help="output evaluation data"),
                     make_option(c("-e","--extract"), action="store", type="integer", help="reduce input data set by extracting every nth instance", metavar="N")
                     )
 parser <- OptionParser(usage="knnDemonstration [option] trainingfile", option_list=option_list)
@@ -34,6 +35,7 @@ library(rgl)
 
 # library for the text box
 library(rpanel)
+
 
 # helper function to extract a fraction amount of the training data
 Nth.rows<-function(dataframe, n)dataframe[(seq(n,to=nrow(dataframe),by=n)),]
@@ -100,7 +102,12 @@ if (modus == "2D"){
         } else {
             testing$Color <- (testing$Class == testing$P)
         } # TODO: does another switch makes sense? Maybe for 'Class', the true Class Attribute without prediction?
-        # TODO: maybe output some evaluation data? e.g. percentage of misclassified instances
+
+        # output some evaluation data? count of (mis)classified instances
+        if (opt$eval){
+        cat("\nThe distribution of color values is :\n")
+        print(table(testing$Color))
+        }
         if (modus == "3D"){
             plot3d(testing[[1]], testing[[2]], testing[[3]], xlab=names(testing)[1], ylab=names(testing)[2], zlab=names(testing)[3], col=testing$Color, size=1, type='s')
         } else if (modus == "2D"){

### 2013-01-27 exercise3/Paper/paper.tex
added haberman data set to paper further work on the paper

@@
 
-You should pick a number of 2D (or optionally 3D) datasets (where the variables correspond to the coordinates in the 2D or 3D space), and demonstrate how different settings of the parameter k in k-NN modify the decision boundary.
+You should pick a number of 2D (or optionally 3D) datasets (where the variables correspond to the coordinates in the 2D or 3D space), and demonstrate how different settings of the parameter k in K-NN modify the decision boundary.
 Your solution should work with arbitrary number of classes.
@@
 We used R as a wrapper for Weka and plotted the result with the native functionality in R.
+Our R-Script can handle 2- and 3-dimensional data as well as separate test and training data.
+It visualizes either the predicted instances of the test data itself or visualizes how many of them were classified correctly.
+After plotting the results, the user is asked to specify another k-parameter and the script runs the visualization once again with the new parameter.
 
 \section{Evaluation}
-Our Data Set were 
+Our Data Sets were 
 \begin{enumerate}
@@
 \item
-A truncation of the iris data set, with a removed column to fit the data to 3 dimensions
+    A truncation of the iris data set, where we have removed the 4th column to adjust the data to three dimensions.
 \item
-\todo[inline]{... a third data set}
+    The 'Haberman's Survival' data set, which contains cases from a study on the survival of patients who had undergone surgery for breast cancer.
 \end{enumerate}
 
-The knn nearest neighbor performs especially well if we have a balanced data set with separated spaces where one of the class values is dominating.
-Figure \ref{fig:skin:predicted:k5} shows the prediction of K-Nearest-Neighbor for the 'Skin-Nonskin' data set.
-Obviously, there are a certain space where all the actual skin colors are located.
+The K-NN nearest neighbor performs especially well if we have a balanced data set with separated spaces where one of the class values is dominating.
+Figure \ref{fig:skin:predicted:k5} shows the prediction of K-NN for the 'Skin-Nonskin' data set.
+Obviously, there is a certain space where all the actual skin colors are located.
 This is an important fact, which facilitates the prediction of instances.
@@
 Black marks belong to correct instances and white marks belong to misclassified instances.
-In this example, k-nn performs very well, only few instances lying on the border to the skin color space are misclassified.
+In this example, K-NN performs very well, only few instances located on the boundary between the two spaces are misclassified.
 
@@
 
+If we raise the K-parameter up to 100, we get worse results.
+As seen in figure \ref{fig:skin:predicted:k100} the K-NN algorithm misclassfies a couple of instances that are located in one corner of the RGB space.
+Hence, we can see more white marks in figure \ref{fig:skin:correct:k100}.
+
 \begin{figure}[\textwidth]
@@
 \end{figure}
-
-If we raise the K-parameter up to 100, we get worse results.
-As seen in figure \ref{fig:skin:predicted:k100} the K-NN algorithm misclassfies more instances that are located in one corner of the RGB space.
-Hence, we can see more white marks in figure \ref{fig:skin:correct:k100}.
 
@@
 
+We observe a similar result for the truncated 'Iris' data set.
+K-NN works very well except for some instances on the boundaries, figure \ref{fig:iris:correct:k5} shows us misclassfied instances exactly on the hyperplane between 'Iris versicolor' and 'Iris virginica'.
+The predicted instances can be seen in figure \ref{fig:iris:predicted:k5}.
+\begin{figure}[\textwidth]
+    \begin{center}
+\includegraphics[width=\textwidth]{Iris_predicted_k5}
+    \end{center}
+\caption['Iris' prediction with k=5]{The 'Iris' Data Set, colored by the predicted class attribute, with 5 neighbors as K parameter.}
+\label{fig:iris:predicted:k5}
+\end{figure}
+
+\begin{figure}[\textwidth]
+    \begin{center}
+\includegraphics[width=\textwidth]{Iris_correct_k5}
+    \end{center}
+\caption['Iris' correctness with k=5]{The 'Iris' Data Set, colored by correct or misclassified instance, with 5 neighbors as K parameter.}
+\label{fig:iris:correct:k5}
+\end{figure}
+
+Unlike the two previous data sets, the 'Haberman' data set does not contain clearly separable spaces where one of the class values dominates.
+Accordingly, the K-NN algorithm performs rather poor on this input set.
+Figure \ref{fig:haberman:correct:k13} shows the best distribution of correct and misclassified instances, namely 31 misclassified and 131 correctly classfied instances.
+It is notably that a k parameter of 13 is a rather high value.
+This fact could be interpreted in a sense that low k-values are preferable in case of easily separable data sets and higher k-values should be used for more overlapping, more scattered data sets.
+\begin{figure}[\textwidth]
+    \begin{center}
+\includegraphics[width=\textwidth]{Haberman_correct_k13}
+    \end{center}
+\caption['Haberman' correctness with k=13]{The 'Haberman' Data Set, colored by correct or misclassified instance, with 13 neighbors as K parameter.}
+\label{fig:haberman:correct:k13}
+\end{figure}
 \end{document}

### 2013-01-27 exercise3/Paper/paper.tex
added java stuff to paper

@@
 
-\author{Benjamin Kiesl \and Philipp Steinwender \and Robert Schäfer} % Your name
+\author{Benjamin Kiesl \and Philipp Steinwender \and Robert Sch\"{a}fer} % Your name
 
@@
 
+For this assignment we have two implementations: one in Java and a second one in R.
+
+\subsection{kNN visualization in Java}
+
+To visualize the decision boundaries and  classification results of weka's kNN implementation iBK a Java gui was implemented.
+We used Swing and plain Java.
+There is also a dataset generator for multiple gaussian cluster that can only be executed after editing the code and recompilation of the application.
+
+To run the application, execute the \texttt{kNN/run.sh} script. There is also a version for windows that was not tested.
+
+\begin{figure}[\textwidth]
+    \begin{center}
+\includegraphics[width=\textwidth]{knn-visualization-marker}
+    \end{center}
+\caption{Java GUI}
+\label{fig:javagui}
+\end{figure}
+
+Figure \ref{fig:javagui} shows how the gui looks like and which controls there are for manipulating the parameters of the data set and the classifier.
+The drawing area on the left displays the data points on a two dimensional coordinate system.
+The coordinate values of the data set get scaled to fit this area.
+The two markers showing the numbers 1 and 2 were added after taking the screenshot and highlight the following features:
+
+\begin{enumerate}
+\item With these 4 blue data points can be described how we visualize the decision boundaries of knn.
+The blue point in the middle has a light grey bigger circle around it.
+Points that look like this are from the test set.
+Data point that do not have this light grey highlighting are data points from the training set.
+Each data point of the test set has \texttt{k} connectors to surrounding data points which are the k nearest neighbors of the point.
+\item Here we can see a half red half green data point of the test set.
+It is connected with 2 green data points and a blue one.
+Therefore the classifier predicts its class as green.
+The predicted class can be seen in the right half of the data point.
+The left half shows which class the data point acutally has which is red.
+If theclassifier predicts the true value of a data point, both halfs have the same color.
+\end{enumerate}
+
+On the right side of the gui is a control panel. From top to bottom:
+
+\begin{itemize}
+\item The "open file" button can be used to read an arff file from disk.
+After selecting a file, it's content is loaded, a classifier is created and the result gets drawn.
+Below is a label that displays the name of the currently selected file.
+\item The spinner with the label \texttt{k} can be used to modify the classifier's value of \texttt{k}.
+\item The training-test-split can be modified with the spinner showing the number 70 whereby this means that we do a 70\% split and use 70\% for training and the rest for classification.
+Although kNN does not train a model, this parameter tells the classifier how much unclassified instances it should use.
+\item The next three checkboxes can be used to view or hide specific visualization details.
+One can hide the data points of the training data, those of the test set and hide the connectors between the test data points and their nearest neighbors.
+\item Then there is a legend telling the attribute names and which coloreach class is assigned to.
+\end{itemize}
+
+All controls have a ChangeListener that redraws the data points and the results immediately after a value got changed by the user.
+
+When loading a new file, the data is read by weka's arff parser.
+The created \texttt{Instances} are split by the specified percentage using the \texttt{RemovePercentageFilter}.
+Before the spilt, the data is shuffled.
+Then the \texttt{iBK} classifier gets instantiated, configured and the data points are drawn.
+
+With this visualization one can easily recognize how the classifier predicts the class of a new instance and how the values of k and the train-test-set-split can change runtime behaviour.
+
+\subsection{kNN visualization in R}
+
 We used R as a wrapper for Weka and plotted the result with the native functionality in R.
@@
 
-\section{Evaluation}
+\subsubsection{Evaluation}
 Our Data Sets were 

### 2013-01-27 exercise3/kNN/src/at/ac/tuwien/knn/gui/DrawingArea.java

@@ -137,9 +137,10 @@ public class DrawingArea extends JPanel {
      * @param percentage
      * @throws Exception
      */
-    public void updatePercentage(Integer percentage) throws Exception {
+    public void updatePercentage(Integer percentage, Integer k) throws Exception {
         // make new split
         dataSets = WekaApi.getInstance().splitDataSet(originalInstances, percentage);
+        resetClassifier(k);
         repaint();
     }
 
@@ -199,9 +200,9 @@ public class DrawingArea extends JPanel {
 
             if (showTestData) {
                 paintTestSetHighlights((Graphics2D) graphics, dataSets.getTestInstances());
-                if(showConnections){
-                	paintTestSetKnnConnectors((Graphics2D) graphics, dataSets.getTestInstances());
-                }
+            }
+            if(showConnections){
+                paintTestSetKnnConnectors((Graphics2D) graphics, dataSets.getTestInstances());
             }
             if (showTrainingData) {
                 paintDataPoints((Graphics2D) graphics, dataSets.getTrainingInstances());

### 2013-01-27 exercise3/R/3dplot.R
cleaned folders. finished java part

@@ -1,0 +1,22 @@
+#!/usr/bin/Rscript
+
+# plotting library
+library(rgl)
+
+# library to wait for user input
+library(tcltk)
+
+Nth.delete<-function(dataframe, n)dataframe[(seq(n,to=nrow(dataframe),by=n)),]
+m <- read.csv(file="Skin_NonSkin.csv", sep = "\t", head=FALSE)
+# We choose every 1000 row as an instance for our sample set
+sample <- Nth.delete(m,1000)
+# Renaming of columns
+names(sample)[names(sample)=="V1"] <- "B"
+names(sample)[names(sample)=="V2"] <- "G"
+names(sample)[names(sample)=="V3"] <- "R"
+names(sample)[names(sample)=="V4"] <- "Class"
+
+
+plot3d(sample$R,sample$G,sample$B, col=sample$Class, size=2, type='s')
+
+tk_messageBox(message = "Press a key to exit")

### 2013-01-27 exercise3/R/knnDemonstration.R

@@ -1,0 +1,123 @@
+#!/usr/bin/Rscript
+
+# library for option parsing
+library("optparse")
+
+option_list <- list(
+                    make_option(c("-t", "--train"), action="store", help="load training data", metavar="ARFF_FILE"),
+                    make_option(c("-T", "--test"), action="store", help="load test data, if ommitted, training data is splitted into parts", metavar="ARFF_FILE"),
+                    make_option(c("-k", "--kneighbors"), action="store", type="integer", help="k parameter for nearest neighbor classifier", default=as.integer(5), metavar="K"),
+                    make_option(c("-c", "--color"), action="store_true", default=FALSE, help="color the points based on the prediction correctness, otherwise on the predicted value"),
+                    make_option(c("-E", "--eval"), action="store_true", default=FALSE, help="output evaluation data"),
+                    make_option(c("-e","--extract"), action="store", type="integer", help="reduce input data set by extracting every nth instance", metavar="N")
+                    )
+parser <- OptionParser(usage="knnDemonstration [option] trainingfile", option_list=option_list)
+opt <- parse_args(parser)
+
+# handle training data
+if(is.null(opt$train)){
+    stop("Please provide some training data.. ")
+}
+if(file.access(opt$train) == -1){
+    stop(sprintf("Specified file ( %s ) does not exist", opt$train))
+}
+
+# handle test data
+if( !is.null(opt$test) && (file.access(opt$test) == -1)){
+    stop(sprintf("Specified file ( %s ) does not exist", opt$test))
+}
+
+# Weka bridge
+library(RWeka)
+
+# plotting library
+library(rgl)
+
+# library for the text box
+library(rpanel)
+
+
+# helper function to extract a fraction amount of the training data
+Nth.rows<-function(dataframe, n)dataframe[(seq(n,to=nrow(dataframe),by=n)),]
+
+
+inputArff <- read.arff(opt$train)
+# If the user specified --extract, we choose every nth row as an instance for our sample set
+if (!(is.null(opt$extract))){
+    sample <-Nth.rows(inputArff, opt$extract)
+} else {
+    sample <- inputArff
+}
+
+# if the user didn't specify a test data set, the training data is splitted
+if (is.null(opt$test)){
+    splits <- split(sample, c("trainset", "testset"))
+    training <- splits$trainset
+    testing <- splits$testset
+} else {
+    training <- sample
+    testing <- read.arff(opt$test)
+}
+
+if ( ncol(training) == 3 ){
[…]

### 2013-01-27 exercise3/knn-java-gui/run.sh

@@ -1,0 +1,1 @@
+java -cp out/production/KNN:lib/uncommons-maths-1.2.3.jar:lib/weka.jar:lib/weka-src.jar at.ac.tuwien.knn.gui.MainWindow

### 2013-01-27 exercise3/knn-java-gui/src/at/ac/tuwien/knn/gui/DrawingArea.java

@@ -1,0 +1,388 @@
+package at.ac.tuwien.knn.gui;
+
+import java.awt.Color;
+import java.awt.Component;
+import java.awt.Graphics;
+import java.awt.Graphics2D;
+import java.awt.Point;
+import java.awt.RenderingHints;
+import java.awt.geom.Arc2D;
+import java.io.File;
+import java.util.Collection;
+import java.util.HashMap;
+import java.util.LinkedList;
+import java.util.Map;
+import java.util.Random;
+
+import javax.swing.JLabel;
+import javax.swing.JPanel;
+
+import weka.classifiers.lazy.IBk;
+import weka.core.Instance;
+import weka.core.Instances;
+import at.ac.tuwien.knn.data.DataSets;
+import at.ac.tuwien.knn.weka.WekaApi;
+
+/**
+ * The panel in which all the data points are drawn
+ *
+ * @author BK
+ */
+public class DrawingArea extends JPanel {
+
+	//MainWindow to render things into it
+	private MainWindow window;
+	
+    // data, holds training and test data sets
+    private DataSets dataSets;
+    private Instances originalInstances;
+    private Collection<Component> legendElements = new LinkedList<Component>();
+
+    // classifier
+    private IBk classifier;
+
+    // drawing flags
+    private boolean showTrainingData = true;
+    private boolean showTestData = false;
+    private boolean showConnections = true;
+
+    // coordinate and drawing stuff
+    private double minX, maxX, minY, maxY;
+
+    private int pointRadius = 5;
+    private int highlightRadius = 15;
+
+    private Color highlightColor = new Color(230, 230, 230);
+    private Color neighborConnectorColor = new Color(255, 163, 0);
+
+    private static Color[] colors = new Color[]{new Color(230, 0, 0),//red
+            new Color(0, 0, 230),//blue
+            new Color(0, 220, 0),//green
+            new Color(102, 0, 102),//violet
+            new Color(255, 153, 0),//orange
+            new Color(255, 255, 100)//beige
+    };
+    private Map<Integer, Color> classColorMap = null;
+
+    public DrawingArea(MainWindow window){
+    	this.window = window;
+    }
+
+    //
+    // CONTROL METHODS FOR USERS OF THIS COMPONENT (MAIN WINDOW)
+    //
+
+    private void resetClassifier(int k) throws Exception {
[…]

### 2013-01-27 exercise3/knn-java-gui/run.sh
wooot

@@ -1,1 +1,1 @@
-java -cp out/production/KNN:lib/uncommons-maths-1.2.3.jar:lib/weka.jar:lib/weka-src.jar at.ac.tuwien.knn.gui.MainWindow
+java -cp bin:lib/uncommons-maths-1.2.3.jar:lib/weka.jar at.ac.tuwien.knn.gui.MainWindow

### 2013-01-27 exercise3/3dplot.R
woooooooot

@@ -1,22 +1,0 @@
-#!/usr/bin/Rscript
-
-# plotting library
-library(rgl)
-
-# library to wait for user input
-library(tcltk)
-
-Nth.delete<-function(dataframe, n)dataframe[(seq(n,to=nrow(dataframe),by=n)),]
-m <- read.csv(file="Skin_NonSkin.csv", sep = "\t", head=FALSE)
-# We choose every 1000 row as an instance for our sample set
-sample <- Nth.delete(m,1000)
-# Renaming of columns
-names(sample)[names(sample)=="V1"] <- "B"
-names(sample)[names(sample)=="V2"] <- "G"
-names(sample)[names(sample)=="V3"] <- "R"
-names(sample)[names(sample)=="V4"] <- "Class"
-
-
-plot3d(sample$R,sample$G,sample$B, col=sample$Class, size=2, type='s')
-
-tk_messageBox(message = "Press a key to exit")

### 2013-01-27 exercise3/kNN/src/at/ac/tuwien/knn/DistributionGenerator.java

@@ -1,52 +1,0 @@
-package at.ac.tuwien.knn;
-
-import org.uncommons.maths.random.GaussianGenerator;
-
-import java.io.File;
-import java.io.FileNotFoundException;
-import java.io.FileOutputStream;
-import java.io.PrintStream;
-import java.util.Random;
-
-public class DistributionGenerator {
-
-    public static void main(String[] args) throws FileNotFoundException {
-        run();
-    }
-
-    public static void run() throws FileNotFoundException {
-        PrintStream out = null;
-        try {
-            File file = new File("data/knn-3-gauss.arff");
-            System.out.println(file.getAbsolutePath());
-            out = new PrintStream(new FileOutputStream(file));
-            System.setOut(out);
-
-            out.println("@RELATION knn-data\n" +
-                    "\n" +
-                    "@ATTRIBUTE x     NUMERIC\n" +
-                    "@ATTRIBUTE y     NUMERIC\n" +
-                    "@ATTRIBUTE class {welldone,medium,rare}\n" +
-                    "\n" +
-                    "@DATA" +
-                    "\n");
-
-            generateGaussFor(10, 0, 10, 0, "welldone", 30);
-            generateGaussFor(10, 0, 10, 30, "medium", 30);
-            generateGaussFor(10, -30, 10, 0, "rare", 30);
-        } finally {
-            out.close();
-        }
-
-
-    }
-
-    private static void generateGaussFor(int xWidth, int xCenter, int yWidth, int yCenter, String classLabel, int count) {
-        GaussianGenerator gg = new GaussianGenerator(0, 1, new Random());
-        for (int i = 0; i < count; i++) {
-            int x = (int) (gg.nextValue() * xWidth + xCenter);
-            int y = (int) (gg.nextValue() * yWidth + yCenter);
-            System.out.println(x + "," + y + "," + classLabel);
-        }
-    }
-}

### 2013-01-27 exercise3/kNN/src/at/ac/tuwien/knn/data/DataSets.java

@@ -1,40 +1,0 @@
-package at.ac.tuwien.knn.data;
-
-import weka.core.Instances;
-
-public class DataSets {
-
-    private Instances trainingInstances;
-    private Instances testInstances;
-
-    public boolean isEmpty(){
-        return trainingInstances.isEmpty();
-    }
-
-    public Instances getTrainingInstances() {
-        return trainingInstances;
-    }
-
-    public void setTrainingInstances(Instances trainingInstances) {
-        this.trainingInstances = trainingInstances;
-    }
-
-    public Instances getTestInstances() {
-        return testInstances;
-    }
-
-    public void setTestInstances(Instances testInstances) {
-        this.testInstances = testInstances;
-    }
-
-    /**
-     * determine the number of classes that appear in in training and test set
-     * @return
-     */
-    public int numClasses() {
-        if (testInstances.numClasses() > trainingInstances.numClasses()) {
-            return testInstances.numClasses();
-        }
-        return trainingInstances.numClasses();
-    }
-}

### 2013-01-27 exercise3/kNN/src/at/ac/tuwien/knn/gui/MainWindow.java

@@ -1,233 +1,0 @@
-package at.ac.tuwien.knn.gui;
-
-import at.ac.tuwien.knn.DistributionGenerator;
-
-import javax.swing.*;
-import javax.swing.border.LineBorder;
-import javax.swing.event.ChangeEvent;
-import javax.swing.event.ChangeListener;
-import java.awt.*;
-import java.awt.event.ItemEvent;
-import java.awt.event.ItemListener;
-import java.awt.event.MouseAdapter;
-import java.awt.event.MouseEvent;
-import java.io.File;
-
-
-public class MainWindow {
-
-    private JFrame frame;
-    private DrawingArea panel;
-    private JLabel lbFilename;
-    private JLabel lblClasses;
-    private JLabel lblXAxis;
-    private JLabel lblYAxis;
-    
-	private int defaultK = 3;
-    private int defaultPercentage = 70;
-
-    public JLabel getLblXAxis() {
-		return lblXAxis;
-	}
-
-	public void setLblXAxis(JLabel lblXAxis) {
-		this.lblXAxis = lblXAxis;
-	}
-
-	public JLabel getLblYAxis() {
-		return lblYAxis;
-	}
-
-	public void setLblYAxis(JLabel lblYAxis) {
-		this.lblYAxis = lblYAxis;
-	}
-
-	public JLabel getLblClasses() {
-		return lblClasses;
-	}
-    
-    public JFrame getFrame() {
-		return frame;
-	}
-
-
-	/**
-     * Launch the application.
-     */
-    public static void main(String[] args) {
-        EventQueue.invokeLater(new Runnable() {
-            public void run() {
-                try {
-                    MainWindow window = new MainWindow();
-                    window.frame.setVisible(true);
-                } catch (Exception e) {
-                    e.printStackTrace();
-                }
-            }
-        });
-    }
-
-    /**
-     * Create the application.
-     */
-    public MainWindow() {
-        initialize();
-
-    }
-
-    /**
-     * Initialize the contents of the frame.
-     */
-    private void initialize() {
-        // MAIN
-        //
-        frame = new JFrame();
-        frame.setBounds(100, 100, 1068, 810);
-        frame.setDefaultCloseOperation(JFrame.EXIT_ON_CLOSE);
-        frame.getContentPane().setLayout(null);
-
-        JLabel jlDesc = new JLabel("Test data points are highlighted by a grey circle. The colors show different classes: (actual : predicted)");
[…]

### 2013-01-27 exercise3/kNN/src/at/ac/tuwien/knn/weka/WekaApi.java

@@ -1,143 +1,0 @@
-package at.ac.tuwien.knn.weka;
-
-import at.ac.tuwien.knn.data.DataSets;
-import weka.classifiers.lazy.IBk;
-import weka.core.Instance;
-import weka.core.Instances;
-import weka.core.converters.ConverterUtils.DataSource;
-import weka.core.neighboursearch.LinearNNSearch;
-import weka.filters.Filter;
-import weka.filters.unsupervised.instance.RemovePercentage;
-
-import java.io.File;
-import java.util.Collection;
-import java.util.Enumeration;
-import java.util.HashSet;
-import java.util.Random;
-
-
-/**
- * This is a stateless api class that is an interface to the weka library.
- */
-public class WekaApi {
-
-    /**
-     * private ctor to disable multiple instances
-     */
-    private WekaApi() {
-    }
-
-    /**
-     * static var that holds the one and only instance of this singleton
-     */
-    private static WekaApi instance = null;
-
-    public static WekaApi getInstance() {
-        if (instance == null) {
-            instance = new WekaApi();
-        }
-        return instance;
-    }
-
-    /**
-     * Give me an instance, i predict its class.
-     *
-     * @param classifier
-     * @param instance
-     * @throws Exception
-     */
-    public void classifyInstance(IBk classifier, Instance instance) throws Exception {
-        double[] probabilities = classifier.distributionForInstance(instance);
-        for (double probability : probabilities) {
-            System.out.println(probability);
-        }
-    }
-
-    /**
-     * takes Instances of trainings data and builds a knn classifier.
-     *
-     * @param instances training data
-     * @param k         main parameter for kNN
-     * @throws Exception kaboom
-     */
-    public IBk buildClassifierFromTrainingInstances(Instances instances, int k) throws Exception {
-        IBk knn = new IBk(k);
-
-        LinearNNSearch nearestNeighbourSearchAlgorithm = new LinearNNSearch();
-        knn.setNearestNeighbourSearchAlgorithm(nearestNeighbourSearchAlgorithm);
-
-        knn.buildClassifier(instances);
-        return knn;
-    }
-
-    /**
-     * Loads arff data set using the parser of weka.
-     *
-     * @param file file to read data from
-     * @return Instances
-     * @throws Exception if something goes wrong
-     */
-    public Instances loadData(File file) throws Exception {
-        DataSource dataSource = new DataSource(file.getAbsolutePath());
-        Instances instances = dataSource.getDataSet();
-        instances.randomize(new Random());
[…]

### 2013-01-27 exercise3/knnDemonstration.R

@@ -1,123 +1,0 @@
-#!/usr/bin/Rscript
-
-# library for option parsing
-library("optparse")
-
-option_list <- list(
-                    make_option(c("-t", "--train"), action="store", help="load training data", metavar="ARFF_FILE"),
-                    make_option(c("-T", "--test"), action="store", help="load test data, if ommitted, training data is splitted into parts", metavar="ARFF_FILE"),
-                    make_option(c("-k", "--kneighbors"), action="store", type="integer", help="k parameter for nearest neighbor classifier", default=as.integer(5), metavar="K"),
-                    make_option(c("-c", "--color"), action="store_true", default=FALSE, help="color the points based on the prediction correctness, otherwise on the predicted value"),
-                    make_option(c("-E", "--eval"), action="store_true", default=FALSE, help="output evaluation data"),
-                    make_option(c("-e","--extract"), action="store", type="integer", help="reduce input data set by extracting every nth instance", metavar="N")
-                    )
-parser <- OptionParser(usage="knnDemonstration [option] trainingfile", option_list=option_list)
-opt <- parse_args(parser)
-
-# handle training data
-if(is.null(opt$train)){
-    stop("Please provide some training data.. ")
-}
-if(file.access(opt$train) == -1){
-    stop(sprintf("Specified file ( %s ) does not exist", opt$train))
-}
-
-# handle test data
-if( !is.null(opt$test) && (file.access(opt$test) == -1)){
-    stop(sprintf("Specified file ( %s ) does not exist", opt$test))
-}
-
-# Weka bridge
-library(RWeka)
-
-# plotting library
-library(rgl)
-
-# library for the text box
-library(rpanel)
-
-
-# helper function to extract a fraction amount of the training data
-Nth.rows<-function(dataframe, n)dataframe[(seq(n,to=nrow(dataframe),by=n)),]
-
-
-inputArff <- read.arff(opt$train)
-# If the user specified --extract, we choose every nth row as an instance for our sample set
-if (!(is.null(opt$extract))){
-    sample <-Nth.rows(inputArff, opt$extract)
-} else {
-    sample <- inputArff
-}
-
-# if the user didn't specify a test data set, the training data is splitted
-if (is.null(opt$test)){
-    splits <- split(sample, c("trainset", "testset"))
-    training <- splits$trainset
-    testing <- splits$testset
-} else {
-    training <- sample
-    testing <- read.arff(opt$test)
-}
-
-if ( ncol(training) == 3 ){
[…]

### 2013-01-27 exercise3/paper/paper.tex

--- 2013-01-27 exercise3/Paper/paper.tex
@@
 
-To run the application, execute the \texttt{kNN/run.sh} script. There is also a version for windows that was not tested.
+To run the application, execute the \texttt{knn-java-gui/run.sh} script. There is also a script to compile the source \texttt{knn-java-gui/compile.sh}. Both scripts also have versions for windows.
 
